### 2022-07-28 others/journal submissions/journal of futures markets/cover letter.qmd
#16 gather references from previous issues

---
title: "cover letter"
format: docx
editor: visual
---

## References

### Co-movement

-   [prokopczuk](https://doi.org/10.1002/fut.22222 "The dynamics of commodity return comovements")[\_dynamics_2021](https://doi.org/10.1002/fut.22222)

### Financialisation

-   [conghui](https://doi.org/10.1002/fut.22127 "Liquidity shocks, commodity financialization, and market comovements")[\_liquidity_2020](https://doi.org/10.1002/fut.22127)
-   [kupabado](https://doi.org/10.1002/fut.22269 "Financialization, common stochastic trends, and commodity prices")[\_financialisation_2021](https://doi.org/10.1002/fut.22269)
-   [yin](https://doi.org/10.1002/fut.22099 "Intermediary asset pricing in commodity futures returns")[\_intermediary_2020](https://doi.org/10.1002/fut.22099)

### Fundamentals

-   [liu_market_2020](https://doi.org/10.1002/fut.22115 "The market quality of commodity futures markets")

#### Risk hedging

-   [isleimeyyeh_role_2020](https://doi.org/10.1002/fut.22122 "The role of financial investors in determining the commodity futures risk premium")

-   [ahmadi](https://doi.org/10.1002/fut.22113 "The theory of storage in the crude oil futures market, the role of financial conditions")[\_theory_2020](https://doi.org/10.1002/fut.22113)

-   [fan_speculative_2020](https://doi.org/10.1002/fut.22085 "Speculative pressure")

#### Price formation

-   [adammer_price_2018](https://doi.org/10.1002/fut.21891 "Price discovery dynamics in European agricultural markets")

-   [bohmann_price_2019](https://doi.org/10.1002/fut.22021 "Price discovery in commodity derivatives: Speculation or hedging?")

-   [llorente_trading_2020](https://doi.org/10.1002/fut.22079 "Trading and information in futures markets")

    ###