### 2024-06-11 others/paper/custom/LaTeX/main.tex

\documentclass[11pt, letterpaper, doublespacing]{article}

\usepackage{setspace}

\makeatletter

\renewcommand{\@seccntformat}[1]{\@nameuse{the#1}.~{}}

\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\Alph{subsection}}
\renewcommand{\thesubsubsection}{\arabic{subsubsection}}

\renewcommand{\section}{\@startsection {section}{1}{\z@}%
  {-3.5ex \@plus -1ex \@minus -.2ex}%
  {2.3ex \@plus.2ex}%
  {\centering\normalfont\large\bfseries}}

\renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
  {-3.25ex\@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\normalfont\large\slshape}}

\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
  {-3.25ex\@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\centering\normalfont\slshape}}

\makeatother

% \newtheoremstyle{plain}{9pt}{9pt}{\itshape}{0pt}{\scshape}{:}{3pt}{}

% \theoremstyle{plain}

% \newtheorem{assumption}{Assumption}
% \newtheorem{corollary}{Corollary}
% \newtheorem{definition}{Definition}
% \newtheorem{lemma}{Lemma}  
% \newtheorem{proposition}{Proposition}
% \newtheorem{remark}{Remark}
% \newtheorem{theorem}{Theorem}

% \usepackage[bf,small,nooneline,normal]{caption}
% \renewcommand*{\captionlabeldelim}{\nobreak}
% \renewcommand*\captionlabeldelim{.}
% \renewcommand{\captionfont}{\bfseries}

\renewcommand{\thetable}{\Roman{table}}

\usepackage{indentfirst}

\renewcommand{\abstractname}{\bf \small ABSTRACT}

\renewcommand{\refname}{\bf \small REFERENCES}

\usepackage{endnotes}
\renewcommand{\footnote}{\endnote}
\renewcommand{\notesname}{\bf Footnotes}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Financialization of commodities: an asset pricing perspective},
            pdfauthor={Devraj Basu, Olivier Bautheac},
            pdfkeywords={Financialization of commodities, futures contracts in agricultural
commodities, speculative bubbles, volume, open interest, convenience
yield.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{jf}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Financialization of commodities: an asset pricing perspective}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Devraj Basu, Olivier Bautheac}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\usepackage{caption}
\usepackage{setspace}
\doublespacing

\begin{document}
\maketitle
\begin{abstract}
The last two decades have seen major changes in the nature of the
commodity markets with the influx of financial investor inflows followed
by the financial crisis. We study the effect of these events on
commodity futures pricing dynamics using commodity asset pricing
techniques. We consider both futures market based factors and the
traditional market factor and find that a hedging pressure based factor
is able to best detect the effects of commodity financialization while
the effects of the financial crisis and after are most clearly visible
through the performance of the market factor. The futures market seems
to have been the main channel for the effects of commodity
financialization, which seems to have delivered an exogenous schock to
the cross-section of global commodities.
\end{abstract}

\newpage

\section{Introduction}\label{introduction}

The last twenty years have seen major upheavals in the commodity markets
with the onset of financialization\footnote{Starting around 2004, with a
  view of commodity markets as an emerging asset class and in an effort
  to hedge against inflation and seek diversification benefits
  \citep{buyuksahin_speculators_2014, singleton_investor_2013},
  institutional investors sent forth an unprecedented flow of capital
  into commodity futures markets. This phenomenon is commonly referred
  to as the financialization of commodities
  \citep{domanski_financial_2007}.} as well as the period of the 2008
financial crisis and its aftermath. Financialization appears to have
affected a number of economic sectors, including agriculture and energy,
and its impact has been hotly debated in the policy, legislative,
regulatory, and academic spheres. A vigorous legislative debate has led
to regulatory changes\footnote{Concerns over the consequences of
  financialization were behind Rule 76 FR 4752 issued by the U.S.
  Commodity Futures Trading Commission (CFTC) on January 26, 2011. This
  provision emanates from the Dodd-Frank Wall Street and Consumer
  Protection Act of 2010 (Title VII, Section 737) that mandates the CFTC
  to use position limits to restrict the flow of speculative capital
  into a number of commodity markets. The Rule was approved in a close
  3-2 vote and the ensuing rulemaking process was extremely contentious
  with several commissioners (Michael Dunn and Scott D. O'Malia in
  particular) expressing reservations about the lack of supporting
  evidence and the Rule also triggering thousands of comment letters as
  well as a lawsuit against the CFTC spearheaded by two Wall Street
  trade groups, namely the International Swaps \& Derivatives
  Association (ISDA) and the Securities Industries \& Financial Markets
  Association (SIFMA) (ISDA \& SIFMA v U.S. CFTC; Complaint,
  1:11-cv-02146; December 2, 2011).} accompanied by an equally intense
policy and academic debate across a number of disciplines about the
perceived effects on commodity prices of the unprecedented inflow of
institutional funds brought into commodity futures markets by new index
type investors\footnote{The first responses to the 2007/2008 crisis of
  escalating food and energy prices took the form of policy reports,
  many of which reasoned that the growth of commodity index funds came
  along with an influx of largely speculative capital that was
  responsible for driving commodity prices beyond their historic highs
  \citep{de_schutter_food_2010_1, gilbert_speculative_2010, herman_not_2011_1, schumann_hunger-makers_2011, unctad_global_2009}.}.
In response to immediate policy concerns, the academic debate was
initially framed rather narrowly\footnote{Early U.S. Senate
  investigations drew on pricing and trading data supplied by the CFTC
  as well as interviews with numerous experts including hedge fund
  manager Michael Masters who linked the growing presence of index
  investors in U.S. wheat markets to the observed surges in both futures
  and cash prices
  \citep{masters_testimony_2008, masters_accidental_2008}. This
  contention has since come to be known as the Masters' hypothesis and
  was largely endorsed by the final U.S. Senate report on the issue
  \citep{USSenate_Excessive_speculation_wheat_2009}. At this stage, the
  question thrust on the academic sphere was whether ``excessive
  speculation'' (understood as the peculiar market activity of index
  type investors) was linked to escalating energy and agricultural
  commodity prices.} and used a fairly restricted set of statistical
techniques\footnote{Most of the studies focused on statistical analyses
  of the Granger type causal connections between ``excessive
  speculation'' and commodity prices.}.

In this paper we take a fundamentally different empirical approach that
is broader in scope. We consider the entire cross-section of actively
traded commodities on futures markets and use a futures based asset
pricing framework that includes factors constructed using both price and
non-price attributes. Factor model techniques are well suited to
isolating common driving factors and detecting co-movement, a central
issue in financialization which we believe the existing literature has
not fully analyzed.\\
The first factor model we consider (CHP) is based on the well-known
hedging pressure based theory
\citep{anderson_hedger_1983, chang_returns_1985, cootner_returns_1960, dusak_futures_1973, hicks_value_1939, hirshleifer_risk_1988, hirshleifer_determinants_1989, hirshleifer_hedging_1990, kolb_is_1992, keynes_treatise_1930}
which postulates that futures prices for a given commodity are inversely
related to the extent that commercial hedgers are short or long and we
use a factor mimicking portfolio that captures the impact of hedging
pressure as a systemic factor \citep{basu_capturing_2013}. Our second
factor model is based on open interest growth following
\citet{hong_what_2012} who find that open interest growth has predictive
power for individual commodity returns and again we construct a
long-short factor mimicking portfolio. Our third model is based on term
structure following \citet{szymanowska_anatomy_2014} and
\citet{fuertes_commodity_2015} who demonstrate that the term structure
factor has explanatory power for the cross-section of commodity returns.
We finally consider an equally weighted portfolio of commodities as a
proxy for the market factor. Our two price based and two non-price based
factors allow us to compare and contrast different channels that could
have led to changes in commodity price dynamics over the last twenty
years.

We study the performance of these models on a set of twenty-four U.S.
traded commodities and six non-U.S. traded (LME traded) industrial
metals over the 1997/2017 period that we divide into four distinct
sub-periods. The first, the pre-financialization phase, spans the
1997/2003 period; the second, the financialization phase, spans the
2004/2007 period\footnote{Starting point based on earlier studies
  \citep{baker_financialization_2014, christoffersen_factor_2014}.}; the
third, 2008/2013, encompasses both the financial crisis and the loose
monetary policy regimes that followed; and the fourth spans the period
starting in 2014 to present.\\
Consistent with earlier studies, we observe that mean returns for the
U.S. traded commodities are sharply higher over the financialization
period compared to pre-financialization and this pattern extends to LME
traded metals.

The cross-sectional nature of financialization
\citep{basak_model_2016, cheng_financialization_2014} leads us to
extends the Keynesian indivudal hedging pressure paradigm by
incorporating aggregate CHP (\(\mathbf{CHP}\)), a market wide measure of
hedging pressure. This paradigm implies that returns for individual
commodities should be high in periods where \(\mathbf{CHP}\) is low
(\(\mathbf{backwardation}\)) and low in periods where \(\mathbf{CHP}\)
is high (\(\mathbf{contango}\)) vice et versa. Over the first period the
cross sectional average returns during \(\mathbf{backwardation}\) for
both the US and UK commodities are positive with those for the US
statistically significant, and significantly higher than during
\(\mathbf{contango}\), providing broad support for an aggregate version
of the Keynesian hedging pressure hypothesis. Over the second period
this pattern reverses, with both the US and UK commodities having
significant positive returns during \(\mathbf{contango}\) with these
returns being significantly higher than during
\(\mathbf{backwardation}\). This change in pattern suggests that the
onset of financialization altered the traditional risk shifting nature
of hedger's behaviour, which is implied by the Keynesian hypothesis.

We analyze the pricing dynamics over the different periods by first
studying the time series pricing performance of the various factors on
all the U.S. traded commodities with the factors constructed using the
latter set of commodities. The market based factor outperforms the other
three in the first period and it has a similar level of performance in
the second period, while the CHP factor shows the greatest improvement
in performance in the second period. The performance of the market
factor is driven by agricultural commodities and metals in both the
first and second period and the dramatic improvement of the CHP factor
in the second period is driven by precious and precious industrial
metals. The best performing commodities for the market factor exhibited
approximately the same level of price co-movement both before and during
financialization. In contrast, the best performing commodities for the
CHP factor exhibited very low pair-wise correlation in the first period,
around the average for U.S. traded commodities, while during
financialization the average pairwise correlation between the top
performing CHP factor commodities is higher than that for the market
factor and double of the average for all the U.S. traded commodities.
This increase is due to the dramatic rise in average pairwise
correlation amongst the U.S. traded metals all of which exhibited high
R\textsuperscript{2} with respect to the CHP factor during
financialization which in turn is due to their common low level of CHP
over this period. Thus the CHP factor model is able to establish the
strengthening link between a non-price attribute and price dynamics in
the U.S. metal sub-sector during financialization.\\
We explore this issue further by constructing the four factors with the
top eight commodities for a given factor (factor picks) in each
period\footnote{See section 2.2 for further details.} and evaluating its
performance on these selected commodities as well as on all U.S. traded
commodities. The logic of this exercise is to determine the extent to
which the pricing performance of these factors in each period is
concentrated in a subset of commodities. The CHP and the market factors
built from CHP factor picks exhibit the most dramatic increase in
pricing performance on the corresponding factor picks in the second
period. In contrast, the increase in pricing performance on all the U.S.
traded commodities is roughly uniform across all sets of factor models
and lower in magnitude. This strongly suggests that the hedging pressure
co-movement and price co-movement link was concentrated in the metals
sub-sector. There is also a dramatic improvement in pricing performance
of the CHP factor built with term structure picks indicating convergence
between Keynesian and Working's theories of the futures term structure
in a U.S. context.

The LME traded industrial metals provide a useful set of assets to test
how widespread were the changes in pricing dynamics brought about by the
onset of financialization. To that end, we examine the performance of
the four factors on the six LME metals, when constructed from each set
of factor picks. The performance of the market factor built with the CHP
and term structure factor picks improves dramatically in the second
period relative to the first while the performance of all factors built
with the other sets of picks does not increase quite as much. The
pricing performance of the market factor built with CHP factor picks as
well as that of the CHP factor built with both CHP and term structure
factor picks show a dramatic increase over the financialization period.
The performance of the market factor built with the CHP factor picks
suggests that there was co-movement in global metals returns in the
second period and that this co-movement could be detected by our hedging
pressure based model while the performance of the CHP factor built with
term structure factor picks indicates the aforementioned convergence
between Keynesian and Working's views of commodity futures term
structure extends to a global context.

Our analysis suggests that financialization appears to have had the
deepest impact on global metals markets and our commodity asset pricing
based approach allows us to better understand the channels through which
it was transmitted. The initial view of financialization was that it
consisted of speculative flows which had the effect to driving up and
creating bubble like conditions in commodity future prices\footnote{\citet{de_schutter_food_2010_1};
  \citet{gilbert_how_2010}; \citet{gilbert_speculative_2010};
  \citet{masters_testimony_2008}; \citet{masters_accidental_2008};
  \citet{herman_not_2011_1}; \citet{schumann_hunger-makers_2011};
  \citet{singleton_investor_2013}; \citet{unctad_global_2009}}. A number
of academic studies, most of which used commodity specific causality and
correlation based analysis, have disproved this contention, particularly
in the case of agricultural and energy commodities\footnote{\citet{irwin_devil_2009},
  \citet{sanders_adequacy_2010}, \citet{sanders_impact_2011},
  \citet{sanders_new_2011}, \citet{irwin_commodity_2013},
  \citet{brunetti_commodity_2014}, \citet{hamilton_effects_2015} and
  \citet{bruno_financialization_2017} study the agricultural markets.
  \citet{buyuksahin_speculators_2011}, \citet{tokic_speculation_2012},
  \citet{fattouh_role_2013}, \citet{kilian_role_2014},
  \citet{knittel_simple_2016} and \citet{manera_modelling_2016} study
  the energy markets. \citet{bohl_does_2013_2}, \citet{kim_does_2015}
  and \citet{boyd_prevalence_2016} study both energy and agricultural
  markets while \citet{irwin_index_2011},
  \citet{irwin_financialization_2012} and \citet{stoll_commodity_2011}
  study the commodity markets in general.}, while other studies have
underlined various alterations in commodity pricing dynamics over the
period\footnote{\citet{gilbert_role_2014} reject the view that
  financialization has not had any effects in the grains markets and
  demonstrate that trades originated by financial actors, and
  specifically index investors, can move prices but tend typically to be
  volatility-reducing. \citet{cheng_financialization_2014} show that in
  the case of index commodities, financialization has transformed the
  risk sharing, and information discovery functions of commodity futures
  markets. \citet{juvenal_speculation_2015} contend that the oil price
  increase between 2004 and 2008 was mainly driven by the strength of
  global demand but that the financialization process of commodity
  markets also played a role. \citet{henderson_new_2015} show that
  non-information-based financial investments have important impacts on
  commodity prices.}. However, the fundamental question of the nature of
the impact of financialization on commodity futures markets has not yet
been completely answered. Our analysis takes a broader perspective and
shifts the focus away from the energy and agricultural sectors. We
provide an empirical complement to a new stream of theoretical studies
that try to model the impact of financialization on various aspects of
commodity futures markets\footnote{\citet{etula_broker-dealer_2013},
  \citet{acharya_limits_2013}, \citet{cheng_convective_2014},
  \citet{leclercq_equilibrium_2014}, \citet{sockin_informational_2015},
  \citet{goldstein_speculation_2013},
  \citet{EkelandSpeculationcommodityfutures2016},
  \citet{goldstein_commodity_2017}.}, as well as extend the empirical
findings of \citet{basak_model_2016} and \citet{tang_index_2012}. In
contrast to the earlier purely statistical studies that were designed to
study speculative behavior alone, our empirical approach allows us to
isolate the more subtle effects of financialization identified in the
theoretical studies above.

The onset of the financial crisis and the monetary policy regimes that
followed appear to have also induced significant changes in pricing
dynamics. Returns on both U.S. and LME traded commodities fell
dramatically over both the third and fourth periods with the pattern of
return for the U.S. traded commodities across \(\mathbf{CHP}\) regimes
reverting to Keynesian paradigm. The pricing results indicate the
presence of a systematic factor across the entire cross-section of the
U.S. commodities as the performance of the market factor built with
every set of picks on the entire set of U.S. traded commodities shows a
dramatic rise across the board in pricing performance. There is also
strong evidence of cross-market linkages as the market factor built with
its own picks (U.S. traded commodities) shows a six fold increase in
average R\textsuperscript{2} on the LME metals compared with the
financialization period. The market factor when built with equity index
(S\&P 500) picks does even better, providing strong evidence of
systematic spill-overs from the equity to the commodity markets.

Financialization was an issue of such policy importance that it
triggered legislative action. The debate was initially framed around
adequacy of speculation, the burning issue of the day, and the
consequent analysis focused on the more mechanical effects of
financialization. With the benefit of hindsight, it appears this
approach was perhaps too narrow and it now seems necessary to address
financialization from a broader perspective. Commodity price dynamics
appear to have altered substantially in quite different ways over the
financialization and the financial crisis periods and our commodity
futures based asset pricing approach seems to be able to provide new
insights into the nature of these changes: financialization was a
phenomenon endogenous to the commodity markets transmitted via the
commodity futures markets and mainly concentrated in a specific sector
(metals) while the crisis and its aftermath seems to have delivered an
exogenous shock across the entire cross-section of global commodities.

\section{Data \& methods}\label{data-methods}

\subsection{Data}\label{data}

Commodity index funds (CIFs), commodity based exchange traded funds
(ETFs) and notes (ETNs) as well as commodity linked notes (CLNs) were
amongst the main investment vehicules that allowed institutional and
retail investors to build up commodity exposure in the early 2000s
\citep{boons_stock_2012, henderson_new_2015, irwin_index_2011, unctad_global_2009, schumann_hunger-makers_2011}.
Most of these products were primarily designed to track prominent
commodity indexes amongst which the S\&P-GSCI, the most popular. Its
constituents at the time of writting, all U.S. traded, form the basis of
the commodity pool we consider in this study where we also include the
cross-section of the most liquid LME traded metals\footnote{See appendix
  1 for details on the commodity pool considered.}. Unleaded gasoline,
the NYMEX legacy contract for gasoline stopped trading in 2006 when it
was replaced by Reformulated Gasoline Blendstock for Oxygen Blending
(RBOB) with the two futures series trading alongside for most of the
year. For our NYMEX gasoline data series, we consider unleaded gasoline
up to Septembre 1\textsuperscript{st}, 2006 and RBOB gasoline
thereafter, date at which liquidity for the latter overtook that for the
former.

The commodity futures trading commission (CFTC) operates a comprehensive
system of collecting information on market participants know as the
large trader reporting program (LTRP) where it records and reports to
the public market position data on market particpants who have position
levels in excess of a particular futures market specific
threshold\footnote{Current reporting level thresholds available at:
  \href{https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=\&SID=970471b8455f4bab7db4110cfde50731\&mc=true\&r=SECTION\&n=se17.1.15_103}{www.ecfr.gov}.}.
The commission collects market data and position information daily from
clearing members, futures commission merchants (FCMs) and foreign
brokers and publishes corresponding summaries weekly in a series of
market specific reports. In these reports, individual traders are
categorised according to the nature of their trading activity on the
basis of self-reported information\footnote{CFTC form 40. Available at:
  \href{https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf}{www.cftc.gov}.}
that is subject to review by CFTC staff for reasonableness. Position
information is then aggregated by category and asset type with each
report providing an aggregatetion for a set of report specific
categories and for each category, a breakdown between futures only
positions and positions in futures and options combined.\\
In its legacy report, the commitment of trader report (COT) with data
dating back to 1962, the CFTC distinguishes ``commercial'' and
``non-commercial'' market participants where a ``commercial''
participant is defined as one ``{[}\ldots{}{]} engaged in business
activities hedged by the use of futures and option markets''\footnote{See
  \href{https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml}{CFTC
  regulation 1.3, 17 CFR 1.3(z)} for details.}. A third category,
``non-reportable'', aggregates positions for participants not meeting
the reporting threshold. In response to concerns related to category
accuracy\footnote{See for example \citet{ederington_who_2002}.} the CFTC
now refines its classification in the disaggregated commitment of trader
report (DCOT) where participant categories include
``Producer/Merhcant/Processor/User'', ``Swap dealer'', ``Money manager''
and ``Other reportable'' as well as in the supplemental commitment of
trader report (SCOT) that details commodity index trader aggregate
positions for thirteen agricultural commodity futures markets.\\
In this study we consider futures only data over the 1997/2017 period.
Because DCOT and SCOT data series only date back to 2006 and 2007
respectively, we rely exclusively on the legacy COT report for market
position data. Futures market price and open interest data are from
Bloomberg.

\subsection{Methods}\label{methods}

On the one hand, futures markets have a long history of serving
commodity producers to allievate their commodity price and output risks.
\citet{keynes_treatise_1930} and \citet{hicks_value_1939} emphasise the
``normal'' behaviour of naturally short hedgers outsourcing their
commodity risk to naturally long commodity specialist speculators. This
market configuration is referred to as ``backwardation'' where futures
is expected to trade at a discount to spot, providing speculators with
incentives to take the long side. The opposite market paradigm is
referred to as ``contango''. Working
\citep{working_theory_1948, working_hedging_1953} and Cootner
\citep{cootner_returns_1960, CootnerSpeculationhedging1967} somehow
relax this ``supply-of-speculative-services''
\citep{Tilllongtermperspectivecommodity2007} approach by introducing
processors and merchants, also commodity experts, on the hedgers side
and relating the notions of backwardation and contango more directly to
the term structure via the theory of storage where the shape of the
front end of the term structure (basis) is directly related to the
market price of storage.\\
Financialization, on the other hand, refers to the entry of financial
investors into the commodity futures markets who, for the most part, are
not commodity experts. Besides, these new market participants tend to
exhibit herding like behaviour, often taking massive long-only positions
on the whole cross-section of indexed commodities in an attempt hedge
against inflation and/or seek further diversification benefits for
portfolios with, in most cases, existing large positions across various
asset classes
\citep{brunetti_speculators_2016, boyd_prevalence_2016, cheng_convective_2014, juvenal_speculation_2015, singleton_investor_2013, tang_index_2012}.
This contrasts with the traditional approaches of Keynes and Working and
in this context the issues of co-movement and aggregate market
participants behaviour seem particularly relevant.\\
Asset pricing techniques are well suited to study co-movement. Building
on theory we thus develop an ad-hoc commodity based asset pricing
framework that we believe is well adapted to study these two issues
simultaneously and implement it over four periods of interest. The
first, the pre-financialization period, starts in 1997, the earliest
year for which LME metals' price data are available on Bloomberg, and is
naturally followed by financialization, the second, that spans the
2004/2007 period with the starting year based on earlier studies
\citep{baker_financialization_2014, christoffersen_factor_2014}. The
third spans the 2008/2013 period, encompassing both the financial crisis
and the loose monetary policy that followed, while the fourth spans the
period starting for the end of this unorthodox monetary policy regime in
2014 to present.\\
Using the weekly price, open interest and market position data described
above we construct market portfolios of commodity nearby futures as well
as mimicking portfolios for commercial hedging pressure, term structure
and liquidity risk factors in returns. While our market portfolios are
long only, equally weighted portfolios of particular commodity sets, our
other mimicking portfolios for risk factors are constructed as
combinations of two sub-portfolios (legs), one held long, one held
short, which respective constituents are pooled on a weekly basis
according to risk factor specific criterias. The weekly return for a
particular mimicking portfolio is calculated as the difference between
the average return for constituents of the long portfolio and that for
the constituents of the short portfolio. i.e, return on the mimicking
portfolio for risk factor \(i\) (\(f_{i}\)) for week \(t\):

\[r_{t}^{f_{i}}=\frac{1}{x}\sum_{j=1}^{x}r_{j,t}-\frac{1}{x}\sum_{j=n-x}^{n}r_{j,t}\]
\(n\equiv\) number of commodities in the set considered for mimicking
portfolio construction.\\
\(x = \left \lceil n \cdot s \right \rceil\)\\
\(s\equiv\) selection threshold (\(\frac{1}{3}\) here).\\
\(r_{j,t}=\frac{C_{j,t}}{C_{j,t-1}}-1\)\\
\(C_{j,t}\equiv\) week \(t\) close price for commodity \(j\).\\
\(j\equiv\) commodity rank in the ordered set \(Y_{t}^{f_{i}}\).\\
\(Y_{t}^{f_{i}}\equiv\) week \(t\) ordered set of the commodities
considered for mimicking portfolio construction where the ordering rule
is specific to \(f_{i}\).

For our portfolio CHP (commercial hedging pressure), we sort the set of
commodities considered on past twenty six-week average CHP from lowest
to highest. The bottom third constituents (lowest \(\overline{CHP}\)s)
of the corresponding ordered set form the long leg of the portfolio
while the top third form the short leg:

\[Y_{t}^{CHP}\equiv\left \{ \left \{ \overline{CHP_{1}}, \overline{CHP_{2}}, ..., \overline{CHP_{n}} \right \}, \leq \right \}\]

\(\overline{CHP_{j}}=\frac{1}{26}\sum_{k=0}^{25}\frac{L_{j,t-k}}{L_{j,t-k}+S_{j,t-k}}\)\\
\(L_{j,t}\equiv\) number of long positions held by commercial hedgers
for week \(t\) on the futures series of commodity \(j\).\\
\(S_{j,t}\equiv\) number of short positions held by commercial hedgers
for week \(t\) on the futures series of commodity \(j\).

For our portfolio TS (term structure), we sort the set of commodities
considered on past twenty six-week average roll-yield from highest to
lowest. The bottom third constituents (highest \(\overline{RY}\)s) of
the corresponding ordered set form the long leg of the portfolio while
the top third form tje short leg:

\[Y_{t}^{TS}\equiv\left \{ \left \{ \overline{RY_{1}}, \overline{RY_{2}}, ..., \overline{RY_{n}} \right \}, \geq \right \}\]

\(\overline{RY_{j}}=\frac{1}{26}\sum_{k=0}^{25}(\frac{F_{j,t-k}^{1}}{F_{j,t-k}^{2}} - 1)\)\\
\(F_{j,t}^{1}\equiv\) week \(t\) close price for commodity \(j\) first
nearby futures contract.\\
\(F_{j,t}^{2}\equiv\) week \(t\) close price for commodity \(j\) second
nearby futures contract.

For our portfolio OI (open interest), we sort the set of commodities
considered on past twenty six-week average aggregate open interest
growth from highest to lowest. The bottom third constituents (highest
\(\overline{\mathbf{OI}^{\Delta \%}}\)s) of the corresponding ordered
set form the long leg of the portfolio while the top third form the
short leg.

\[Y_{t}^{OI}\equiv\left \{ \left \{ \overline{\mathbf{OI}_{1}^{\Delta \%}}, \overline{\mathbf{OI}_{2}^{\Delta \%}}, ..., \overline{\mathbf{OI}_{n}^{\Delta \%}} \right \}, \geq \right \}\]

\(\overline{\mathbf{OI}_{j}^{\Delta \%}}=\frac{1}{26}\sum_{k=0}^{25}(\frac{\mathbf{OI}_{j,t-k}}{\mathbf{OI}_{j,t-k-1}} - 1)\)\\
\(\mathbf{OI}_{j,t}=\sum_{k=1}^{l_{j,t}}OI_{j,t,k}\)\\
\(OI_{j,t,k}\equiv\) open interest at close of week \(t\) for the
\(k^{th}\) nearby contract on the futures term structure for commodity
\(j\).\\
\(l_{j,t}\equiv\) length of commodity \(j\)'s futures term structure on
week \(t\) close.

We use a time series regression approach where we start with regressions
of the U.S. traded individual commodity weekly front returns on returns
to the mimicking portfolio for a particular risk factor where the
portfolio is constructed using the whole cross-section of U.S. traded
commodities. The corresponding adjusted \(R^{2}\)s are averaged and the
process is repeated for each of our four risk factors (market, hedging
pressure, term structure and liquidity) and periods (1997/2003,
2004/2007, 2008/2013 and 2014/2017). i.e, average adjusted \(R^{2}\) for
risk factor \(i\) (\(f_{i}\)) over period \(t\):

\[\overline{\tilde{R}_{f_{i},t}^{2}}=\frac{1}{24}\sum_{j=1}^{24}\tilde{R}_{j,t}^{2}\]
\(\tilde{R_{j}^{2}}=1-(1-R_{j}^{2})\frac{T-1}{T-p-1}\)\footnote{The
  Wherry formula-1.}\\
\(T\equiv\) length of period \(t\) (number of weeks here).\\
\(p\equiv\) number of regressors (1 here).\\
\(R_{j,t}^{2}=1-\frac{\sum_{k=1}^{T}(y_{j,k}-\hat{y_{j,k}})^{2}}{\sum_{k=1}^{T}(y_{j,k}-\overline{y_{j}})^{2}}\)\\
\(y_{j,t}\equiv\) week \(t\) nearby return for commodity \(j\).\\
\(\hat{y_{j,t}}=\beta_{0}+\beta_{1}f_{i,t}+\epsilon\)\\
\(\beta s \equiv\) classic OLS coefficient estimates.\\
\(f_{i,t}\equiv\) week \(t\) return on mimicking portfolio for risk
factor \(f_{i}\)

The same operation is repeated for the two (\(p=2\)) and three (\(p=3\))
factor model cases, where regressors include in turn all combinations of
any two and three risk factors respectivly, as well as for the four
(\(p=4\)) factor model case where all the risk factor mimicking
portfolios are included. Results are reported for the one-factor model
case with the other results available from the authors upon request.\\
For each period we also implement the analysis described above over
periods of low and high aggregate CHP (\(\mathbf{CHP}\)), a notion that
we introduce in this study and that we believe is particularly relevant
in the context of financialization where issues of co-movement and
aggregate market participants' behaviour play a central role. We define
\(\mathbf{CHP}\) as the cross-sectional average of CHPs for a particular
pool of asset. i.e, Week \(t\) \(\mathbf{CHP}\) for asset pool \(x\):

\[\mathbf{CHP}_{t}^{x}=\frac{1}{n}\sum_{i=1}^{n}CHP_{i,t}\] \(n\equiv\)
number of commodities in set \(x\).\\
\(CHP_{i,t}\equiv\) week \(t\) CHP for commodity \(i\) as defined above.

We further define aggregate backwardation (\(\mathbf{backwardation}\))
and contango (\(\mathbf{contango}\)) \(\mathbf{CHP}\) regimes as, for a
given period, \(\mathbf{CHP}\) levels below and above the period's
median respectively.\\
For the rest of the study, the asset pool used for \(\mathbf{CHP}\)
construction include the whole cross-section of U.S. traded commodities.

In the next part of the analysis, our mimicking portfolios for risk
factors, constructed from the set of all U.S. traded commodities are
used as commodity picking devices on the latter commodity set. For each
risk factor and period, individual U.S. traded commodity front returns
are regressed on returns for the corresponding mimicking portfolio.
Individual commodities are sorted by adjusted \(R^{2}\)
(\(\tilde{R^{2}}\)) accordingly from highest to lowest and the bottom
third (highest \(\tilde{R^{2}}\)s) are selected as commodity picks for
the risk factor for the period. i.e, set of picks \(P_{t}^{f_{i}}\) for
risk factor \(i\) (\(f_{i}\)) over period \(t\):

\[P_{t}^{f_{i}}\equiv\left \{ P_{1}, P_{2}, ..., P_{x} \right \}\]
\(x = \left \lceil n \cdot s \right \rceil\)\\
\(n\equiv\) number of commodities in the set considered for the picking
excersise (24 here).\\
\(s\equiv\) selection threshold (\(\frac{1}{3}\) here).\\
\(P_{j}\equiv\) commodity pick \(j\) for risk factor \(f_{i}\) over
period \(t\)\\
\(j\equiv\) rank in the ordered set \(Y_{t}^{f_{i}}\)\\
\[Y_{t}^{f_{i}}\equiv\left \{ \left \{ \tilde{R_{1}^{2}}, \tilde{R_{2}^{2}}, ..., \tilde{R_{n}^{2}} \right \}, \geq \right \}\]
\(n\equiv\) number of commodities in the set considered for picking (24
here).\\
\(\tilde{R_{j}^{2}}\equiv\) risk factor \(f_{i}\) returns' explanatory
power on commodity \(j\) nearby returns over period \(t\) as defined
above.

For each period and \(\mathbf{CHP}\) regime we record the proportion of
time each pick is held in the long and (/or) short leg of the
corresponding risk factor mimicking portfolio as well as the average
pairwise correlation amongst every sets of picks.

In the last part of the analysis, for each period our mimicking
portfolios for risk factors in returns are constructed using each set of
commodity picks independently as the asset pool for portfolio
construction. A similar time series approach to that described above is
then implemented using these newly formed mimicking portfolios. In this
case, a first set of dependent variables includes the sets of commodity
picks themselves, a second set includes the whole cross-section of U.S.
traded commodities while a third set includes the six LME traded metals
considered. The latter are also used as dependent variables in the last
series of regressions where the long and short legs of the newly formed
risk factor mimicking portfolios described above are considered
separately as regressor.

\section{Results \& discussion}\label{results-discussion}

The descriptive statistics over the various time periods for our two
sets of commodities (U.S. and non U.S.) as well as the four factors
considered are shown in Table 1. The traditional Keynesian hedging
pressure paradigm postulates a negative relationship between the return
on an individual commodity and its hedging pressure. Financialization
has been seen as an exogenous cross-sectional commodity market
phenomenon. This naturally leads us to extend the individual hedging
pressure paradigm by incorporating aggregate CHP (\(\mathbf{CHP}\)), a
market wide measure of hedging pressure. The extended Keynesian hedging
pressure paradigm should also imply that returns for individual
commodities should be high in periods where \(\mathbf{CHP}\) is low and
vice et versa.\\
In table 2 the mean returns are shown over high (\(\mathbf{contango}\))
and low (\(\mathbf{backwardation}\)) \(\mathbf{CHP}\) regimes for each
time period. Over the first period 18 of the 24 US commodities have
higher mean returns during periods of \(\mathbf{backwardation}\) as
compared to \(\mathbf{contango}\), with the difference being
statistically significant at the 5\% level for one commodity and for the
remaining 6 commodities the higher mean return during
\(\mathbf{contango}\) is not statistically significant for any of them.
For the 6 UK commodities, all metals, 4 of them had higher mean returns
during phases of \(\mathbf{backwardation}\) as compared to
\(\mathbf{contango}\) in the first period, with one of these differences
being statistically significant at the 5\% level while for the remaining
two the higher mean return during \(\mathbf{contango}\) is not
statistically significant for either. An equally weighted portfolio of
the 24 US commodities had a mean return of 10.98\% during
\(\mathbf{backwardation}\) phases, statistically significant at the 1\%
level, and a mean return of 0.93\% during \(\mathbf{contango}\), both in
the first period, with the difference between them statistically
significant at the 5\% level. The corresponding figures for an equally
weighted portfolio of the UK metals were 8.35\% and -1.71\%, the
difference again statistically significant at the 5\% level. These
results provide broad support for an aggregate version of the Keynesian
hedging pressure hypothesis and suggests that hedgers in the aggregate
were engaged in risk transfer.\\
Over the second period the pattern essentially reverses. For the 24 US
commodities over this period, 18 have higher mean returns during phases
of \(\mathbf{backwardation}\) over \(\mathbf{contango}\), with the
difference statistically significant for 4 commodities and of the
remaining 7 the higher mean return during \(\mathbf{backwardation}\) is
statistically significant for one commodity. All of the 6 UK metals
achieved higher mean returns over \(\mathbf{backwardation}\) than
contango with the difference being statistically significant for one
commodity. The equally weighted US portfolio had a mean return of
11.16\% over \(\mathbf{backwardation}\) while it was 26.68\% during
\(\mathbf{contango}\), the first statistically significant at the 5\%
level and the second at the 1\% level, with the difference statistically
significant at the 1\% level. The difference was considerably more
pronounced for the equally weighted UK metals portfolio with a mean
return of 6.43\% during \(\mathbf{backwardation}\) rising to 41.22\%
during \(\mathbf{contango}\), with the difference also significant at
the 1\% level.\\
Thus the onset of financialization seems to have engendered a change in
the nature of hedger's behaviour in the aggregate, taking it away from
risk shifting, the traditional Keynesian view. This phenomenon has been
noted in earlier theoretical and empirical work
\citep[\citet{stout_why_1998}]{DanthineInformationfuturesprices1978} in
different context from which two competing models of behaviour have
emerged. The information arbitrage model implies that hedgers may often
be seeking out counterparties to trade with rather than being purely
passive. This issue is also raised in both
\citet{cheng_financialization_2014} and
\citet{StulzRethinkingriskmanagement1996} who point out that hedgers may
be taking a view on prices just as speculators do. In fact, by hedging
away some of their risk, hedgers are able to speculate more heavily
based on their disagreement against speculators regarding futures price
movement \citep{SimsekSpeculationrisksharing2013}, which fits into a
heterogenous expectation theory of speculation. As
\citet{stout_why_1998} points out, this disagreement-based theory of
trading is one of the main reason the public and the law disapproves the
speculators as this form of trading is regarded as
non-productive\footnote{\citet{DuffieChallengespolicytreatment2014} also
  discusses some of the challenges faced by a policy treatment of
  speculative trading motivated by differences in beliefs.}. In the
second model hedgers are seen as liquidity providers as in
\citet{Kangtaletwopremiums2017}. This explanation is also consistent
with the nature of financialization which led to the arrival of long
term, long only investors who require substantial liquidity to roll-over
their positions.

The average time series pricing performance of the various one factor
models on the U.S. traded assets over the four periods is shown in Table
3. Over the first period the market factor has the best average time
series performance with the highest average adjusted
R\textsuperscript{2} with all three futures market factors performing
poorly and the overall pricing performance consistent across periods of
high and low \(\mathbf{CHP}\) (below period median). The two factor
models average R\textsuperscript{2}s (results available upon request)
are roughly the sum of the one factor R\textsuperscript{2}s indicating
that the factors are almost orthogonal over this period. The onset of
financialization leads to a substantial improvement in the performance
of the CHP factor with its average R\textsuperscript{2} increasing to
5.4\% in the second period relative to 1.4\% in the first. The
performance of the market factor also improves though not by as much in
a relative sense (18.6\% from 12.2\%). Both the term structure factor
and the open interest growth factor show a much smaller improvement in
pricing performance. Over all periods, all of our factor models show
similar pricing performance across regimes of high and low
\(\mathbf{CHP}\).

We next investigate the top eight of the U.S. traded commodities which
achieved the highest time series R\textsuperscript{2} for the CHP,
market, open interest and term structure factors, in each of the four
time periods with the commodity picks shown in Table 4. For the CHP
factor the commodities that achieve high R\textsuperscript{2} in a given
period are likely to be those with relatively high or low CHP over that
period, or those commodities that co-vary with the short or the long
side of the factor. The same applies to the open interest and term
structure factors with CHP replaced by open interest growth and roll
yield respectively; while for the market factor the picks are those that
co-vary most strongly with the market factor over that period. For the
CHP factor, we observe a change in the picks between the first and the
second period; the picks in the first period included five agricultural
commodities and three metals while in the second period include all of
the five industrial and precious metals and only three agricultural
commodities. This pattern is also visible for the term structure factor
for which there are four metals (gold, copper, palladium and platinum)
picks in the second period against three in the first (gold, palladium
and platinum). The market factor has two metal picks in the second
period (palladium and silver) while all of the picks in the first period
were agricultural or energy commodities. Over the third period, the
average CHP factor has four metal picks while both the term structure
factor and market factor have three. Table 4 also shows the proportion
of time that these picks were constituents of the corresponding factor.
All of the picks of the long-short factors (average CHP, open interest
growth and term structure factors) were constituents of the
corresponding factor for some proportion of time in all of the time
periods. In this respect, there is a contrast between the first and
second periods for the average CHP with the picks appearing
predominantly on the long side in the second period, as opposed to both
legs in the first. This indicates that the picks in the second period
consistently exhibited low CHP. This behaviour continued over the third
period.

In table 5, we show the average pair wise correlation between each set
of picks in each time period. This exercise gives an indication of the
extent to which attribute co-movement reveals price co-movement, which
is particularly relevant for the long-short factors for which the
corresponding attribute is not directly related to price co-movement. In
the first period, the average CHP picks pairwise correlation is similar
to the average across the U.S. traded 24 commodities. In the second
period however it is the highest, comparable with the market and double
of the U.S. traded commodities average. This increase is a consequence
of the dramatic increase in average pairwise price correlation amongst
the U.S. traded metals, all of which are CHP factor picks. Taken
together with the fact that four of these metals had consistently low
CHP, we see that CHP co-movement in the second period is strongly
correlated with price co-movement. It is also interesting to note that
U.S. traded metals returns were negatively correlated with the rest of
U.S. traded commodities in the first period and positively in the second
(Results available upon request).

In order further understand the link between attribute and price
co-movement we build all of the factors from every set of factor picks
and assess their pricing performance on these sets of picks. The results
are shown in table 6. The long-short factors for a particular attribute
select the top and bottom three of the corresponding picks based on that
attribute while the market factor takes an equally weighted combination
of the eight corresponding picks. In the absence of price co-movement
between the picks, this market factor will achieve an average
R\textsuperscript{2} of around 12.5\%\footnote{Market factor lemma:
  \(\overline{R_{r_{mkt}, p}^{2}} = \frac{1}{n} \sum_{j} R_{r_{j}, r_{mkt}, p}^{2} = \frac{1}{n}; \: n\equiv\)
  size of market portfolio (8 here). See appendix 2 for a formal proof.},
while there is no non-zero R\textsuperscript{2} lower bound for the
corresponding long-short factors. A high average R\textsuperscript{2}
for the market factor is indicative of strong price co-movement while
for the long-short factors it is indicative of a simultaneous attribute
and price co-movement. The most dramatic change in pricing performance
occurs for the average CHP factor picks where the market factor built
from these picks has an average R\textsuperscript{2} of 21.3 \% in the
first period (close to the 0 correlation situation) and jumps to 37.4 \%
in the second period. The average CHP factor built with average CHP
factor picks also shows a sharp increase in pricing performance over the
second period. The most striking result is the dramatic improvement in
pricing performance of the CHP factor built with term structure picks
(3.4\% vs.~21.2\%)\footnote{In contrast to the market factor, there is
  no pricing performance lower bound for the CHP factor built with a set
  of picks.}, providing evidence of convergence of Keynesian and
Working's theories of the term structure in detecting price co-movement
during financialization. In table 7 we assess the performance of the
same factors on all the commodities instead of just the picks. While the
average R\textsuperscript{2}s are expectedly lower, the relative
performance of the market factors built with different sets of picks is
similar to table 6. In contrast, the increase in R\textsuperscript{2}
for the average CHP factor built with average CHP factor picks is lower
indicating that the link between CHP co-movement and price co-movement
is strongest amongst the picks which are predominantly metals.\\
The conclusion from Tables 6 and 7 is that the onset of financialization
seems to have had a major impact on the dynamics of U.S. traded precious
and precious industrial metals and we now seek to analyse whether this
phenomenon is true for the non US metal futures. To that end we examine
the pricing performance of the factors constructed used in Table 6 on
the six LME metals futures. The results are reported in Table 8. In the
first period, the average R\textsuperscript{2} s across the board are
always less than 4.5\%. In the second period we see a dramatic increase
in R\textsuperscript{2} for the market factor built with average CHP
factor picks and average term structure factor picks (4.48\% to 25.1\%
and 1.3\% to 20.6\% respectively) showing dramatically higher
co-movement between LME traded metals and the corresponding picks all of
which are U.S. traded commodities. The increase in R\textsuperscript{2}
for the market factor built with market factor picks is much more
modest, providing further evidence of concentration of co-movement in
the metal sector during financialization.

In order to understand this better, in table 9, we examine the pricing
performance of both long and short legs of the corresponding factors
separately on the same set of LME traded metals. The most striking
result here is the dramatic increase in R\textsuperscript{2} for the
long leg of the term structure factor built with CHP factor picks which
achieves the highest overall R\textsuperscript{2} in the second period
(1.91\% vs.~27.37\%). This provides further evidence that global metals
co-movement could be detected by US based factor models that combined
the Keynesian and Working's notions of backwardation.

We then turn our analysis to the crisis and post-crisis period
(2008-2013). In table 1 we see that the returns on both U.S. and
non-U.S. commodities fall dramatically over this period, going down to
-4.3\% for the LME metals and 8\% for the US traded commodities although
volatility for both sets of assets is similar to that during the second
period. Two of the LME metals have negative mean returns over the third
period compared to one over the first and none over the second. This
trend continues in from 2014 to 2017 with the US traded commodities
recording an average mean return of -0.6\% and the LME metals an average
of 4.9\% over this period, with fourteen US traded commodities and one
LME metals having negative mean returns. The volatility of both sets of
assets is lower than over the 2004-2007 period, particularly for the UK
metals, suggesting the impact of the investment outflows from the
commodity futures markets has had a much lower impact on volatility than
the corresponding inflows witnessed over the financialization period.
The market factor dominates pricing performance over this period with an
increase of mean average R\textsuperscript{2} of over 50\% while that
for the other factors increases slightly at best (table 3). From table 5
we see that there is an across the board increase in pairwise
correlations with a doubling in the average pairwise correlation of all
U.S. traded commodities and a levelling in the pairwise correlation of
the picks of the three long-short factors. Taken together, these results
indicate the presence of a systematic factor across the entire
cross-section of the U.S. commodities. This is clearly evident in the
performance of the market factor built with any set of picks on the
entire set of U.S. traded commodities (table 7) as we see a dramatic
rise across the board in pricing performance. Table 8 provides further
evidence of systematic linkages, in this case between U.S. and
non-U.S.traded commodities as the market factor built with its own picks
(U.S. traded commodities) achieves an average R\textsuperscript{2} of
33.7\% on the LME metals compared to 9.9\% in the previous period.

\section{Conclusion}\label{conclusion}

In the early 2000s, against a backdrop of a low yield environment and
poor stock market performance, a combination of financial
innovation\footnote{Tradable commodity price indexes (commonly referred
  to as commodity index funds or CIFs), and commodity-based exchange
  traded funds (ETFs) were the main financial innovations which allowed
  the large global banks to offer commodity investment products to
  institutional and retail investors. In 1991 Goldman Sachs created the
  S\&P-GSCI which provides investors with buy-side exposure to
  commodities via the OTC swap market and thus without having to
  participate in the formal futures markets with their position limit
  restrictions. At this stage these restrictions still applied to the
  bank though, as it hedged its commodity swap exposure in the futures
  markets. The first commodity-based ETFs were created through buying
  physical precious metals with gold and silver ETFs offered as early as
  2002/2003. The regulatory hurdle here related to the licensing of
  commodity trading professionals. Typically, investors had to sign a
  statement with their broker stating that they understood the risks of
  commodity investments; a rather inconvenient paperwork for a product
  designed to trade like a stock, as set forth by a number of industry
  players at the time.} and deregulation\footnote{In 2000, the Commodity
  Futures Modernization Act (CFMA) granted non-agricultural commodity
  futures statutory exemption from regulation (``Enron loophole'') while
  it required that agricultural commodity derivatives be traded on a
  CFTC-regulated exchange. Notwithstanding, the CFTC classified swap
  dealers as ``bona fide'' hedgers, granting them position limits
  exemption (``swap dealer loophole'').\\
  In 2005, the CFTC waived the rule that required commodity investors to
  sign a statement saying they understood the risks, letting the funds
  replace it with their prospectus and thus unleashed commodity-based
  ETFs investments.} led to a large inflow of institutional and
financial capital into the commodity futures markets. This process known
as financialization lead to a heated public policy debate about whether
the increase in open interest and trading volume in commodities exerted
upward pressure on prices. Perhaps in response to this, most of the
early studies focused on the more mechanical effects of
financialization, relying on commodity specific causality and
correlation based analysis. With the benefit of hindsight this focus
appears to have been too narrow and it seems important to analyze the
effects of financialization from a broader perspective which we do in
this study. We consider the entire cross-section of actively traded
commodities on futures markets and use a futures based asset pricing
framework that includes factors constructed using both price and
non-price attributes. Factor model techniques are well suited to
isolating common driving factors and detecting co-movement, a central
issue in financialization which we believe the existing literature has
not fully analyzed. Our analysis suggests that financialization was a
phenomenon endogenous to the commodity markets transmitted via the
futures markets and appears to have had the deepest impact on global
metal futures markets. The onset of the financial crisis and the
monetary policy regimes that followed appear to have also induced
significant changes in commodity pricing dynamics and our analysis
suggest that in contrast with financialization, the crisis and its
aftermath have delivered an exogenous shock across the entire
cross-section of global commodities.

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:mean_returns_no_CHP}This table shows descriptive statistics for nearby returns on various commodity groups as well as for mimicking portfolios for risk factors where the factor portfolios are constructed using the whole cross-section of the U.S. traded commodities considered in the study. Figures labelled $***$ ($**$, $*$) are significant at the 1\% (5\%, 10\%) level. See section 2 for more details.}\\
\toprule
Asset pool & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
US-Agriculturals-Grains & Mean return & 3.75\% & ***18.99\% & 5.34\% & -4.14\%\\
US-Agriculturals-Grains & Volatility & 27.14\% & 29.17\% & 33.52\% & 25.18\%\\
US-Agriculturals-Livestock & Mean return & 3.81\% & 5.84\% & *9.53\% & 0.16\%\\
US-Agriculturals-Livestock & Volatility & 26.53\% & 21.59\% & 23.94\% & 26.55\%\\
US-Agriculturals-Softs & Mean return & 1.57\% & **14.26\% & 9.53\% & 2.97\%\\
\addlinespace
US-Agriculturals-Softs & Volatility & 33.47\% & 30.19\% & 35.53\% & 28.08\%\\
US-Metals-Industrial & Mean return & 2.42\% & **33.08\% & 6.74\% & 1.35\%\\
US-Metals-Industrial & Volatility & 20.2\% & 30.44\% & 31.66\% & 19.95\%\\
US-Metals-Precious & Mean return & *9.39\% & ***22.7\% & 9.99\% & 2.13\%\\
US-Metals-Precious & Volatility & 26.2\% & 25.61\% & 31.76\% & 21.57\%\\
\addlinespace
US-Agriculturals & Mean return & 2.89\% & ***14.45\% & **7.86\% & -0.44\%\\
US-Agriculturals & Volatility & 29.72\% & 28.24\% & 32.69\% & 26.64\%\\
US-Energy & Mean return & 13.56\% & ***28.33\% & 6.94\% & -4.13\%\\
US-Energy & Volatility & 44.03\% & 42.77\% & 39.5\% & 36.52\%\\
US-Metals & Mean return & *7.99\% & ***24.77\% & 9.34\% & 1.98\%\\
\addlinespace
US-Metals & Volatility & 25.12\% & 26.64\% & 31.73\% & 21.24\%\\
US-Commodity & Mean return & **5.73\% & ***18.9\% & ***8.01\% & -0.55\%\\
US-Commodity & Volatility & 31.73\% & 30.84\% & 33.72\% & 27.56\%\\
UK-Commodity & Mean return & 2.95\% & ***23.91\% & 4.25\% & 4.87\%\\
UK-Commodity & Volatility & 20.33\% & 33.43\% & 35.99\% & 22.33\%\\
\bottomrule
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:mean_returns_aggregate_CHP}This table shows descriptive statistics for nearby returns on various commodity groups as well as for mimicking portfolios for risk factors where the factor portfolios are constructed using the whole cross-section of the U.S. traded commodities considered in the study. For each period the results are shown for periods of high (contango) and low (backwardation) aggregate CHP regimes. Figures labelled $***$ ($**$, $*$) are significant at the 1\% (5\%, 10\%) level. See section 2 for more details.}\\
\toprule
Assets & CHP regime & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Assets & CHP regime & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
US-Agriculturals-Grains & High & Mean return & 4.25\% & ***29.78\% & -2.92\% & -0.55\%\\
US-Agriculturals-Grains & High & Volatility & 27.61\% & 30.17\% & 36.06\% & 24.93\%\\
US-Agriculturals-Grains & Low & Mean return & 3.24\% & 8.26\% & *13.66\% & -7.76\%\\
US-Agriculturals-Grains & Low & Volatility & 26.67\% & 28.09\% & 30.73\% & 25.43\%\\
US-Agriculturals-Livestock & High & Mean return & -2.15\% & -1.68\% & 4.05\% & -15.8\%\\
\addlinespace
US-Agriculturals-Livestock & High & Volatility & 29.3\% & 20.65\% & 23.59\% & 25.46\%\\
US-Agriculturals-Livestock & Low & Mean return & 9.84\% & 13.43\% & *15.04\% & 16.28\%\\
US-Agriculturals-Livestock & Low & Volatility & 23.38\% & 22.47\% & 24.29\% & 27.48\%\\
US-Agriculturals-Softs & High & Mean return & -1.31\% & ***24.85\% & 11.43\% & -3.34\%\\
US-Agriculturals-Softs & High & Volatility & 32.54\% & 30.48\% & 36.06\% & 28.72\%\\
\addlinespace
US-Agriculturals-Softs & Low & Mean return & 4.47\% & 3.67\% & 7.62\% & 9.33\%\\
US-Agriculturals-Softs & Low & Volatility & 34.4\% & 29.84\% & 35\% & 27.42\%\\
US-Metals-Industrial & High & Mean return & -0.45\% & **51.48\% & 1.3\% & -13.36\%\\
US-Metals-Industrial & High & Volatility & 20.55\% & 31.58\% & 36.97\% & 18.91\%\\
US-Metals-Industrial & Low & Mean return & 5.3\% & 14.86\% & 12.22\% & 16.2\%\\
\addlinespace
US-Metals-Industrial & Low & Volatility & 19.89\% & 29.2\% & 25.31\% & 20.85\%\\
US-Metals-Precious & High & Mean return & -3.38\% & ***37.17\% & -2.16\% & -3.58\%\\
US-Metals-Precious & High & Volatility & 28.72\% & 23.7\% & 32.66\% & 22.64\%\\
US-Metals-Precious & Low & Mean return & ***22.26\% & 8.36\% & **22.23\% & 7.9\%\\
US-Metals-Precious & Low & Volatility & 23.28\% & 27.26\% & 30.77\% & 20.42\%\\
\addlinespace
US-Agriculturals & High & Mean return & 0.74\% & ***21.45\% & 4.21\% & -4.72\%\\
US-Agriculturals & High & Volatility & 30\% & 28.67\% & 33.93\% & 26.61\%\\
US-Agriculturals & Low & Mean return & 5.06\% & 7.46\% & **11.52\% & 3.88\%\\
US-Agriculturals & Low & Volatility & 29.44\% & 27.79\% & 31.39\% & 26.67\%\\
US-Energy & High & Mean return & 3.9\% & *29.75\% & -3.1\% & -23.15\%\\
\addlinespace
US-Energy & High & Volatility & 43.76\% & 45.83\% & 45.46\% & 40.32\%\\
US-Energy & Low & Mean return & *23.3\% & *26.93\% & *17.04\% & 15.07\%\\
US-Energy & Low & Volatility & 44.3\% & 39.55\% & 32.39\% & 32.06\%\\
US-Metals & High & Mean return & -2.79\% & ***40.04\% & -1.47\% & -5.54\%\\
US-Metals & High & Volatility & 27.27\% & 25.45\% & 33.54\% & 21.93\%\\
\addlinespace
US-Metals & Low & Mean return & ***18.86\% & 9.66\% & ***20.22\% & 9.56\%\\
US-Metals & Low & Volatility & 22.65\% & 27.63\% & 29.75\% & 20.49\%\\
US-Commodity & High & Mean return & 0.53\% & ***26.68\% & 1.81\% & *-7.96\%\\
US-Commodity & High & Volatility & 32.19\% & 31.59\% & 36.02\% & 28.52\%\\
US-Commodity & Low & Mean return & ***10.98\% & **11.16\% & ***14.25\% & *6.93\%\\
\addlinespace
US-Commodity & Low & Volatility & 31.25\% & 30.04\% & 31.22\% & 26.52\%\\
UK-Commodity & High & Mean return & -1.71\% & ***41.22\% & -1.32\% & **-13.66\%\\
UK-Commodity & High & Volatility & 21.05\% & 33.54\% & 38.82\% & 21.69\%\\
UK-Commodity & Low & Mean return & *8.35\% & 6.43\% & 10.04\% & ***23.4\%\\
UK-Commodity & Low & Volatility & 19.45\% & 33.17\% & 32.8\% & 22.69\%\\
\addlinespace
CHP factor & High & Mean return & 1.45\% & 7.88\% & 0.89\% & -3.48\%\\
CHP factor & High & Volatility & 8.08\% & 8.28\% & 9.81\% & 7.11\%\\
CHP factor & Low & Mean return & -0.98\% & -5.63\% & -1.2\% & -0.47\%\\
CHP factor & Low & Volatility & 7.43\% & 7.75\% & 7.97\% & 7.81\%\\
Market factor & High & Mean return & 0.67\% & ***26.46\% & 1.81\% & -7.96\%\\
\addlinespace
Market factor & High & Volatility & 10.27\% & 11.96\% & 18.97\% & 11.24\%\\
Market factor & Low & Mean return & **11.43\% & 11.16\% & 14.25\% & 6.93\%\\
Market factor & Low & Volatility & 10.1\% & 12.78\% & 15.37\% & 9.19\%\\
OI factor & High & Mean return & -0.76\% & 4.89\% & -6.9\% & 3.84\%\\
OI factor & High & Volatility & 7.92\% & 8.79\% & 8.64\% & 6.7\%\\
\addlinespace
OI factor & Low & Mean return & 0.46\% & 0.72\% & -1.65\% & 0.7\%\\
OI factor & Low & Volatility & 8.66\% & 7.52\% & 7.53\% & 6.7\%\\
TS factor & High & Mean return & **-13.59\% & -7.38\% & -8.14\% & *-11.23\%\\
TS factor & High & Volatility & 10.38\% & 8.71\% & 10.76\% & 8.39\%\\
TS factor & Low & Mean return & -4.6\% & -6.92\% & -3.37\% & -6.79\%\\
TS factor & Low & Volatility & 10.08\% & 7.94\% & 9\% & 8.69\%\\*
\end{longtabu}
\end{landscape}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:fact_regs}This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby returns on the commodities forming the entire cross-section of U.S. traded commodities considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using whole set of U.S. traded commodities considered. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes independently. See section 2 for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
US commodities & \textasciitilde{} CHP factor & High & 1.54\% & 5.5\% & 5.89\% & 3.73\%\\
US commodities & \textasciitilde{} CHP factor & Low & 1.39\% & 5.64\% & 7.33\% & 5.43\%\\
US commodities & \textasciitilde{} CHP factor & None & 1.37\% & 5.43\% & 5.99\% & 4.47\%\\
US commodities & \textasciitilde{} Market factor & High & 13.04\% & 16.96\% & 30.45\% & 17.62\%\\
US commodities & \textasciitilde{} Market factor & Low & 11.61\% & 19.83\% & 28.92\% & 13.53\%\\
\addlinespace
US commodities & \textasciitilde{} Market factor & None & 12.2\% & 18.55\% & 29.45\% & 15.57\%\\
US commodities & \textasciitilde{} OI factor & High & 0.82\% & 1.32\% & 1.19\% & 0.31\%\\
US commodities & \textasciitilde{} OI factor & Low & 1.42\% & -0.39\% & 0.54\% & 0.56\%\\
US commodities & \textasciitilde{} OI factor & None & 0.66\% & 0.36\% & 0.31\% & 0.31\%\\
US commodities & \textasciitilde{} TS factor & High & 2.12\% & 3.36\% & 4.68\% & 2.69\%\\
\addlinespace
US commodities & \textasciitilde{} TS factor & Low & 2.88\% & 2.72\% & 5.57\% & 3.09\%\\
US commodities & \textasciitilde{} TS factor & None & 2.4\% & 2.7\% & 4.69\% & 2.52\%\\
\bottomrule
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:factor_picks}For each risk factor and period, this table displays the list of corresponding commodity factor picks along with the proportion of time they are individually held on the long and short legs of the corresponding risk factor mimicking portfolio. The factor portfolios are constructed from the whole set of the twenty four U.S. traded commodities considered which is also used as asset pool for the commodity picking exercise. See section 2 for more details.}\\
\toprule
Factor & Period & Asset & Long & Short\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor & Period & Asset & Long & Short\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
CHP & 1997/2003 & XCBT - Oats & 97\% & 0\%\\
CHP & 1997/2003 & XCBT - Soybean oil & 42\% & 32\%\\
CHP & 1997/2003 & XCBT - Wheat-SR & 29\% & 9\%\\
CHP & 1997/2003 & XNYM - Copper & 43\% & 18\%\\
CHP & 1997/2003 & XNYM - Platinum & 91\% & 4\%\\
\addlinespace
CHP & 1997/2003 & XNYM - Silver & 100\% & 0\%\\
CHP & 1997/2003 & IFUS - Coffee & 22\% & 17\%\\
CHP & 1997/2003 & IFUS - Orange juice & 67\% & 18\%\\
Market & 1997/2003 & XCBT - Corn & 100\% & 0\%\\
Market & 1997/2003 & XCBT - Oats & 100\% & 0\%\\
\addlinespace
Market & 1997/2003 & XCBT - Soybean meal & 100\% & 0\%\\
Market & 1997/2003 & XCBT - Wheat-SR & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Crude oil & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Gasoline & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Heating oil & 100\% & 0\%\\
\addlinespace
Market & 1997/2003 & XNYM - Natural gas & 99\% & 0\%\\
OI & 1997/2003 & XCME - Feeder cattle & 27\% & 39\%\\
OI & 1997/2003 & XCME - Lean hogs & 36\% & 30\%\\
OI & 1997/2003 & XNYM - Crude oil & 24\% & 28\%\\
OI & 1997/2003 & XNYM - Natural gas & 46\% & 24\%\\
\addlinespace
OI & 1997/2003 & XNYM - Gold & 39\% & 28\%\\
OI & 1997/2003 & XNYM - Palladium & 46\% & 33\%\\
OI & 1997/2003 & IFUS - Coffee & 39\% & 23\%\\
OI & 1997/2003 & XCME - Lumber & 42\% & 33\%\\
TS & 1997/2003 & XCBT - Soybean meal & 71\% & 0\%\\
\addlinespace
TS & 1997/2003 & XNYM - Crude oil & 53\% & 9\%\\
TS & 1997/2003 & XNYM - Gasoline & 61\% & 17\%\\
TS & 1997/2003 & XNYM - Heating oil & 35\% & 24\%\\
TS & 1997/2003 & XNYM - Gold & 10\% & 1\%\\
TS & 1997/2003 & XNYM - Palladium & 61\% & 0\%\\
\addlinespace
TS & 1997/2003 & XNYM - Platinum & 96\% & 0\%\\
TS & 1997/2003 & IFUS - Sugar & 60\% & 26\%\\
CHP & 2004/2007 & XCBT - Oats & 100\% & 0\%\\
CHP & 2004/2007 & XNYM - Copper & 44\% & 37\%\\
CHP & 2004/2007 & XNYM - Gold & 100\% & 0\%\\
\addlinespace
CHP & 2004/2007 & XNYM - Palladium & 100\% & 0\%\\
CHP & 2004/2007 & XNYM - Platinum & 100\% & 0\%\\
CHP & 2004/2007 & XNYM - Silver & 100\% & 0\%\\
CHP & 2004/2007 & IFUS - Coffee & 15\% & 5\%\\
CHP & 2004/2007 & IFUS - Orange juice & 66\% & 15\%\\
\addlinespace
Market & 2004/2007 & XCBT - Corn & 100\% & 0\%\\
Market & 2004/2007 & XCBT - Soybeans & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Heating oil & 100\% & 0\%\\
\addlinespace
Market & 2004/2007 & XNYM - Natural gas & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Palladium & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Silver & 100\% & 0\%\\
OI & 2004/2007 & XCME - Lean hogs & 38\% & 22\%\\
OI & 2004/2007 & XCME - Live cattle & 32\% & 35\%\\
\addlinespace
OI & 2004/2007 & XNYM - Natural gas & 32\% & 35\%\\
OI & 2004/2007 & XNYM - Palladium & 43\% & 28\%\\
OI & 2004/2007 & IFUS - Cocoa & 18\% & 36\%\\
OI & 2004/2007 & IFUS - Coffee & 34\% & 28\%\\
OI & 2004/2007 & IFUS - Cotton & 33\% & 17\%\\
\addlinespace
OI & 2004/2007 & IFUS - Sugar & 48\% & 16\%\\
TS & 2004/2007 & XNYM - Copper & 86\% & 0\%\\
TS & 2004/2007 & XCME - Feeder cattle & 86\% & 0\%\\
TS & 2004/2007 & XCME - Lean hogs & 44\% & 46\%\\
TS & 2004/2007 & XCME - Live cattle & 56\% & 23\%\\
\addlinespace
TS & 2004/2007 & XNYM - Gold & 28\% & 0\%\\
TS & 2004/2007 & XNYM - Palladium & 0\% & 6\%\\
TS & 2004/2007 & XNYM - Platinum & 95\% & 0\%\\
TS & 2004/2007 & IFUS - Cocoa & 15\% & 12\%\\
CHP & 2008/2013 & XCBT - Soybean meal & 79\% & 4\%\\
\addlinespace
CHP & 2008/2013 & XNYM - Gasoline & 46\% & 0\%\\
CHP & 2008/2013 & XNYM - Heating oil & 1\% & 18\%\\
CHP & 2008/2013 & XNYM - Gold & 96\% & 0\%\\
CHP & 2008/2013 & XNYM - Palladium & 100\% & 0\%\\
CHP & 2008/2013 & XNYM - Platinum & 100\% & 0\%\\
\addlinespace
CHP & 2008/2013 & XNYM - Silver & 90\% & 0\%\\
CHP & 2008/2013 & IFUS - Orange juice & 96\% & 0\%\\
Market & 2008/2013 & XCBT - Corn & 100\% & 0\%\\
Market & 2008/2013 & XCBT - Oats & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Copper & 100\% & 0\%\\
\addlinespace
Market & 2008/2013 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Heating oil & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Palladium & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Silver & 100\% & 0\%\\
\addlinespace
OI & 2008/2013 & XCBT - Corn & 21\% & 37\%\\
OI & 2008/2013 & XCBT - Soybean meal & 37\% & 34\%\\
OI & 2008/2013 & XNYM - Crude oil & 23\% & 31\%\\
OI & 2008/2013 & XNYM - Gasoline & 33\% & 27\%\\
OI & 2008/2013 & XNYM - Heating oil & 27\% & 34\%\\
\addlinespace
OI & 2008/2013 & IFUS - Orange juice & 32\% & 43\%\\
OI & 2008/2013 & IFUS - Sugar & 36\% & 31\%\\
OI & 2008/2013 & XCME - Lumber & 35\% & 38\%\\
TS & 2008/2013 & XCBT - Soybean meal & 69\% & 1\%\\
TS & 2008/2013 & XCBT - Soybeans & 59\% & 0\%\\
\addlinespace
TS & 2008/2013 & XNYM - Crude oil & 25\% & 30\%\\
TS & 2008/2013 & XNYM - Gasoline & 61\% & 6\%\\
TS & 2008/2013 & XNYM - Heating oil & 33\% & 6\%\\
TS & 2008/2013 & XNYM - Palladium & 38\% & 0\%\\
TS & 2008/2013 & XNYM - Platinum & 47\% & 1\%\\
\addlinespace
TS & 2008/2013 & XNYM - Silver & 64\% & 0\%\\
CHP & 2014/2017 & XNYM - Copper & 0\% & 79\%\\
CHP & 2014/2017 & XNYM - Gasoline & 13\% & 0\%\\
CHP & 2014/2017 & XNYM - Gold & 78\% & 0\%\\
CHP & 2014/2017 & XNYM - Palladium & 100\% & 0\%\\
\addlinespace
CHP & 2014/2017 & XNYM - Platinum & 100\% & 0\%\\
CHP & 2014/2017 & XNYM - Silver & 69\% & 0\%\\
CHP & 2014/2017 & IFUS - Orange juice & 62\% & 2\%\\
CHP & 2014/2017 & XCME - Lumber & 38\% & 53\%\\
Market & 2014/2017 & XCBT - Soybean meal & 100\% & 0\%\\
\addlinespace
Market & 2014/2017 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Heating oil & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Natural gas & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Palladium & 100\% & 0\%\\
\addlinespace
Market & 2014/2017 & IFUS - Coffee & 100\% & 0\%\\
Market & 2014/2017 & IFUS - Sugar & 100\% & 0\%\\
OI & 2014/2017 & XCBT - Oats & 32\% & 45\%\\
OI & 2014/2017 & XCBT - Soybean meal & 39\% & 27\%\\
OI & 2014/2017 & XNYM - Copper & 48\% & 15\%\\
\addlinespace
OI & 2014/2017 & XNYM - Gold & 39\% & 23\%\\
OI & 2014/2017 & XNYM - Silver & 44\% & 28\%\\
OI & 2014/2017 & IFUS - Cocoa & 33\% & 34\%\\
OI & 2014/2017 & IFUS - Cotton & 42\% & 31\%\\
OI & 2014/2017 & XCME - Lumber & 50\% & 30\%\\
\addlinespace
TS & 2014/2017 & XCBT - Soybean meal & 68\% & 10\%\\
TS & 2014/2017 & XCBT - Soybeans & 56\% & 10\%\\
TS & 2014/2017 & XCME - Feeder cattle & 75\% & 10\%\\
TS & 2014/2017 & XCME - Live cattle & 65\% & 24\%\\
TS & 2014/2017 & IFUS - Cocoa & 54\% & 1\%\\
\addlinespace
TS & 2014/2017 & IFUS - Cotton & 40\% & 2\%\\
TS & 2014/2017 & IFUS - Orange juice & 41\% & 23\%\\
TS & 2014/2017 & XCME - Lumber & 41\% & 39\%\\*
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:correlations}This table displays the average pairwise correlations for various sets of commodities including the multiple sets of risk factor commodity picks, the whole cross-section of U.S. traded commodities considered in the study, the latter with metals omitted, and the cross-section of U.S. traded metals. For each period and commodity set the correlations are calculated independently over the whole period as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes. See section 2 for more details.}\\
\toprule
Asset pool & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
CHP factor picks & High & 0.09 & 0.24 & 0.32 & 0.22\\
CHP factor picks & Low & 0.10 & 0.28 & 0.37 & 0.19\\
CHP factor picks & None & 0.10 & 0.27 & 0.34 & 0.21\\
Market factor picks & High & 0.23 & 0.24 & 0.40 & 0.22\\
Market factor picks & Low & 0.20 & 0.28 & 0.42 & 0.19\\
\addlinespace
Market factor picks & None & 0.21 & 0.26 & 0.41 & 0.21\\
OI factor picks & High & 0.03 & 0.04 & 0.28 & 0.15\\
OI factor picks & Low & 0.03 & 0.09 & 0.22 & 0.07\\
OI factor picks & None & 0.03 & 0.07 & 0.25 & 0.11\\
TS factor picks & High & 0.16 & 0.15 & 0.41 & 0.11\\
\addlinespace
TS factor picks & Low & 0.11 & 0.17 & 0.48 & 0.09\\
TS factor picks & None & 0.13 & 0.16 & 0.44 & 0.10\\
US - Commodity & High & 0.08 & 0.11 & 0.25 & 0.13\\
US - Commodity & Low & 0.07 & 0.15 & 0.22 & 0.09\\
US - Commodity & None & 0.08 & 0.13 & 0.24 & 0.11\\
\addlinespace
US - Commodity - No metals & High & 0.09 & 0.10 & 0.26 & 0.12\\
US - Commodity - No metals & Low & 0.08 & 0.13 & 0.17 & 0.10\\
US - Commodity - No metals & None & 0.09 & 0.12 & 0.22 & 0.11\\
US - Metals & High & 0.23 & 0.55 & 0.52 & 0.44\\
US - Metals & Low & 0.19 & 0.52 & 0.61 & 0.42\\
US - Metals & None & 0.21 & 0.53 & 0.55 & 0.43\\
\bottomrule
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:fact_pick_regs_picks_picks}This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby returns on risk factors commodity picks while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in Table 4. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes independently. See section 2 for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & 1.45\% & 20.37\% & 16.95\% & 8.81\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & 7.6\% & 21.82\% & 10.29\% & 4.84\%\\
TS factor picks & \textasciitilde{} CHP factor & None & 3.42\% & 21.2\% & 14.09\% & 5.06\%\\
TS factor picks & \textasciitilde{} Market factor & High & 28.64\% & 27.15\% & 48.8\% & 21.54\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 25.53\% & 28.87\% & 54.4\% & 21.22\%\\
\addlinespace
TS factor picks & \textasciitilde{} Market factor & None & 27.23\% & 28.26\% & 50.95\% & 21.63\%\\
TS factor picks & \textasciitilde{} OI factor & High & 0.07\% & 6.42\% & 4.08\% & 2.13\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 2.48\% & 0.81\% & 2.68\% & 0.61\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.68\% & 2.26\% & 1.81\% & 0.56\%\\
TS factor picks & \textasciitilde{} TS factor & High & 2.04\% & 7.65\% & 5.2\% & 2.98\%\\
\addlinespace
TS factor picks & \textasciitilde{} TS factor & Low & 2.22\% & 6.03\% & 2\% & 1.85\%\\
TS factor picks & \textasciitilde{} TS factor & None & 0.53\% & 6.91\% & 4.16\% & 2.02\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 3.02\% & 5.9\% & 4.56\% & 10.95\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & 1.32\% & 9.37\% & 7.5\% & 5.5\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 1.96\% & 6.96\% & 3.95\% & 6.36\%\\
\addlinespace
OI factor picks & \textasciitilde{} Market factor & High & 14.4\% & 15.55\% & 39.51\% & 24.98\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 14.99\% & 20.4\% & 32.69\% & 17.31\%\\
OI factor picks & \textasciitilde{} Market factor & None & 14.86\% & 18\% & 36.91\% & 21.08\%\\
OI factor picks & \textasciitilde{} OI factor & High & 1.67\% & 1.82\% & 0.87\% & 0.64\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 0.67\% & -0.38\% & 0.95\% & 3.52\%\\
\addlinespace
OI factor picks & \textasciitilde{} OI factor & None & 0.29\% & 0.77\% & 0.09\% & 0.74\%\\
OI factor picks & \textasciitilde{} TS factor & High & 3.24\% & 7.15\% & 13.59\% & 3.69\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 2.92\% & 2.7\% & 4.11\% & 3\%\\
OI factor picks & \textasciitilde{} TS factor & None & 3.07\% & 5.36\% & 9.97\% & 3.08\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 5.74\% & 15.05\% & 12.05\% & 12.46\%\\
\addlinespace
CHP factor picks & \textasciitilde{} CHP factor & Low & 7.05\% & 10.85\% & 8.7\% & 14.6\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 5.83\% & 13.23\% & 10.46\% & 12.91\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 21.09\% & 35.23\% & 41.13\% & 32.66\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 21.6\% & 38.7\% & 46.97\% & 29.72\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 21.27\% & 37.36\% & 43.49\% & 31.36\%\\
\addlinespace
CHP factor picks & \textasciitilde{} OI factor & High & 1.69\% & 3.9\% & 4.04\% & 0.74\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 3.51\% & 1.61\% & 1.03\% & 2.96\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 1.57\% & 1.64\% & 2.07\% & 0.23\%\\
CHP factor picks & \textasciitilde{} TS factor & High & 4.77\% & 7.1\% & 1.98\% & 4.78\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & 5.36\% & 5.57\% & 0.95\% & -0.21\%\\
\addlinespace
CHP factor picks & \textasciitilde{} TS factor & None & 4.54\% & 5.1\% & 1.84\% & 1.66\%\\
Market factor picks & \textasciitilde{} CHP factor & High & 7.79\% & 8.62\% & 7.1\% & 7.84\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & 4.6\% & 7.78\% & 7.63\% & 7.73\%\\
Market factor picks & \textasciitilde{} CHP factor & None & 6.06\% & 8.31\% & 6.22\% & 7.07\%\\
Market factor picks & \textasciitilde{} Market factor & High & 32.85\% & 34.83\% & 48.14\% & 33.71\%\\
\addlinespace
Market factor picks & \textasciitilde{} Market factor & Low & 30.4\% & 37.35\% & 49.31\% & 29.36\%\\
Market factor picks & \textasciitilde{} Market factor & None & 31.43\% & 36.1\% & 48.63\% & 32.31\%\\
Market factor picks & \textasciitilde{} OI factor & High & 3.74\% & 0.77\% & 2.62\% & 0.52\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.41\% & -0.45\% & 1.22\% & -0.05\%\\
Market factor picks & \textasciitilde{} OI factor & None & 1.33\% & 0.29\% & 0.78\% & 0.53\%\\
\addlinespace
Market factor picks & \textasciitilde{} TS factor & High & 6.64\% & 5.3\% & 2.45\% & 7.59\%\\
Market factor picks & \textasciitilde{} TS factor & Low & 8.53\% & 4.21\% & 11.16\% & 4.7\%\\
Market factor picks & \textasciitilde{} TS factor & None & 7.65\% & 5.2\% & 3.25\% & 6.24\%\\*
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:fact_pick_regs_picks_UScommos}This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby returns on the commodities forming the entire cross-section of U.S. traded commodities considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in Table 4. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes independently. See section 2 for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & 0.36\% & 8.95\% & 7.8\% & 3.34\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & 2.38\% & 11.41\% & 4.96\% & 1.26\%\\
TS factor picks & \textasciitilde{} CHP factor & None & 1.03\% & 10.14\% & 5.65\% & 1.66\%\\
TS factor picks & \textasciitilde{} Market factor & High & 10.52\% & 10.93\% & 26.99\% & 9.22\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 9.7\% & 13.71\% & 27.33\% & 7.99\%\\
\addlinespace
TS factor picks & \textasciitilde{} Market factor & None & 10.13\% & 12.29\% & 26.94\% & 8.74\%\\
TS factor picks & \textasciitilde{} OI factor & High & 0.07\% & 2.36\% & 1.42\% & 0.31\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 0.69\% & 0.03\% & 0.89\% & 0.23\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.22\% & 0.77\% & 0.74\% & 0.06\%\\
TS factor picks & \textasciitilde{} TS factor & High & 0.6\% & 2.8\% & 2.09\% & 1.18\%\\
\addlinespace
TS factor picks & \textasciitilde{} TS factor & Low & 0.89\% & 3.07\% & 0.73\% & 0.72\%\\
TS factor picks & \textasciitilde{} TS factor & None & 0.18\% & 2.49\% & 1.64\% & 0.93\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 0.94\% & 2.61\% & 1.57\% & 4.45\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & 0.35\% & 6.44\% & 3.39\% & 2.23\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 0.61\% & 3.7\% & 1.54\% & 2.53\%\\
\addlinespace
OI factor picks & \textasciitilde{} Market factor & High & 6.65\% & 9.49\% & 25.93\% & 13.83\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 6.51\% & 13.26\% & 21.69\% & 9.02\%\\
OI factor picks & \textasciitilde{} Market factor & None & 6.49\% & 11.42\% & 23.87\% & 11.59\%\\
OI factor picks & \textasciitilde{} OI factor & High & 0.38\% & 0.29\% & 0.36\% & 1.01\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 0.63\% & -0.58\% & 0.23\% & 1.89\%\\
\addlinespace
OI factor picks & \textasciitilde{} OI factor & None & 0.05\% & 0.08\% & 0.05\% & 0.48\%\\
OI factor picks & \textasciitilde{} TS factor & High & 1.17\% & 2.83\% & 7.46\% & 1.87\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 1.09\% & 0.63\% & 1.92\% & 1.56\%\\
OI factor picks & \textasciitilde{} TS factor & None & 1.22\% & 2.08\% & 5.16\% & 1.44\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 1.97\% & 4.78\% & 6.3\% & 3.8\%\\
\addlinespace
CHP factor picks & \textasciitilde{} CHP factor & Low & 3.02\% & 3.81\% & 3.72\% & 5.47\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 2.3\% & 4.23\% & 4.14\% & 4.39\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 10.52\% & 13.87\% & 25.07\% & 13.35\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 9.51\% & 16.21\% & 26.19\% & 11.69\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 9.96\% & 15.3\% & 25.4\% & 12.43\%\\
\addlinespace
CHP factor picks & \textasciitilde{} OI factor & High & 0.65\% & 1.39\% & 1.39\% & 0.41\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 1.46\% & 0.29\% & 0.55\% & 1.45\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 0.65\% & 0.45\% & 0.83\% & 0.12\%\\
CHP factor picks & \textasciitilde{} TS factor & High & 1.96\% & 2.85\% & 1.19\% & 1.49\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & 2.44\% & 2.87\% & 0.27\% & 0.07\%\\
\addlinespace
CHP factor picks & \textasciitilde{} TS factor & None & 2\% & 2.32\% & 0.97\% & 0.5\%\\
Market factor picks & \textasciitilde{} CHP factor & High & 2.57\% & 4.52\% & 4.29\% & 2.37\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & 1.89\% & 4.5\% & 4.17\% & 2.69\%\\
Market factor picks & \textasciitilde{} CHP factor & None & 1.99\% & 4.31\% & 3.47\% & 2.49\%\\
Market factor picks & \textasciitilde{} Market factor & High & 12.42\% & 15.19\% & 27.88\% & 14.87\%\\
\addlinespace
Market factor picks & \textasciitilde{} Market factor & Low & 11.32\% & 16.98\% & 27.68\% & 12.24\%\\
Market factor picks & \textasciitilde{} Market factor & None & 11.69\% & 16.13\% & 27.53\% & 13.42\%\\
Market factor picks & \textasciitilde{} OI factor & High & 1.63\% & 0.52\% & 1.29\% & 0.28\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.14\% & 0.43\% & 0.3\% & -0.05\%\\
Market factor picks & \textasciitilde{} OI factor & None & 0.6\% & 0.36\% & 0.52\% & 0.41\%\\
\addlinespace
Market factor picks & \textasciitilde{} TS factor & High & 2.66\% & 1.84\% & 1.09\% & 2.6\%\\
Market factor picks & \textasciitilde{} TS factor & Low & 3.18\% & 2.47\% & 5.13\% & 2.36\%\\
Market factor picks & \textasciitilde{} TS factor & None & 2.97\% & 2.18\% & 1.28\% & 2.38\%\\*
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:fact_pick_regs_picks_UKcommos}This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby returns on the six LME traded metals considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in Table 4. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes independently. See section 2 for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & -0.01\% & 2.1\% & 3.87\% & 1.05\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & -0.37\% & 17.79\% & 3.32\% & 0.8\%\\
TS factor picks & \textasciitilde{} CHP factor & None & -0.05\% & 7.89\% & 0.65\% & -0.26\%\\
TS factor picks & \textasciitilde{} Market factor & High & 0.65\% & 17.55\% & 30.72\% & 5.69\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 1.62\% & 22.57\% & 27.92\% & 0.68\%\\
\addlinespace
TS factor picks & \textasciitilde{} Market factor & None & 1.3\% & 20.58\% & 29.6\% & 3.01\%\\
TS factor picks & \textasciitilde{} OI factor & High & -0.36\% & 4.97\% & 0.31\% & -0.65\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 0.06\% & 1.29\% & -0.33\% & -0.11\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.04\% & 3.02\% & -0.02\% & -0.18\%\\
TS factor picks & \textasciitilde{} TS factor & High & 0.45\% & 4.72\% & 0.22\% & -0.22\%\\
\addlinespace
TS factor picks & \textasciitilde{} TS factor & Low & 0.01\% & -0.67\% & 0.22\% & 1.72\%\\
TS factor picks & \textasciitilde{} TS factor & None & -0.12\% & 2.11\% & 0.44\% & 1.33\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 1.63\% & 0.34\% & -0.37\% & 5.03\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & -0.04\% & 8.57\% & 3.01\% & 0.43\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 0.01\% & 3.54\% & 0.55\% & 0.72\%\\
\addlinespace
OI factor picks & \textasciitilde{} Market factor & High & -0.24\% & 4.93\% & 28.78\% & 12.91\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 3.59\% & 10.46\% & 16.76\% & 7.99\%\\
OI factor picks & \textasciitilde{} Market factor & None & 1.17\% & 7.97\% & 24.03\% & 10.37\%\\
OI factor picks & \textasciitilde{} OI factor & High & -0.18\% & -0.68\% & 0.43\% & 0.76\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 1.79\% & 0.78\% & -0.32\% & 3.01\%\\
\addlinespace
OI factor picks & \textasciitilde{} OI factor & None & 0.54\% & -0.33\% & 0.07\% & -0.12\%\\
OI factor picks & \textasciitilde{} TS factor & High & -0.26\% & -0.34\% & 7.07\% & -0.26\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 0.3\% & 1.01\% & 0.58\% & -0.31\%\\
OI factor picks & \textasciitilde{} TS factor & None & 0.17\% & 0.11\% & 4.57\% & -0.17\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 0.48\% & -0.1\% & 2.96\% & 0.46\%\\
\addlinespace
CHP factor picks & \textasciitilde{} CHP factor & Low & 0.15\% & -0.55\% & 2.7\% & 0.77\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 0.05\% & -0.43\% & 0.48\% & 0.05\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 4.22\% & 18.7\% & 26.41\% & 15.02\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 4.4\% & 30.16\% & 27.34\% & 14.77\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 4.48\% & 25.09\% & 26.83\% & 14.99\%\\
\addlinespace
CHP factor picks & \textasciitilde{} OI factor & High & 0.28\% & 6.51\% & 0\% & -0.57\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 0.41\% & 0.18\% & 0.39\% & 0.65\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 0.49\% & 3.89\% & 0.29\% & -0.1\%\\
CHP factor picks & \textasciitilde{} TS factor & High & -0.36\% & 4.19\% & -0.19\% & 2.09\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & -0.26\% & -0.52\% & -0.17\% & 0.28\%\\
\addlinespace
CHP factor picks & \textasciitilde{} TS factor & None & -0.24\% & 2.07\% & 0.13\% & 1.33\%\\
Market factor picks & \textasciitilde{} CHP factor & High & -0.22\% & 3.62\% & 0.64\% & 1.88\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & -0.33\% & 9.62\% & 1.72\% & 1.12\%\\
Market factor picks & \textasciitilde{} CHP factor & None & -0.1\% & 6.32\% & -0.08\% & 1.61\%\\
Market factor picks & \textasciitilde{} Market factor & High & -0.11\% & 8.95\% & 36.58\% & 11.77\%\\
\addlinespace
Market factor picks & \textasciitilde{} Market factor & Low & 2.08\% & 9.85\% & 29.63\% & 2.79\%\\
Market factor picks & \textasciitilde{} Market factor & None & 0.88\% & 9.93\% & 33.73\% & 7.69\%\\
Market factor picks & \textasciitilde{} OI factor & High & 0.24\% & -0.58\% & 0.84\% & -0.04\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.66\% & 0.56\% & -0.17\% & -0.42\%\\
Market factor picks & \textasciitilde{} OI factor & None & -0.04\% & 0.12\% & 0.62\% & -0.24\%\\
\addlinespace
Market factor picks & \textasciitilde{} TS factor & High & -0.38\% & 1.37\% & 2.9\% & -0.37\%\\
Market factor picks & \textasciitilde{} TS factor & Low & -0.43\% & 1.41\% & 0.32\% & 0.39\%\\
Market factor picks & \textasciitilde{} TS factor & None & -0.23\% & 1.65\% & 2.15\% & 0.12\%\\*
\end{longtabu}\endgroup{}

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:fact_pick_regs_picks_UKcommos_legs}This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby returns on the six LME traded metals considered in the study while the regressors are returns on the long and short legs of mimicking portfolios for risk factors considered independently, where the factor portfolios are constructed using the corresponding set of risk factor commodity picks. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over periods of high (contango) and low (backwardation) aggregate CHP regimes independently. See section 2 for more details.}\\
\toprule
Factor asset pool & Model & Leg & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & Leg & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & Long & High & 0.91\% & 11.5\% & 13.22\% & -0.15\%\\
TS factor picks & \textasciitilde{} CHP factor & Long & Low & 1.51\% & 29.68\% & 34.77\% & 1.21\%\\
TS factor picks & \textasciitilde{} CHP factor & Long & None & 1.39\% & 19.71\% & 20.35\% & 1.09\%\\
TS factor picks & \textasciitilde{} CHP factor & Short & High & 0.01\% & 1.47\% & 28.59\% & 4.63\%\\
TS factor picks & \textasciitilde{} CHP factor & Short & Low & -0.15\% & -0.48\% & 26.72\% & -0.11\%\\
\addlinespace
TS factor picks & \textasciitilde{} CHP factor & Short & None & 0.16\% & 0.9\% & 27.93\% & 1.23\%\\
TS factor picks & \textasciitilde{} Market factor & Long & High & 0.65\% & 17.55\% & 30.72\% & 5.69\%\\
TS factor picks & \textasciitilde{} Market factor & Long & Low & 1.62\% & 22.57\% & 27.92\% & 0.68\%\\
TS factor picks & \textasciitilde{} Market factor & Long & None & 1.3\% & 20.58\% & 29.6\% & 3.01\%\\
TS factor picks & \textasciitilde{} OI factor & Long & High & 0.36\% & 4.73\% & 23.81\% & 1.72\%\\
\addlinespace
TS factor picks & \textasciitilde{} OI factor & Long & Low & 1.54\% & 11.1\% & 30.88\% & -0.57\%\\
TS factor picks & \textasciitilde{} OI factor & Long & None & 0.99\% & 8.22\% & 25.87\% & 0.26\%\\
TS factor picks & \textasciitilde{} OI factor & Short & High & 0.56\% & 22.57\% & 20.33\% & 2.35\%\\
TS factor picks & \textasciitilde{} OI factor & Short & Low & 2.01\% & 20.14\% & 28.09\% & 0.2\%\\
TS factor picks & \textasciitilde{} OI factor & Short & None & 1.42\% & 21.62\% & 22.99\% & 1.8\%\\
\addlinespace
TS factor picks & \textasciitilde{} TS factor & Long & High & -0.14\% & 24.07\% & 19.08\% & 0.51\%\\
TS factor picks & \textasciitilde{} TS factor & Long & Low & 1.18\% & 22.59\% & 23.52\% & -0.54\%\\
TS factor picks & \textasciitilde{} TS factor & Long & None & 0.37\% & 23.8\% & 20.46\% & -0.36\%\\
TS factor picks & \textasciitilde{} TS factor & Short & High & 1.14\% & 4.16\% & 25.71\% & 3.89\%\\
TS factor picks & \textasciitilde{} TS factor & Short & Low & 0.6\% & 9.27\% & 34.98\% & 2.59\%\\
\addlinespace
TS factor picks & \textasciitilde{} TS factor & Short & None & 1.06\% & 7.24\% & 28.35\% & 4.01\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & High & 0.19\% & 8.36\% & 15.6\% & 1.65\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & Low & 3.42\% & 16.64\% & 8.36\% & 6.14\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & None & 0.03\% & 12.55\% & 13.29\% & 3.64\%\\
OI factor picks & \textasciitilde{} CHP factor & Short & High & 0.9\% & 0.79\% & 23.17\% & 13.29\%\\
\addlinespace
OI factor picks & \textasciitilde{} CHP factor & Short & Low & 0.45\% & 0.65\% & 20.93\% & 3.06\%\\
OI factor picks & \textasciitilde{} CHP factor & Short & None & 1.01\% & 1.03\% & 22.48\% & 8.42\%\\
OI factor picks & \textasciitilde{} Market factor & Long & High & -0.24\% & 4.93\% & 28.78\% & 12.91\%\\
OI factor picks & \textasciitilde{} Market factor & Long & Low & 3.59\% & 10.46\% & 16.76\% & 7.99\%\\
OI factor picks & \textasciitilde{} Market factor & Long & None & 1.17\% & 7.97\% & 24.03\% & 10.37\%\\
\addlinespace
OI factor picks & \textasciitilde{} OI factor & Long & High & -0.39\% & 1.54\% & 17.05\% & 3.59\%\\
OI factor picks & \textasciitilde{} OI factor & Long & Low & -0.18\% & 7.12\% & 16.74\% & 6.64\%\\
OI factor picks & \textasciitilde{} OI factor & Long & None & -0.22\% & 3.7\% & 17.08\% & 5.56\%\\
OI factor picks & \textasciitilde{} OI factor & Short & High & 0.01\% & 3.92\% & 24.04\% & 9.07\%\\
OI factor picks & \textasciitilde{} OI factor & Short & Low & 3.83\% & 2.41\% & 15.01\% & -1.06\%\\
\addlinespace
OI factor picks & \textasciitilde{} OI factor & Short & None & 1.55\% & 3.14\% & 20.57\% & 2.71\%\\
OI factor picks & \textasciitilde{} TS factor & Long & High & -0.24\% & 5.47\% & 22.98\% & 3.39\%\\
OI factor picks & \textasciitilde{} TS factor & Long & Low & 2.1\% & 15.43\% & 16.87\% & 0.23\%\\
OI factor picks & \textasciitilde{} TS factor & Long & None & 0.74\% & 10.28\% & 20.88\% & 1.57\%\\
OI factor picks & \textasciitilde{} TS factor & Short & High & -0.35\% & 1.99\% & 10.81\% & 8.45\%\\
\addlinespace
OI factor picks & \textasciitilde{} TS factor & Short & Low & 0.56\% & 1.8\% & 8.96\% & 0.35\%\\
OI factor picks & \textasciitilde{} TS factor & Short & None & 0.16\% & 2.25\% & 10.32\% & 4.29\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & High & 0.9\% & 10.35\% & 12.52\% & 5.81\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & Low & 3.58\% & 28.68\% & 29.36\% & 11.92\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & None & 1.87\% & 18.04\% & 18.32\% & 8.25\%\\
\addlinespace
CHP factor picks & \textasciitilde{} CHP factor & Short & High & 4.73\% & 21.64\% & 25.57\% & 13.39\%\\
CHP factor picks & \textasciitilde{} CHP factor & Short & Low & 0.15\% & 22.77\% & 21.53\% & 7.38\%\\
CHP factor picks & \textasciitilde{} CHP factor & Short & None & 2.08\% & 22.85\% & 24.23\% & 10.59\%\\
CHP factor picks & \textasciitilde{} Market factor & Long & High & 4.22\% & 18.7\% & 26.41\% & 15.02\%\\
CHP factor picks & \textasciitilde{} Market factor & Long & Low & 4.4\% & 30.16\% & 27.34\% & 14.77\%\\
\addlinespace
CHP factor picks & \textasciitilde{} Market factor & Long & None & 4.48\% & 25.09\% & 26.83\% & 14.99\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & High & 0.6\% & 1.73\% & 19.04\% & 8.75\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & Low & 1.32\% & 21.86\% & 29.51\% & 16.19\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & None & 1.03\% & 9.08\% & 22.58\% & 11.61\%\\
CHP factor picks & \textasciitilde{} OI factor & Short & High & 2.03\% & 21.75\% & 17.79\% & 8.86\%\\
\addlinespace
CHP factor picks & \textasciitilde{} OI factor & Short & Low & 3.48\% & 23.02\% & 30.31\% & 3.59\%\\
CHP factor picks & \textasciitilde{} OI factor & Short & None & 2.83\% & 22.98\% & 21.6\% & 6.75\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & High & 1.55\% & 24.47\% & 16.98\% & 4.23\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & Low & 2.4\% & 29.72\% & 21.59\% & 4.46\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & None & 1.91\% & 27.37\% & 18.54\% & 5.2\%\\
\addlinespace
CHP factor picks & \textasciitilde{} TS factor & Short & High & 2.04\% & 2.25\% & 21.16\% & 11.49\%\\
CHP factor picks & \textasciitilde{} TS factor & Short & Low & 0.13\% & 15.94\% & 28.57\% & 15.29\%\\
CHP factor picks & \textasciitilde{} TS factor & Short & None & 1.15\% & 8.14\% & 23.54\% & 12.87\%\\
Market factor picks & \textasciitilde{} CHP factor & Long & High & 0.14\% & 14.47\% & 24.85\% & 13.11\%\\
Market factor picks & \textasciitilde{} CHP factor & Long & Low & 1.14\% & 19\% & 33.01\% & 7.1\%\\
\addlinespace
Market factor picks & \textasciitilde{} CHP factor & Long & None & 0.66\% & 17\% & 27.85\% & 10.89\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & High & -0.19\% & 1.81\% & 31.9\% & 5.71\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & Low & 1.53\% & 3.86\% & 35.32\% & -0.39\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & None & 0.66\% & 2.85\% & 32.76\% & 2.67\%\\
Market factor picks & \textasciitilde{} Market factor & Long & High & -0.11\% & 8.95\% & 36.58\% & 11.77\%\\
\addlinespace
Market factor picks & \textasciitilde{} Market factor & Long & Low & 2.08\% & 9.85\% & 29.63\% & 2.79\%\\
Market factor picks & \textasciitilde{} Market factor & Long & None & 0.88\% & 9.93\% & 33.73\% & 7.69\%\\
Market factor picks & \textasciitilde{} OI factor & Long & High & -0.32\% & 2.73\% & 34.04\% & 7.23\%\\
Market factor picks & \textasciitilde{} OI factor & Long & Low & 0.07\% & 5.36\% & 35.07\% & 1.56\%\\
Market factor picks & \textasciitilde{} OI factor & Long & None & 0.07\% & 4.4\% & 34.38\% & 4.78\%\\
\addlinespace
Market factor picks & \textasciitilde{} OI factor & Short & High & -0.03\% & 6.86\% & 24.78\% & 9.62\%\\
Market factor picks & \textasciitilde{} OI factor & Short & Low & 2.28\% & 9.51\% & 27.82\% & 1.38\%\\
Market factor picks & \textasciitilde{} OI factor & Short & None & 0.43\% & 8.38\% & 25.68\% & 6.05\%\\
Market factor picks & \textasciitilde{} TS factor & Long & High & -0.03\% & 11.75\% & 30.86\% & 8.57\%\\
Market factor picks & \textasciitilde{} TS factor & Long & Low & 0.84\% & 8.74\% & 30.58\% & 4.49\%\\
\addlinespace
Market factor picks & \textasciitilde{} TS factor & Long & None & 0.48\% & 10.83\% & 30.76\% & 7.34\%\\
Market factor picks & \textasciitilde{} TS factor & Short & High & -0.43\% & 2.19\% & 25.29\% & 6.6\%\\
Market factor picks & \textasciitilde{} TS factor & Short & Low & 2.06\% & 2.77\% & 24.04\% & 0.2\%\\
Market factor picks & \textasciitilde{} TS factor & Short & None & 0.53\% & 2.81\% & 24.88\% & 3.76\%\\*
\end{longtabu}
\end{landscape}\endgroup{}

\section{Appendix}\label{appendix}

\subsection{Assets}\label{assets}

\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:appendix_assets}This table shows details on the individual commodities considered in this study including trading country, sector, subsector as well as trading exchange Market Identifier Codes (MIC, ISO 10383).}\\
\toprule
Country & Sector & Subsector & Exchange & Name\\
\midrule
US & Agriculturals & Grains & XCBT & Corn\\
US & Agriculturals & Grains & XCBT & Oats\\
US & Agriculturals & Grains & XCBT & Soybean meal\\
US & Agriculturals & Grains & XCBT & Soybean oil\\
US & Agriculturals & Grains & XCBT & Soybeans\\

US & Agriculturals & Grains & XCBT & Wheat-SR\\
US & Agriculturals & Livestock & XCME & Feeder cattle\\
US & Agriculturals & Livestock & XCME & Lean hogs\\
US & Agriculturals & Livestock & XCME & Live cattle\\
US & Agriculturals & Softs & IFUS & Cocoa\\

US & Agriculturals & Softs & IFUS & Coffee\\
US & Agriculturals & Softs & IFUS & Cotton\\
US & Agriculturals & Softs & IFUS & Orange juice\\
US & Agriculturals & Softs & IFUS & Sugar\\
US & Agriculturals & Softs & XCME & Lumber\\

US & Energy & Petroleum & XNYM & Crude oil\\
US & Energy & Petroleum & XNYM & Gasoline\\
US & Energy & Petroleum & XNYM & Heating oil\\
US & Energy & Petroleum & XNYM & Natural gas\\
US & Metals & Industrial & XNYM & Copper\\

US & Metals & Precious & XNYM & Gold\\
US & Metals & Precious & XNYM & Palladium\\
US & Metals & Precious & XNYM & Platinum\\
US & Metals & Precious & XNYM & Silver\\
UK & Metals & Industrial & XLME & Aluminium\\

UK & Metals & Industrial & XLME & Copper\\
UK & Metals & Industrial & XLME & Lead\\
UK & Metals & Industrial & XLME & Nickel\\
UK & Metals & Industrial & XLME & Tin\\
UK & Metals & Industrial & XLME & Zinc\\
\bottomrule
\end{longtabu}

\endgroup{}

\clearpage

\subsection{Market factor lemma}\label{market-factor-lemma}

\[\overline{R_{r_{mkt}, p}^{2}} = \frac{1}{n} \sum_{j} R_{r_{j}, r_{mkt}, p}^{2} = \frac{1}{n}\]
\(\overline{R_{mkt, p}^{2}}\equiv\) period \(p\) average \(R^{2}\) for
market risk factor returns.\\
\(R_{j, mkt, p}^{2}\equiv\) period \(p\) \(R^{2}\) for the regression
where the depent variable is the vector of returns on commodity \(j\)
nearby contract and the regressor is the vector of returns on the market
portfolio; \(j \in Y_{p}^{mkt}\).\\
\(Y_{p}^{mkt}\equiv\) commodity set considered for the formation of the
mimicking portfolio for market risk over period \(p\).\\
\(n\equiv\) number of commodities in \(Y_{p}^{mkt}\).

\textbf{Proof}

Week \(t\) market factor return:
\[r_{mkt, t} = \sum_{j} w_{j, p} \cdot r_{j, t}\] \(w_{j, p}\equiv\)
period \(p\) weight for commodity \(j\) in the mimicking portfolio for
market risk in returns.\\
\(r_{j, t}\equiv\) week \(t\) return for commodity \(j\).

Assumptions:\\
-
\(w_{j, p} = w_{k, p} = \frac{1}{n} \: \forall \: \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt}\).\\
-
\(\sigma_{r_{j}, p} = \sigma_{r_{k}, p} \: \forall \: \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt}\).\\
-
\(cov \left ( r_{j}, r_{k} \right )_{p} = 0 \: \forall \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt}\).

\[
\begin{array}{l c l}
  \sigma_{r_{j, p}}^{2} & = & \beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2} + \sigma_{\epsilon}^{2} \\
  \beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2} & = & \sigma_{r_{j, p}}^{2} - \sigma_{\epsilon}^{2} \\
  \frac{\beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2}}{\sigma_{r_{j, p}}^{2}} & = & \frac{\sigma_{r_{j, p}}^{2} - \sigma_{\epsilon}^{2}}{\sigma_{r_{j, p}}^{2}} \\
    \frac{\beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2}}{\sigma_{r_{j, p}}^{2}}  & = & R_{r_{j}, r_{mkt}, p}^{2} \\ \\
  \sigma_{r_{mkt, p}}^{2} & = & \sum_{j} w_{j, p}^{2} \cdot \sigma_{j, p}^{2} + 2 \cdot \sum_{j} \sum_{k \neq j} w_{j, p} \cdot w_{k, p} \cdot cov \left ( r_{j, p}, r_{k, p} \right ) \\
                          & = & \sum_{j} \left ( \frac{1}{n} \right )^{2} \cdot \sigma_{j, p}^{2} \\
                          & = & \frac{\sum_{j} \sigma_{j, p}^{2}}{n^{2}} \\ \\
  cov \left ( r_{j}, r_{mkt} \right )_{p} & = & cov \left ( r_{j}, \frac{1}{n} \sum_{k} r_{k} \right )_{p} \\
                                          & = & \frac{1}{n} \sum_{k} cov \left ( r_{j}, r_{k} \right )_{p} \\
                                          & = & \frac{1}{n} cov \left ( r_{j}, r_{j} \right )_{p} \\
                                          & = & \frac{\sigma_{r_{j}, p}^{2}}{n} \\ \\
  \beta_{j, mkt, p}^{2} & = & \left [ \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}}{\sigma_{r_{mkt, p}}^{2}} \right ]^{2} \\
                        & = & \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}^{2}}{\sigma_{r_{mkt, p}}^{4}} \\ \\
  \beta_{j, mkt, p}^{2} \cdot  \sigma_{r_{mkt, p}}^{2}  & = & \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}^{2}}{\sigma_{r_{mkt, p}}^{2}} \\
                                                        & = & \left ( \frac{\sigma_{r_{j}, p}^{2}}{n} \right )^{2} \cdot \frac{1}{\sigma_{r_{mkt, p}}^{2}} \\
                                                        & = & \frac{\sigma_{r_{j}, p}^{4}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\ \\
  R_{r_{j}, r_{mkt}, p}^{2} & = & \frac{\sigma_{r_{j}, p}^{4}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \cdot \frac{1}{\sigma_{r_{j, p}}^{2}} \\
                            & = & \frac{\sigma_{r_{j}, p}^{2}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\ \\  
                            & \therefore  & \\ \\
  \overline{R_{r_{mkt}, p}^{2}} & = & \frac{1}{n} \sum_{j} \frac{\sigma_{r_{j}, p}^{2}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\
                                & = & \frac{1}{n} \cdot \frac{1}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \cdot \sum_{j} \sigma_{r_{j}, p}^{2} \\
                                & = & \frac{1}{n} \cdot \frac{\sum_{j} \sigma_{r_{j}, p}^{2}}{n^{2} \cdot \frac{1}{n^{2}} \cdot \sum_{j} \sigma_{r_{j}, p}^{2}} \\
                                & = & \frac{1}{n} \\
\end{array}
\]

\newpage

\bibliography{BibTeX.bib}

\clearpage
\theendnotes

\end{document}