### 2024-06-11 others/paper/custom/title page.Rmd

---
documentclass: article
# classoption: letterpaper
geometry: "letterpaper, left = 1in, right = 1in, top = 1.5in, bottom = 1.5in"
link-citations: TRUE
bibliography: BibTeX.bib
biblio-style: jf
fontsize: 12pt

output:
  pdf_document:
    template: "template.tex"
    citation: natbib
    fig_caption: TRUE
    keep_tex: TRUE
    number_sections: TRUE
    latex_engine: pdflatex
    includes:
     in_header: preamble.tex

title: 'Financialization of Commodities: an Asset Pricing Perspective'

author: 
  - name: Devraj Basu
    surname: Basu (corresponding author)
    email: devraj.basu@strath.ac.uk
    affiliation: University of Strathclyde, Glasgow, UK
  - name: Olivier Bauthéac
    surname: Bauthéac
    email: olivier.bautheac@strath.ac.uk
    affiliation: University of Strathclyde, Glasgow, UK

acknowledgements: |
  We thank seminar participants at IIM Ahmedabad, IIM Calcutta, SKEMA Business School and University of Gottingen, as well as Griffith inaugural alternative investment conference attendants for their feedback. We also thank Joelle Miffre, Martijn Boons, John Fan, Robert Bianchi, Chardin Wese Simen, Andrew Kaleel and particularly Olaf Korn for their extremely useful comments and suggestions.

---

\newpage

**Disclosure**

* Devraj Basu: 
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](pics/signature - Devraj.png)\

\newpage

**Disclosure**

* Olivier Bauthéac:
  I am a co-author of this paper and wish to state that I have no conflict of interest, financial or otherwise. I do not hold paid or unpaid positions at any organization whose goals relate to the article.  
  ![](pics/signature - Olivier.png)\

\newpage