### 2024-06-11 others/paper/elsevier article/main.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Financialization of Commodities: an Asset Pricing Perspective},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Financialization of Commodities: an Asset Pricing Perspective}
    \author[strathclyde]{Devraj Basu}
   \ead{devraj.basu@strath.ac.uk}

    \author[strathclyde]{Olivier Bauthéac}
   \ead{olivier.bautheac@strath.ac.uk}

      \address[strathclyde]{University of Strathclyde, Glasgow, UK}

  \begin{abstract}
  The last two decades have seen major changes in the nature of the
  commodity markets with the influx of financial investor inflows followed
  by the financial crisis. We study the effect of these events on
  commodity futures pricing dynamics using commodity asset pricing
  techniques. We find that a non-price based factor is best able to detect
  the effects of commodity financialization, while the effects of the
  financial crisis and after are most clearly visible through the
  performance of the market factor. The effects of financialization seem
  to go beyond mechanical effects induced by the rise of indexing.
  \end{abstract}
   \begin{keyword} Commodity futures, financialization, hedging pressure, asset pricing\end{keyword}
 \end{frontmatter}

\newpage

\textbf{Disclosure}

\begin{itemize}
\tightlist
\item
  Devraj Basu: I am a co-author of this paper and wish to state that I
  have no conflict of interest, financial or otherwise. I do not hold
  paid or unpaid positions at any organization whose goals relate to the
  article.\\
  \includegraphics{Devraj.jpg}
\end{itemize}

\newpage

\textbf{Disclosure}

\begin{itemize}
\tightlist
\item
  Olivier Bauthéac: I am a co-author of this paper and wish to state
  that I have no conflict of interest, financial or otherwise. I do not
  hold paid or unpaid positions at any organization whose goals relate
  to the article.\\
  \includegraphics{Olivier.jpg}
\end{itemize}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The last twenty years have seen major upheavals in the commodity markets
with the onset of financialization\footnote{Starting around 2004, with a
  view of commodity markets as an emerging asset class and in an effort
  to hedge against inflation and seek diversification benefits
  (Büyükşahin and Robe, 2014; Singleton, 2013), institutional investors
  sent forth an unprecedented flow of capital into commodity futures
  markets. This phenomenon is commonly referred to as the
  financialization of commodities (Domanski and Heath, 2007).} as well
as the period of the 2008 financial crisis and its aftermath.
Financialization appears to have affected a number of economic sectors,
including agriculture and energy, and its impact has been hotly debated
in the policy, legislative, regulatory, and academic spheres. A vigorous
legislative debate has led to regulatory changes accompanied by an
equally intense policy and academic debate across a number of
disciplines about the perceived effects on commodity prices of the
unprecedented inflow of institutional funds brought into commodity
futures markets by new index type investors\footnote{See section
  \ref{background} for a detailled discussion.}. In response to
immediate policy concerns, the academic debate was initially framed
rather narrowly. In this paper we take a fundamentally different
empirical approach that is broader in scope. We consider the entire
cross-section of actively traded commodities on futures markets and use
a futures based asset pricing framework that includes factors
constructed using both price and non-price attributes.

Commodity futures markets have a history of being the classic physical
commodity risks shifting conduit for commodity specialists, and
financialization may have altered this long standing paradigm with the
entry of long only commodity index traders (CITs) with very different
hedging and volume demands from traditional participants. This change
could have altered the hedging ontology in these markets and our asset
pricing framework is well suited to detect these more subtle effects
driven by changes in the nature of traders. Factor model techniques are
well suited to isolating common driving factors and detecting
co-movement and building on theory we develop a bespoke commodity-based
asset pricing framework that seems well adapted to study these two
issues simultaneously.

The first factor model we consider (CHP) is based on the well-known
hedging pressure based theory (Anderson and Danthine, 1983; Chang, 1985;
Cootner, 1960; Dusak, 1973; Hicks, 1939; Hirshleifer, 1990, 1989, 1988;
Keynes, 1930; Kolb, 1992) which postulates that futures prices for a
given commodity are inversely related to the extent that commercial
hedgers are short or long and we use a factor mimicking portfolio that
captures the impact of hedging pressure as a systemic factor (Basu and
Miffre, 2013). Our second factor model is based on open interest growth
following Hong and Yogo (2012) who find that open interest growth has
predictive power for individual commodity returns and again we construct
a long-short factor mimicking portfolio. Our third model is based on
term structure following Szymanowska et al. (2014) and Fuertes et al.
(2015) who demonstrate that the term structure factor has explanatory
power for the cross-section of commodity returns. We finally consider an
equally weighted portfolio of commodities as a proxy for the market
factor. Our two price based and two non-price based factors allow us to
compare and contrast different channels that could have led to changes
in commodity price dynamics over the last twenty years.

We study the performance of these models on a set of twenty-four US
traded commodities and six UK traded (LME exchange) base metals over the
1997/2017 period that we divide into four distinct sub-periods. The
first, the pre-financialization phase, spans the 1997/2003 period; the
second, the financialization phase, spans the 2004/2007 period\footnote{Starting
  point based on earlier studies (Baker, 2014; Christoffersen et al.,
  2014).}; the third, 2008/2013, encompasses both the financial crisis
and the loose monetary policy regimes that followed; and the fourth
spans the period starting in 2014 to present.\\
Consistent with earlier studies, we observe that mean returns for the US
traded commodities are sharply higher over the financialization period
compared to pre-financialization and this pattern extends to UK traded
metals.

The cross-sectional nature of financialization (Basak and Pavlova, 2016;
Cheng and Xiong, 2014) leads us to extend the Keynesian individual
hedging pressure paradigm by incorporating aggregate CHP (\textbf{CHP}),
a market wide measure of hedging pressure first considered in Hong and
Yogo (2012). This paradigm implies that returns for individual
commodities should be high in periods where \textbf{CHP} is low
(\textbf{backwardation}) and low in periods where \textbf{CHP} is high
(\textbf{contango}). Over the first period the cross sectional average
returns during \textbf{backwardation} for both the US and UK commodities
are positive with those for the US statistically significant, and
significantly higher than during \textbf{contango}, providing broad
support for an aggregate version of the Keynesian hedging pressure
hypothesis. Over the second period this pattern reverses, with both the
US and UK commodities having significant positive returns during
\textbf{contango} with these returns being significantly higher than
during **backwardation\}. This change in pattern suggests that the onset
of financialization altered the traditional risk shifting nature of
hedger's behaviour, which is implied by the Keynesian hypothesis. The
arrival of financial investors raises the question of whether they
functioned as liquidity providers or liquidity demanders. Arguments have
been made for both of these possibilities with Moskowitz et al. (2012)
arguing that financial investors provide liquidity while Kang et al.
(2017) argue that they demand liquidity. Our analysis of returns in
periods of low and high aggregate CHP before and during financialization
allows us to analyse this issue from a new perspective and suggests that
financial investors may have been liquidity demanders leading hedgers to
become liquidity providers.

Goldstein and Yang (2017) as well as Goldstein et al. (2014) argue that
the arrival of financial investors has an impact on the extent of normal
backwardation in commodity futures markets. Based on their analysis we
would expect our average CHP factor, which is based on backwardation
measured by hedger's positions, as well our term structure factor, which
is based on backwardation measured by roll yield, to be best suited to
measure the impact of financialization.\\
We analyze the pricing dynamics over the different periods by first
studying the time series pricing performance of the various factors on
all the US traded commodities with the factors constructed using the
latter set of commodities. The market based factor outperforms the other
three in the first period and it has a similar level of performance in
the second period, while the CHP factor shows the greatest improvement
in performance in the second period. The performance of the market
factor is driven by agricultural commodities and metals in both the
first and second period and the dramatic improvement of the CHP factor
in the second period is driven by precious and precious-base metals. The
best performing commodities for the market factor exhibited
approximately the same level of price co-movement both before and during
financialization. In contrast, the best performing commodities for the
CHP factor exhibited very low pair-wise correlation in the first period,
around the average for US traded commodities, while during
financialization the average pairwise correlation between the top
performing CHP factor commodities is higher than that for the market
factor and double of the average for all the US traded commodities. This
increase is due to the dramatic rise in average pairwise correlation
amongst the US traded metals all of which exhibited high
R\textsuperscript{2} with respect to the CHP factor during
financialization which in turn is due to their common low level of CHP
over this period. Thus the CHP factor model is able to establish the
strengthening link between a non-price attribute and price dynamics in
the US metal sub-sector during financialization.\\
We explore this issue further by constructing the four factors with the
top eight commodities for a given factor (factor picks) in each
period\footnote{See section \ref{methods} for further details.} and
evaluating its performance on these selected commodities as well as on
all US traded commodities. The results of this exercise strongly suggest
that the hedging pressure co-movement and price co-movement link was
concentrated in the metals sub-sector.

The UK traded base metals provide a useful set of assets to test how
widespread were the changes in pricing dynamics brought about by the
onset of financialization. To that end, we examine the performance of
the four factors on the six UK metals, when constructed from each set of
factor picks. The resuslts show that here was co-movement in global
metals returns in the second period and that this co-movement could be
detected by our hedging pressure based model.

The onset of the financial crisis and the monetary policy regimes that
followed, on the other hand, also appear to have induced significant
changes in pricing dynamics. Returns on both US and UK traded
commodities fell dramatically over both the third and fourth periods
with the pattern of return for the US traded commodities across
\textbf{CHP} regimes reverting to Keynesian paradigm. The pricing
results indicate the emergence of a systematic factor across the entire
cross-section of the US commodities as well as strong evidence of
cross-market linkages.

Financialization was an issue of such policy importance that it
triggered legislative action. The debate was initially framed around
adequacy of speculation, the burning issue of the day, and the
consequent analysis focused on the more mechanical effects of
financialization. With the benefit of hindsight, it appears this
approach was perhaps too narrow and it now seems necessary to address
financialization from a broader perspective. Commodity price dynamics
appear to have altered substantially in quite different ways over the
financialization and the financial crisis periods and our commodity
futures based asset pricing approach seems able to provide new insights
into the nature of these changes: financialization was a phenomenon
endogenous to the commodity markets transmitted via the commodity
futures markets and mainly concentrated in a specific sector (metals)
while the crisis and its aftermath seems to have delivered an exogenous
shock across the entire cross-section of global commodities. The initial
view of financialization was that it consisted of speculative flows
which had the effect of driving up and creating bubble like conditions
in commodity future prices\footnote{De Schutter (2010); Gilbert (2010a);
  Gilbert (2010b); Masters (2008); Masters and White (2008); Herman et
  al. (2011); Schumann (2011); Singleton (2013); UNCTAD and Cooperation
  (2009).}. However, the fundamental question of the nature of the
impact of financialization across the entire cross-section of commodity
futures markets has not yet been completely answered. We provide an
empirical complement to a new stream of theoretical studies that try to
model the impact of financialization on various aspects of commodity
futures markets\footnote{Etula (2013), Acharya et al. (2013), Cheng et
  al. (2014), Leclercq and Praz (2014), Sockin and Xiong (2015),
  Goldstein et al. (2014), Ekeland et al. (2016), Goldstein and Yang
  (2017).} and demonstrate that the effects of financialization extend
beyond the mechanical effects induced by indexation as outlined in Basak
and Pavlova (2016) as well as Tang and Xiong (2012).

\hypertarget{data-methods}{%
\section{Data \& methods}\label{data-methods}}

\hypertarget{background}{%
\subsection{Background}\label{background}}

The period from 1998 to 2008 saw an influx of investment into commodity
index linked products most of which made its way into the commodity
futures market. These commodity indexes whose goal was to track the
broad movement of commodity prices could be accessed by means of
exchange traded funds and exchange traded notes. At least \$100 billion
of new investment flowed into these products over the 1998-2008 period
with this financialization phenomenon leading to substantial legislative
changes. Concerns over the consequences of financialization were behind
Rule 76 FR 4752 issued by the US Commodity Futures Trading Commission
(CFTC) on January 26\textsuperscript{th}, 2011. This provision emanates
from the Dodd-Frank Wall Street and Consumer Protection Act of 2010
(Title VII, Section 737) that mandates the CFTC to use position limits
to restrict the flow of speculative capital into a number of commodity
markets. The Rule was approved in a close 3-2 vote and the ensuing
rulemaking process was extremely contentious with several commissioners
(Michael Dunn and Scott D. O'Malia in particular) expressing
reservations about the lack of supporting evidence and the Rule also
triggering thousands of comment letters as well as a lawsuit against the
CFTC spearheaded by two Wall Street trade groups, the International
Swaps \& Derivatives Association (ISDA) and the Securities Industries \&
Financial Markets Association (SIFMA)\footnote{ISDA \& SIFMA v US CFTC;
  Complaint, 1:11-cv-02146; December 2\textsuperscript{nd}, 2011.}.

A world-wide debate has ensued about the role of index funds in
commodity markets. The first responses to the 2007/2008 crisis of
escalating food and energy prices took the form of policy reports, many
of which reasoned that the growth of commodity index funds came along
with an influx of largely speculative capital that was responsible for
driving commodity prices beyond their historic highs (De Schutter, 2010;
Gilbert, 2010b; Herman et al., 2011; Schumann, 2011; UNCTAD and
Cooperation, 2009). The US and various european countries also became
involved in this issue. Early US Senate investigations drew on pricing
and trading data supplied by the CFTC as well as interviews with
numerous experts including hedge fund manager Michael Masters who linked
the growing presence of index investors in US wheat markets to the
observed surges in both futures and cash prices (Masters, 2008; Masters
and White, 2008)). This contention has since come to be known as the
Masters' hypothesis and was largely endorsed by the final US Senate
report on the issue (Senate, 2009). At this stage, the question thrust
on the academic sphere was whether ``excessive speculation'' (understood
as the market activity peculiar to index type investors) was linked to
escalating energy and agricultural commodity prices. A number of
academic studies, most of which used commodity specific causality and
correlation based analysis have disproved this contention, particularly
in the case of agricultural and energy commodities. Irwin et al. (2009),
Sanders et al. (2010), Sanders and Irwin (2011a), Sanders and Irwin
(2011b), Irwin (2013), Brunetti and Reiffen (2014), Hamilton and Wu
(2015) and Bruno et al. (2017) study the agricultural markets.
Büyükşahin and Harris (2011), Tokic (2012), Fattouh et al. (2013),
Kilian and Murphy (2014), Knittel and Pindyck (2016) and Manera et al.
(2016) study the energy markets. Bohl and Stephan (2013), Kim (2015) and
Boyd et al. (2016) study both energy and agricultural markets while
Irwin and Sanders (2011), Irwin and Sanders (2012) and Stoll and Whaley
(2011) study the commodity markets in general.

Other studies have underlined various alterations in commodity pricing
dynamics over the period. Gilbert and Pfuderer (2014) reject the view
that financialization has not had any effects in the grains markets and
demonstrate that trades originated by financial actors, and specifically
index investors, can move prices but tend typically to be
volatility-reducing. Cheng and Xiong (2014) show that in the case of
index commodities, financialization has transformed the risk sharing,
and information discovery functions of commodity futures markets.
Juvenal and Petrella (2015) contend that the oil price increase between
2004 and 2008 was mainly driven by the strength of global demand but
that the financialization process of commodity markets also played a
role. Henderson et al. (2015) show that non-information-based financial
investments have important impacts on commodity prices.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Commodity index funds (CIFs), commodity based exchange traded funds
(ETFs) and notes (ETNs) as well as commodity linked notes (CLNs) were
amongst the main investment vehicules that allowed institutional and
retail investors to build up commodity exposure in the early 2000s
(Boons et al., 2012; Henderson et al., 2015; Irwin and Sanders, 2011;
Schumann, 2011; UNCTAD and Cooperation, 2009). Most of these products
were primarily designed to track prominent commodity indexes amongst
which the S\&P-GSCI, the most popular. Most of its constituents at the
time of writting form the basis of the broad cross-section of
commodities that we consider in this study\footnote{See section
  \ref{assets} for details on the commodity pool considered.}. Unleaded
gasoline, the NYMEX legacy contract for gasoline stopped trading in 2006
when it was replaced by Reformulated Gasoline Blendstock for Oxygen
Blending (RBOB) with the two futures series trading alongside for most
of the year. For our NYMEX gasoline data series, we consider unleaded
gasoline up to September 1\textsuperscript{st}, 2006 and RBOB gasoline
thereafter, date at which liquidity for the latter overtook that for the
former.

The commodity futures trading commission (CFTC) operates a comprehensive
system of collecting information on market participants known as the
large trader reporting program (LTRP) where it records and reports to
the public market position data on market particpants who have position
levels in excess of a particular futures market specific
threshold\footnote{Current reporting level thresholds available at:
  \href{https://www.ecfr.gov/cgi-bin/retrieveECFR?gp=\&SID=970471b8455f4bab7db4110cfde50731\&mc=true\&r=SECTION\&n=se17.1.15_103}{www.ecfr.gov}.}.
The commission collects market data and position information daily from
clearing members, futures commission merchants (FCMs) and foreign
brokers and publishes corresponding summaries weekly in a series of
market specific reports. In these reports, individual traders are
categorised according to the nature of their trading activity on the
basis of self-reported information\footnote{CFTC form 40. Available at:
  \href{https://www.cftc.gov/sites/default/files/idc/groups/public/@forms/documents/file/cftcform40.pdf}{www.cftc.gov}.}
that is subject to review by CFTC staff for reasonableness. Position
information is then aggregated by category and asset type with each
report providing an aggregation for a set of report specific categories
and, for each category, a breakdown between futures only positions and
positions in futures and options combined.\\
In its legacy report, the commitment of trader report (COT) with data
dating back to 1962, the CFTC distinguishes ``commercial'' and
``non-commercial'' market participants where a ``commercial''
participant is defined as one ``{[}\ldots{}{]} engaged in business
activities hedged by the use of futures and option markets''\footnote{See
  \href{https://www.gpo.gov/fdsys/pkg/CFR-1998-title17-vol1/xml/CFR-1998-title17-vol1-sec1-3.xml}{CFTC
  regulation 1.3, 17 CFR 1.3(z)} for details.}. A third category,
``non-reportable'', aggregates positions for participants not meeting
the reporting threshold. In response to concerns related to category
accuracy\footnote{See for example Ederington and Lee (2002).} the CFTC
now refines its classification in the disaggregated commitment of trader
report (DCOT) where participant categories include
``Producer/Merchant/Processor/User'', ``Swap dealer'', ``Money manager''
and ``Other reportable'' as well as in the supplemental commitment of
trader report (SCOT) that details commodity index trader aggregate
positions for thirteen agricultural commodity futures markets. Data for
DCOT and SCOT dates back to 2006 and 2007 respectively.\\
We study the 1997/2017 period where we observe futures only position
data from the COT report, the only CFTC report with data available for
the whole period of interest; as well as futures term structure price
and open interest data from Bloomberg.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

On the one hand, futures markets have a long history of serving
commodity producers to allievate their commodity price and output risks.
Keynes (1930) and Hicks (1939) emphasise the ``normal'' behaviour of
naturally short hedgers outsourcing their commodity risk to naturally
long commodity specialist speculators. This market configuration is
referred to as ``backwardation'' where futures is expected to trade at a
discount to spot, providing speculators with incentives to take the long
side. The opposite market paradigm is referred to as ``contango''.
Working (Working, 1953, 1948) and Cootner (Cootner, 1967, 1960) somehow
relax this ``supply-of-speculative-services'' (Till, 2007) approach by
introducing processors and merchants, also commodity experts, on the
hedgers side and relating the notions of backwardation and contango more
directly to the term structure via the theory of storage where the shape
of the front end of the term structure (basis) is directly related to
the market price of storage.\\
Financialization, on the other hand, refers to the entry of financial
investors into the commodity futures markets who, for the most part, are
not commodity experts. Besides, these new market participants tend to
exhibit herding like behaviour, often taking massive long-only positions
on the whole cross-section of indexed commodities in an attempt to hedge
against inflation and/or seek further diversification benefits for
portfolios with, in most cases, existing large positions across various
asset classes (Boyd et al., 2016; Brunetti et al., 2016; Cheng et al.,
2014; Juvenal and Petrella, 2015; Singleton, 2013; Tang and Xiong,
2012). This contrasts with the traditional approaches of Keynes and
Working and in this context the issues of co-movement and aggregate
market participants behaviour seem particularly relevant.\\
Asset pricing techniques are well suited to study co-movement. Building
on theory we develop an ad-hoc commodity based asset pricing framework
that we believe is well adapted to study these two issues simultaneously
and implement it over four periods of interest. The first, the
pre-financialization period, starts in 1997 and is naturally followed by
financialization, spanning the 2004/2007 period, and the financial
crisis and its aftermath, spanning the 2008/2013 period. The last period
spans the 2014/present period sarting with the beginning of the
tapering/end of the US Quantitative Easing program (QE).\\
Using the weekly price, open interest and market position data described
above we construct market portfolios of commodity nearby futures as well
as mimicking portfolios for market, commercial hedging pressure, term
structure and liquidity risk factors in returns. While our market
portfolios are long only, equally weighted portfolios of particular
commodity sets, our other mimicking portfolios for risk factors are
constructed as combinations of two sub-portfolios (legs), one held long,
one held short, which respective constituents are pooled on a weekly
basis according to risk factor specific criterias. The weekly return for
a particular mimicking portfolio is calculated as the difference between
the average return for constituents of the long portfolio and that for
the constituents of the short portfolio. i.e, return on the mimicking
portfolio for risk factor \(i\) (\(f_{i}\)) for week \(t\):

\[r_{t}^{f_{i}}=\frac{1}{x}\sum_{j=1}^{x}r_{j,t}-\frac{1}{x}\sum_{j=n-x}^{n}r_{j,t}\]
\(n\equiv\) number of commodities in the set considered for mimicking
portfolio construction.\\
\(x = \left \lceil n \cdot s \right \rceil\)\\
\(s\equiv\) selection threshold (\(\frac{1}{3}\) here).\\
\(r_{j,t}=\frac{C_{j,t}}{C_{j,t-1}}-1\)\\
\(C_{j,t}\equiv\) week \(t\) close price for commodity \(j\).\\
\(j\equiv\) commodity rank in the ordered set \(Y_{t}^{f_{i}}\).\\
\(Y_{t}^{f_{i}}\equiv\) week \(t\) ordered set of the commodities
considered for mimicking portfolio construction where the ordering rule
is specific to \(f_{i}\).

For our portfolio CHP (commercial hedging pressure), we sort the set of
commodities considered on past twenty six-week average CHP from lowest
to highest. The bottom third constituents (lowest \(\overline{CHP}\)s)
of the corresponding ordered set form the long leg of the portfolio
while the top third form the short leg:

\[Y_{t}^{CHP}\equiv\left \{ \left \{ \overline{CHP_{1}}, \overline{CHP_{2}}, ..., \overline{CHP_{n}} \right \}, \leq \right \}\]

\(\overline{CHP_{j}}=\frac{1}{26}\sum_{k=0}^{25}\frac{L_{j,t-k}}{L_{j,t-k}+S_{j,t-k}}\)\\
\(L_{j,t}\equiv\) number of long positions held by commercial hedgers
for week \(t\) on the futures series of commodity \(j\).\\
\(S_{j,t}\equiv\) number of short positions held by commercial hedgers
for week \(t\) on the futures series of commodity \(j\).

For our portfolio TS (term structure), we sort the set of commodities
considered on past twenty six-week average roll-yield from highest to
lowest. The bottom third constituents (highest \(\overline{RY}\)s) of
the corresponding ordered set form the long leg of the portfolio while
the top third form the short leg:

\[Y_{t}^{TS}\equiv\left \{ \left \{ \overline{RY_{1}}, \overline{RY_{2}}, ..., \overline{RY_{n}} \right \}, \geq \right \}\]

\(\overline{RY_{j}}=\frac{1}{26}\sum_{k=0}^{25}(\frac{F_{j,t-k}^{1}}{F_{j,t-k}^{2}} - 1)\)\\
\(F_{j,t}^{1}\equiv\) week \(t\) close price for commodity \(j\) first
nearby futures contract.\\
\(F_{j,t}^{2}\equiv\) week \(t\) close price for commodity \(j\) second
nearby futures contract.

For our portfolio OI (open interest), we sort the set of commodities
considered on past twenty six-week average aggregate open interest
growth from highest to lowest. The bottom third constituents (highest
\(\overline{\mathbf{OI}^{\Delta \%}}\)s) of the corresponding ordered
set form the long leg of the portfolio while the top third form the
short leg.

\[Y_{t}^{OI}\equiv\left \{ \left \{ \overline{\mathbf{OI}_{1}^{\Delta \%}}, \overline{\mathbf{OI}_{2}^{\Delta \%}}, ..., \overline{\mathbf{OI}_{n}^{\Delta \%}} \right \}, \geq \right \}\]

\(\overline{\mathbf{OI}_{j}^{\Delta \%}}=\frac{1}{26}\sum_{k=0}^{25}(\frac{\mathbf{OI}_{j,t-k}}{\mathbf{OI}_{j,t-k-1}} - 1)\)\\
\(\mathbf{OI}_{j,t}=\sum_{k=1}^{l_{j,t}}OI_{j,t,k}\)\\
\(OI_{j,t,k}\equiv\) open interest at close of week \(t\) for the
\(k^{th}\) nearby contract on the futures term structure for commodity
\(j\).\\
\(l_{j,t}\equiv\) length of commodity \(j\)'s futures term structure on
week \(t\) close.

We use a time series regression approach where we start with regressions
of the US traded individual commodity weekly front returns on returns to
the mimicking portfolio for a particular risk factor where the portfolio
is constructed using the whole cross-section of US traded commodities.
The corresponding adjusted \(R^{2}\)s are averaged and the process is
repeated for each of our four risk factors (market, hedging pressure,
term structure and liquidity) and periods (1997/2003, 2004/2007,
2008/2013 and 2014/2017). i.e, average adjusted \(R^{2}\) for risk
factor \(i\) (\(f_{i}\)) over period \(t\):

\[\overline{\tilde{R}_{f_{i},t}^{2}}=\frac{1}{24}\sum_{j=1}^{24}\tilde{R}_{j,t}^{2}\]
\(\tilde{R_{j}^{2}}=1-(1-R_{j}^{2})\frac{T-1}{T-p-1}\)\footnote{The
  Wherry formula-1.}\\
\(T\equiv\) length of period \(t\) (number of weeks here).\\
\(p\equiv\) number of regressors (1 here).\\
\(R_{j,t}^{2}=1-\frac{\sum_{k=1}^{T}(y_{j,k}-\hat{y_{j,k}})^{2}}{\sum_{k=1}^{T}(y_{j,k}-\overline{y_{j}})^{2}}\)\\
\(y_{j,t}\equiv\) week \(t\) nearby return for commodity \(j\).\\
\(\hat{y_{j,t}}=\beta_{0}+\beta_{1}f_{i,t}+\epsilon\)\\
\(\beta s \equiv\) classic OLS coefficient estimates.\\
\(f_{i,t}\equiv\) week \(t\) return on mimicking portfolio for risk
factor \(f_{i}\)

The same operation is repeated for the two (\(p=2\)) and three (\(p=3\))
factor model cases, where regressors include in turn all combinations of
any two and three risk factors respectivly, as well as for the four
(\(p=4\)) factor model case where all the risk factor mimicking
portfolios are included. Results are reported for the one-factor model
case with the other results available from the authors upon request.\\
For each period we also implement the analysis described above over
periods of low and high aggregate CHP (\textbf{CHP}), a notion which we
introduce in this study and which we believe is particularly relevant in
the context of financialization where issues of co-movement and
aggregate market participants' behaviour play a central role. We define
\textbf{CHP} as the cross-sectional average of CHPs for a particular
pool of assets. i.e, Week \(t\) \textbf{CHP} for asset pool \(x\):

\[\mathbf{CHP}_{t}^{x}=\frac{1}{n}\sum_{j=1}^{n}CHP_{j,t}\] \(n\equiv\)
number of commodities in set \(x\).\\
\(CHP_{i,t}\equiv\) week \(t\) CHP for commodity \(i\) as defined above.

We further define aggregate backwardation (\textbf{backwardation}) and
contango (\textbf{contango}) \textbf{CHP} regimes as, for a given
period, \textbf{CHP} levels below and above the period's median
respectively and, in contrast with (Hong and Yogo, 2012) who aggregate
across sub-sectors, the asset pool we consider for \textbf{CHP}
construction include the whole cross-section of US traded commodities.

In the next part of the analysis, our mimicking portfolios for risk
factors, constructed from the set of all US traded commodities are used
as commodity picking devices on the latter commodity set. For each risk
factor and period, individual US traded commodity front returns are
regressed on returns for the corresponding mimicking portfolio.
Individual commodities are sorted by adjusted \(R^{2}\)
(\(\tilde{R^{2}}\)) accordingly from highest to lowest and the bottom
third (highest \(\tilde{R^{2}}\)s) are selected as commodity picks for
the risk factor for the period. i.e, set of picks \(P_{t}^{f_{i}}\) for
risk factor \(i\) (\(f_{i}\)) over period \(t\):

\[P_{t}^{f_{i}}\equiv\left \{ P_{1}, P_{2}, ..., P_{x} \right \}\]
\(x = \left \lceil n \cdot s \right \rceil\)\\
\(n\equiv\) number of commodities in the set considered for the picking
excersise (twenty-four here).\\
\(s\equiv\) selection threshold (\(\frac{1}{3}\) here).\\
\(P_{j}\equiv\) commodity pick \(j\) for risk factor \(f_{i}\) over
period \(t\)\\
\(j\equiv\) rank in the ordered set \(Y_{t}^{f_{i}}\)\\
\[Y_{t}^{f_{i}}\equiv\left \{ \left \{ \tilde{R_{1}^{2}}, \tilde{R_{2}^{2}}, ..., \tilde{R_{n}^{2}} \right \}, \geq \right \}\]
\(n\equiv\) number of commodities in the set considered for picking
(twenty-four here).\\
\(\tilde{R_{j}^{2}}\equiv\) risk factor \(f_{i}\) returns' explanatory
power on commodity \(j\) nearby futures returns over period \(t\) as
defined above.

For each period and \textbf{CHP} regime we record the proportion of time
each pick is held in the long and (/or) short leg of the corresponding
risk factor mimicking portfolio as well as the average pairwise
correlation amongst every sets of picks.

In the last part of the analysis, for each period our mimicking
portfolios for risk factors in returns are constructed using each set of
commodity picks independently as the asset pool for portfolio
construction. A similar time series approach to that described above is
then implemented using these newly formed mimicking portfolios. In this
case, a first set of dependent variables includes the sets of commodity
picks themselves, a second set includes the whole cross-section of US
traded commodities while a third set includes the six UK traded metals
considered. The latter are also used as dependent variables in the last
series of regressions where the long and short legs of the newly formed
risk factor mimicking portfolios described above are considered
separately as regressors.

\hypertarget{results-discussion}{%
\section{Results \& discussion}\label{results-discussion}}

The descriptive statistics over the various time periods for our two
sets of commodities (US and non-US) as well as the four factors
considered are shown in Table \ref{table1}. The traditional Keynesian
hedging pressure paradigm postulates a negative relationship between the
return on an individual commodity and its hedging pressure.
Financialization has been seen as a cross-sectional commodity market
phenomenon; this naturally leads us to extend the individual hedging
pressure paradigm by incorporating aggregate CHP (\textbf{CHP}), a
market wide measure of hedging pressure. The extended Keynesian hedging
pressure paradigm should also imply that returns for individual
commodities should be high in periods where \textbf{CHP} is low and vice
et versa.\\
In table \ref{table2} the mean returns are shown over high
(\textbf{contango}) and low (\textbf{backwardation}) \textbf{CHP}
regimes for each time period. Over the first period eighteen of the
twenty-four US commodities have higher mean returns during periods of
\textbf{backwardation} as compared to \textbf{contango}, with the
difference being statistically significant at the 5\% level for one
commodity and for the remaining six commodities the higher mean return
during \textbf{contango} is not statistically significant for any of
them. For the six UK commodities, all metals, four of them had higher
mean returns during phases of \textbf{backwardation} as compared to
\textbf{contango} in the first period, with one of these differences
being statistically significant at the 5\% level while for the remaining
two the higher mean return during \textbf{contango} is not statistically
significant for either. An equally weighted portfolio of the twenty-four
US commodities had a mean return of 10.98\% during
\textbf{backwardation} phases, statistically significant at the 1\%
level, and a mean return of 0.93\% during \textbf{contango}, both in the
first period, with the difference between them statistically significant
at the 5\% level. The corresponding figures for an equally weighted
portfolio of the UK metals were 8.35\% and -1.71\%, the difference again
statistically significant at the 5\% level. These results provide broad
support for an aggregate version of the Keynesian hedging pressure
hypothesis and suggests that hedgers in the aggregate were engaged in
risk transfer during this period.\\
Over the second period the pattern essentially reverses. For the
twenty-four US commodities over this period, eighteen have higher mean
returns during phases of \textbf{backwardation} over \textbf{contango},
with the difference statistically significant for four commodities and
of the remaining seven the higher mean return during
\textbf{backwardation} is statistically significant for one commodity.
All of the six UK metals achieved higher mean returns over
\textbf{backwardation} than contango with the difference being
statistically significant for one commodity. The equally weighted US
portfolio had a mean return of 11.16\% over \textbf{backwardation} while
it was 26.68\% during \textbf{contango}, the first statistically
significant at the 5\% level and the second at the 1\% level, with the
difference statistically significant at the 1\% level. The difference
was considerably more pronounced for the equally weighted UK metals
portfolio with a mean return of 6.43\% during \textbf{backwardation}
rising to 41.22\% during **contango\}, with the difference also
significant at the 1\% level.\\
Thus the onset of financialization seems to have engendered a change in
the nature of hedger's behaviour in the aggregate, taking it away from
risk shifting, the traditional Keynesian view. This phenomenon has been
noted in earlier theoretical and empirical work (Danthine, 1978, p.
@stout\_why\_1998) in different context from which two competing models
of behaviour have emerged. The information arbitrage model implies that
hedgers may often be seeking out counterparties to trade with rather
than being purely passive. This issue is also raised in both Cheng and
Xiong (2014) and Stulz (1996) who point out that hedgers may be taking a
view on prices just as speculators do. In fact, by hedging away some of
their risk, hedgers are able to speculate more heavily based on their
disagreement against speculators regarding futures price movement
(Simsek, 2013), which fits into a heterogenous expectation theory of
speculation. As Stout (1998) points out, this disagreement-based theory
of trading is one of the main reason the public and the law disapproves
the speculators as this form of trading is regarded as
non-productive\footnote{Duffie (2014) also discusses some of the
  challenges faced by a policy treatment of speculative trading
  motivated by differences in beliefs.}. In the second model hedgers are
seen as liquidity providers as in Kang et al. (2017). This explanation
is also consistent with the nature of financialization which led to the
arrival of long term, long only investors who require substantial
liquidity to roll-over their positions.

The average time series pricing performance of the various one factor
models on the US traded assets over the four periods is shown in Table
\ref{table3}. Over the first period the market factor has the best
average time series performance with the highest average adjusted
R\textsuperscript{2} with all three futures market factors performing
poorly and the overall pricing performance consistent across periods of
high and low \textbf{CHP} (below period median). The two factor models
average R\textsuperscript{2}s (results available upon request) are
roughly the sum of the one factor R\textsuperscript{2}s indicating that
the factors are almost orthogonal over this period. The onset of
financialization leads to a substantial improvement in the performance
of the CHP factor with its average R\textsuperscript{2} increasing to
5.4\% in the second period relative to 1.4\% in the first. The
performance of the market factor also improves though not by as much in
a relative sense (18.6\% from 12.2\%). Both the term structure factor
and the open interest growth factor show a much smaller improvement in
pricing performance. Over all periods, all of our factor models show
similar pricing performance across regimes of high and low **CHP\}.

We next investigate the top eight of the US traded commodities which
achieved the highest time series R\textsuperscript{2} for the CHP,
market, open interest and term structure factors, in each of the four
time periods with the commodity picks shown in Table \ref{table4}. For
the CHP factor the commodities that achieve high R\textsuperscript{2} in
a given period are likely to be those with relatively high or low CHP
over that period, or those commodities that co-vary with the short or
the long side of the factor. The same applies to the open interest and
term structure factors with CHP replaced by open interest growth and
roll yield respectively; while for the market factor the picks are those
that co-vary most strongly with the market factor over that period. For
the CHP factor, we observe a change in the picks between the first and
the second period; the picks in the first period included five
agricultural commodities and three metals while in the second period
include all of the five base and precious metals and only three
agricultural commodities. This pattern is also visible for the term
structure factor for which there are four metals (gold, copper,
palladium and platinum) picks in the second period against three in the
first (gold, palladium and platinum). The market factor has two metal
picks in the second period (palladium and silver) while all of the picks
in the first period were agricultural or energy commodities. Over the
third period, the average CHP factor has four metal picks while both the
term structure factor and market factor have three. Table \ref{table4}
also shows the proportion of time that these picks were constituents of
the corresponding factor. All of the picks of the long-short factors
(average CHP, open interest growth and term structure factors) were
constituents of the corresponding factor for some proportion of time in
all of the time periods. In this respect, there is a contrast between
the first and second periods for the average CHP with the picks
appearing predominantly on the long side in the second period, as
opposed to both legs in the first. This indicates that the picks in the
second period consistently exhibited low CHP. This behaviour continued
over the third period.

In table \ref{table5}, we show the average pairwise correlation between
each set of picks in each time period. This exercise gives an indication
of the extent to which attribute co-movement reveals price co-movement,
which is particularly relevant for the long-short factors for which the
corresponding attribute is not directly related to price co-movement. In
the first period, the average CHP factor picks pairwise correlation is
similar to the average across the US traded twenty-four commodities
(around 0.1) with the market factor picks showing the highest figure
(0.21). In the second period however it is the highest (0.27),
comparable with the market (0.26) and double of the US traded
commodities average (0.13). This increase is a consequence of the
dramatic increase in average pairwise price correlation amongst the US
traded metals, all of which are CHP factor picks. Taken together with
the fact that four of these metals had consistently low CHP, we see that
CHP co-movement in the second period is strongly correlated with price
co-movement. It is also interesting to note that US traded metals
returns were negatively correlated with the rest of US traded
commodities in the first period and positively in the second (Results
available upon request).

In order to further understand the link between attribute and price
co-movement we build all of the factors from every set of factor picks
and assess their pricing performance on these sets of picks. The results
are shown in table \ref{table6}. The long-short factors for a particular
attribute select the top and bottom three of the corresponding picks
based on that attribute while the market factor takes an equally
weighted combination of the eight corresponding picks. In the absence of
price co-movement between the picks, this market factor will achieve an
average R\textsuperscript{2} of around 12.5\%\footnote{Market factor
  lemma:
  \(\overline{R_{r_{mkt}, p}^{2}} = \frac{1}{n} \sum_{j} R_{r_{j}, r_{mkt}, p}^{2} = \frac{1}{n}; \: n\equiv\)
  size of market portfolio (eight here). See section
  \ref{market-factor-lemma} for a formal proof.}, while there is no
non-zero R\textsuperscript{2} lower bound for the corresponding
long-short factors. A high average R\textsuperscript{2} for the market
factor is indicative of strong price co-movement while for the
long-short factors it is indicative of simultaneous attribute and price
co-movement. The most dramatic change in pricing performance occurs for
the average CHP factor picks where the market factor built from these
picks has an average R\textsuperscript{2} of 21.3 \% in the first period
(close to the 0 correlation situation) and jumps to 37.4 \% in the
second period. The average CHP factor built with average CHP factor
picks also shows a sharp increase in pricing performance over the second
period. There is also substancial improvement in pricing performance of
the CHP factor built with term structure picks (3.4\%
vs.~21.2\%)\footnote{In contrast to the market factor, there is no
  pricing performance lower bound for the CHP factor built with a set of
  picks.}, providing evidence of convergence of Keynesian and Working's
theories of the term structure in detecting price co-movement during
financialization. In table \ref{table7} we assess the performance of the
same factors on all the commodities instead of just the picks. While the
average R\textsuperscript{2}s are expectedly lower, the relative
performance of the market factors built with different sets of picks is
similar to table \ref{table6}. In contrast, the increase in
R\textsuperscript{2} for the average CHP factor built with average CHP
factor picks is lower indicating that the link between CHP co-movement
and price co-movement is strongest amongst the picks which are
predominantly metals.\\
The conclusion from Tables \ref{table6} and \ref{table7} is that the
onset of financialization seems to have had a major impact on the
dynamics of US traded precious and precious-base metals and we now seek
to analyse whether this phenomenon is true for the non-US metal futures.
To that end we examine the pricing performance of the factors used in
Table \ref{table6} on the six UK metals futures. The results are
reported in Table \ref{table8}. In the first period, the average
R\textsuperscript{2}s across the board are always less than 4.5\%. In
the second period we see a dramatic increase in R\textsuperscript{2} for
the market factor built with average CHP factor picks and average term
structure factor picks (4.48\% to 25.1\% and 1.3\% to 20.6\%
respectively) showing dramatically higher co-movement between UK traded
metals and the corresponding picks all of which are US traded
commodities. The increase in R\textsuperscript{2} for the market factor
built with market factor picks is much more modest, providing further
evidence of concentration of co-movement in the metal sector during
financialization.

In order to understand this better, in table \ref{table9}, we examine
the pricing performance of both long and short legs of the corresponding
factors separately on the same set of UK traded metals. The most
striking result here is the dramatic increase in R\textsuperscript{2}
for the long leg of the term structure factor built with CHP factor
picks which achieves the highest overall R\textsuperscript{2} in the
second period (1.91\% vs.~27.37\%). This provides further evidence that
global metals co-movement could be detected by US based factor models
that combined the Keynesian and Working's notions of backwardation.

We then turn our analysis to the crisis and post-crisis period
(2008-2013). In table \ref{table1} we see that the returns on both US
and non-US commodities fall dramatically over this period, going down to
-4.3\% for the UK metals and 8\% for the US traded commodities although
volatility for both sets of assets is similar to that during the second
period. Two of the UK metals have negative mean returns over the third
period compared to one over the first and none over the second. This
trend continues in from 2014 to 2017 with the US traded commodities
recording an average mean return of -0.6\% and the UK metals an average
of 4.9\% over this period, with fourteen US traded commodities and one
UK metal having negative mean returns. The volatility of both sets of
assets is lower than over the 2004-2007 period, particularly for the UK
metals, suggesting the impact of the investment outflows from the
commodity futures markets has had a much lower impact on volatility than
the corresponding inflows witnessed over the financialization period.
The market factor dominates pricing performance over this period with an
increase of mean average R\textsuperscript{2} of over 50\% while that
for the other factors increases slightly at best (table \ref{table3}).
From table \ref{table5} we see that there is an across the board
increase in pairwise correlations with a doubling in the average
pairwise correlation of all US traded commodities and a levelling in the
pairwise correlation of the picks of the three long-short factors. Taken
together, these results indicate the presence of a systematic factor
across the entire cross-section of the US commodities. This is clearly
evident in the performance of the market factor built with any set of
picks on the entire set of US traded commodities (table \ref{table7}) as
we see a dramatic rise across the board in pricing performance. Table
\ref{table8} provides further evidence of systematic linkages, in this
case between US and non-US traded commodities as the market factor built
with its own picks (US traded commodities) achieves an average
R\textsuperscript{2} of 33.7\% on the UK metals compared to 9.9\% in the
previous period.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In the early 2000s, against a backdrop of a low yield environment and
poor stock market performance, a combination of financial
innovation\footnote{Tradable commodity price indexes (commonly referred
  to as commodity index funds or CIFs), and commodity-based exchange
  traded funds (ETFs) were the main financial innovations which allowed
  the large global banks to offer commodity investment products to
  institutional and retail investors. In 1991 Goldman Sachs created the
  S\&P-GSCI which provides investors with buy-side exposure to
  commodities via the OTC swap market and thus without having to
  participate in the formal futures markets with their position limit
  restrictions. At this stage these restrictions still applied to the
  bank though, as it hedged its commodity swap exposure in the futures
  markets. The first commodity-based ETFs were created through buying
  physical precious metals with gold and silver ETFs offered as early as
  2002/2003. The regulatory hurdle here related to the licensing of
  commodity trading professionals. Typically, investors had to sign a
  statement with their broker stating that they understood the risks of
  commodity investments; a rather inconvenient paperwork for a product
  designed to trade like a stock, as set forth by a number of industry
  players at the time.} and deregulation\footnote{In 2000, the Commodity
  Futures Modernization Act (CFMA) granted non-agricultural commodity
  futures statutory exemption from regulation (``Enron loophole'') while
  it required that agricultural commodity derivatives be traded on a
  CFTC-regulated exchange. Notwithstanding, the CFTC classified swap
  dealers as ``bona fide'' hedgers, granting them position limits
  exemption (``swap dealer loophole'').\\
  In 2005, the CFTC waived the rule that required commodity investors to
  sign a statement saying they understood the risks, letting the funds
  replace it with their prospectus and thus unleashed commodity-based
  ETFs investments.} led to a large inflow of institutional and
financial capital into the commodity futures markets. This process known
as financialization lead to a heated public policy debate about whether
the increase in open interest and trading volume in commodities exerted
upward pressure on prices. Perhaps in response to this, most of the
early studies focused on the more mechanical effects of
financialization, relying on commodity specific causality and
correlation based analysis. With the benefit of hindsight this focus
appears to have been too narrow and it seems important to analyze the
effects of financialization from a broader perspective which we do in
this study. We consider the entire cross-section of actively traded
commodities on futures markets and use a futures based asset pricing
framework that includes factors constructed using both price and
non-price attributes. Factor model techniques are well suited to
isolating common driving factors and detecting co-movement, a central
issue in financialization which we believe the existing literature has
not fully analyzed. Our analysis suggests that financialization was a
phenomenon endogenous to the commodity markets transmitted via the
futures markets and appears to have had the deepest impact on global
metal futures markets. The onset of the financial crisis and the
monetary policy regimes that followed appear to have also induced
significant changes in commodity pricing dynamics and our analysis
suggest that in contrast with financialization, the crisis and its
aftermath have delivered an exogenous shock across the entire
cross-section of global commodities.

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`mean returns`}\label{table1} This table shows descriptive statistics for nearby futures returns on various commodity groups as well as for mimicking portfolios for risk factors where the factor portfolios are constructed using the whole cross-section of the US traded commodities considered in the study. Figures labelled *** (**, *) are significant at the 1\% (5\%, 10\%) level. See section \ref{methods} for more details.}\\
\toprule
Assets & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
US-Agriculturals-Grains & Mean return & 3.75\% & ***18.99\% & 5.34\% & -4.14\%\\
US-Agriculturals-Grains & Volatility & 27.14\% & 29.17\% & 33.52\% & 25.18\%\\
US-Agriculturals-Livestock & Mean return & 3.81\% & 5.84\% & *9.53\% & 0.16\%\\
US-Agriculturals-Livestock & Volatility & 26.53\% & 21.59\% & 23.94\% & 26.55\%\\
US-Agriculturals-Softs & Mean return & 1.57\% & **14.26\% & 9.53\% & 2.97\%\\
US-Agriculturals-Softs & Volatility & 33.47\% & 30.19\% & 35.53\% & 28.08\%\\
US-Metals-Industrial & Mean return & 2.42\% & **33.08\% & 6.74\% & 1.35\%\\
US-Metals-Industrial & Volatility & 20.2\% & 30.44\% & 31.66\% & 19.95\%\\
US-Metals-Precious & Mean return & *9.39\% & ***22.7\% & 9.99\% & 2.13\%\\
US-Metals-Precious & Volatility & 26.2\% & 25.61\% & 31.76\% & 21.57\%\\
US-Agriculturals & Mean return & 2.89\% & ***14.45\% & **7.86\% & -0.44\%\\
US-Agriculturals & Volatility & 29.72\% & 28.24\% & 32.69\% & 26.64\%\\
US-Energy & Mean return & 13.56\% & ***28.33\% & 6.94\% & -4.13\%\\
US-Energy & Volatility & 44.03\% & 42.77\% & 39.5\% & 36.52\%\\
US-Metals & Mean return & *7.99\% & ***24.77\% & 9.34\% & 1.98\%\\
US-Metals & Volatility & 25.12\% & 26.64\% & 31.73\% & 21.24\%\\
US-Commodity & Mean return & **5.73\% & ***18.9\% & ***8.01\% & -0.55\%\\
US-Commodity & Volatility & 31.73\% & 30.84\% & 33.72\% & 27.56\%\\
UK-Commodity & Mean return & 2.95\% & ***23.91\% & 4.25\% & 4.87\%\\
UK-Commodity & Volatility & 20.33\% & 33.43\% & 35.99\% & 22.33\%\\
CHP factor & Mean return & 0.34\% & 1.09\% & -0.15\% & -1.98\%\\
CHP factor & Volatility & 7.78\% & 8.05\% & 8.93\% & 7.45\%\\
Market factor & Mean return & 5.99\% & ***18.81\% & 8.01\% & -0.55\%\\
Market factor & Volatility & 10.2\% & 12.39\% & 17.26\% & 10.3\%\\
OI factor & Mean return & -0.2\% & 2.79\% & -4.28\% & 2.28\%\\
OI factor & Volatility & 8.26\% & 8.16\% & 8.1\% & 6.69\%\\
TS factor & Mean return & **-9.49\% & *-7.15\% & -5.77\% & **-9.02\%\\
TS factor & Volatility & 10.25\% & 8.31\% & 9.91\% & 8.52\%\\
\bottomrule
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`mean returns - aggregate CHP regimes`}\label{table2} This table shows descriptive statistics for nearby futures returns on various commodity groups as well as for mimicking portfolios for risk factors where the factor portfolios are constructed using the whole cross-section of the US traded commodities considered in the study. For each period the results are shown for \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. Figures labelled *** (**, *) are significant at the 1\% (5\%, 10\%) level. See section \ref{methods} for more details.}\\
\toprule
Assets & CHP regime & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Assets & CHP regime & Variable & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
US-Agriculturals-Grains & High & Mean return & 4.25\% & ***29.78\% & -2.92\% & -0.55\%\\
US-Agriculturals-Grains & High & Volatility & 27.61\% & 30.17\% & 36.06\% & 24.93\%\\
US-Agriculturals-Grains & Low & Mean return & 3.24\% & 8.26\% & *13.66\% & -7.76\%\\
US-Agriculturals-Grains & Low & Volatility & 26.67\% & 28.09\% & 30.73\% & 25.43\%\\
US-Agriculturals-Livestock & High & Mean return & -2.15\% & -1.68\% & 4.05\% & -15.8\%\\
US-Agriculturals-Livestock & High & Volatility & 29.3\% & 20.65\% & 23.59\% & 25.46\%\\
US-Agriculturals-Livestock & Low & Mean return & 9.84\% & 13.43\% & *15.04\% & 16.28\%\\
US-Agriculturals-Livestock & Low & Volatility & 23.38\% & 22.47\% & 24.29\% & 27.48\%\\
US-Agriculturals-Softs & High & Mean return & -1.31\% & ***24.85\% & 11.43\% & -3.34\%\\
US-Agriculturals-Softs & High & Volatility & 32.54\% & 30.48\% & 36.06\% & 28.72\%\\
US-Agriculturals-Softs & Low & Mean return & 4.47\% & 3.67\% & 7.62\% & 9.33\%\\
US-Agriculturals-Softs & Low & Volatility & 34.4\% & 29.84\% & 35\% & 27.42\%\\
US-Metals-Industrial & High & Mean return & -0.45\% & **51.48\% & 1.3\% & -13.36\%\\
US-Metals-Industrial & High & Volatility & 20.55\% & 31.58\% & 36.97\% & 18.91\%\\
US-Metals-Industrial & Low & Mean return & 5.3\% & 14.86\% & 12.22\% & 16.2\%\\
US-Metals-Industrial & Low & Volatility & 19.89\% & 29.2\% & 25.31\% & 20.85\%\\
US-Metals-Precious & High & Mean return & -3.38\% & ***37.17\% & -2.16\% & -3.58\%\\
US-Metals-Precious & High & Volatility & 28.72\% & 23.7\% & 32.66\% & 22.64\%\\
US-Metals-Precious & Low & Mean return & ***22.26\% & 8.36\% & **22.23\% & 7.9\%\\
US-Metals-Precious & Low & Volatility & 23.28\% & 27.26\% & 30.77\% & 20.42\%\\
US-Agriculturals & High & Mean return & 0.74\% & ***21.45\% & 4.21\% & -4.72\%\\
US-Agriculturals & High & Volatility & 30\% & 28.67\% & 33.93\% & 26.61\%\\
US-Agriculturals & Low & Mean return & 5.06\% & 7.46\% & **11.52\% & 3.88\%\\
US-Agriculturals & Low & Volatility & 29.44\% & 27.79\% & 31.39\% & 26.67\%\\
US-Energy & High & Mean return & 3.9\% & *29.75\% & -3.1\% & -23.15\%\\
US-Energy & High & Volatility & 43.76\% & 45.83\% & 45.46\% & 40.32\%\\
US-Energy & Low & Mean return & *23.3\% & *26.93\% & *17.04\% & 15.07\%\\
US-Energy & Low & Volatility & 44.3\% & 39.55\% & 32.39\% & 32.06\%\\
US-Metals & High & Mean return & -2.79\% & ***40.04\% & -1.47\% & -5.54\%\\
US-Metals & High & Volatility & 27.27\% & 25.45\% & 33.54\% & 21.93\%\\
US-Metals & Low & Mean return & ***18.86\% & 9.66\% & ***20.22\% & 9.56\%\\
US-Metals & Low & Volatility & 22.65\% & 27.63\% & 29.75\% & 20.49\%\\
US-Commodity & High & Mean return & 0.53\% & ***26.68\% & 1.81\% & *-7.96\%\\
US-Commodity & High & Volatility & 32.19\% & 31.59\% & 36.02\% & 28.52\%\\
US-Commodity & Low & Mean return & ***10.98\% & **11.16\% & ***14.25\% & *6.93\%\\
US-Commodity & Low & Volatility & 31.25\% & 30.04\% & 31.22\% & 26.52\%\\
UK-Commodity & High & Mean return & -1.71\% & ***41.22\% & -1.32\% & **-13.66\%\\
UK-Commodity & High & Volatility & 21.05\% & 33.54\% & 38.82\% & 21.69\%\\
UK-Commodity & Low & Mean return & *8.35\% & 6.43\% & 10.04\% & ***23.4\%\\
UK-Commodity & Low & Volatility & 19.45\% & 33.17\% & 32.8\% & 22.69\%\\
CHP factor & High & Mean return & 1.45\% & 7.88\% & 0.89\% & -3.48\%\\
CHP factor & High & Volatility & 8.08\% & 8.28\% & 9.81\% & 7.11\%\\
CHP factor & Low & Mean return & -0.98\% & -5.63\% & -1.2\% & -0.47\%\\
CHP factor & Low & Volatility & 7.43\% & 7.75\% & 7.97\% & 7.81\%\\
Market factor & High & Mean return & 0.67\% & ***26.46\% & 1.81\% & -7.96\%\\
Market factor & High & Volatility & 10.27\% & 11.96\% & 18.97\% & 11.24\%\\
Market factor & Low & Mean return & **11.43\% & 11.16\% & 14.25\% & 6.93\%\\
Market factor & Low & Volatility & 10.1\% & 12.78\% & 15.37\% & 9.19\%\\
OI factor & High & Mean return & -0.76\% & 4.89\% & -6.9\% & 3.84\%\\
OI factor & High & Volatility & 7.92\% & 8.79\% & 8.64\% & 6.7\%\\
OI factor & Low & Mean return & 0.46\% & 0.72\% & -1.65\% & 0.7\%\\
OI factor & Low & Volatility & 8.66\% & 7.52\% & 7.53\% & 6.7\%\\
TS factor & High & Mean return & **-13.59\% & -7.38\% & -8.14\% & *-11.23\%\\
TS factor & High & Volatility & 10.38\% & 8.71\% & 10.76\% & 8.39\%\\
TS factor & Low & Mean return & -4.6\% & -6.92\% & -3.37\% & -6.79\%\\
TS factor & Low & Volatility & 10.08\% & 7.94\% & 9\% & 8.69\%\\*
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor regressions`}\label{table3} This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby futures returns on the commodities forming the entire cross-section of US traded commodities considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using whole set of US traded commodities considered. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. See section \ref{methods} for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
US commodities & \textasciitilde{} CHP factor & High & 1.54\% & 5.5\% & 5.89\% & 3.73\%\\
US commodities & \textasciitilde{} CHP factor & Low & 1.39\% & 5.64\% & 7.33\% & 5.43\%\\
US commodities & \textasciitilde{} CHP factor & None & 1.37\% & 5.43\% & 5.99\% & 4.47\%\\
US commodities & \textasciitilde{} Market factor & High & 13.04\% & 16.96\% & 30.45\% & 17.62\%\\
US commodities & \textasciitilde{} Market factor & Low & 11.61\% & 19.83\% & 28.92\% & 13.53\%\\
US commodities & \textasciitilde{} Market factor & None & 12.2\% & 18.55\% & 29.45\% & 15.57\%\\
US commodities & \textasciitilde{} OI factor & High & 0.82\% & 1.32\% & 1.19\% & 0.31\%\\
US commodities & \textasciitilde{} OI factor & Low & 1.42\% & -0.39\% & 0.54\% & 0.56\%\\
US commodities & \textasciitilde{} OI factor & None & 0.66\% & 0.36\% & 0.31\% & 0.31\%\\
US commodities & \textasciitilde{} TS factor & High & 2.12\% & 3.36\% & 4.68\% & 2.69\%\\
US commodities & \textasciitilde{} TS factor & Low & 2.88\% & 2.72\% & 5.57\% & 3.09\%\\
US commodities & \textasciitilde{} TS factor & None & 2.4\% & 2.7\% & 4.69\% & 2.52\%\\
\bottomrule
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright\arraybackslash}p{3.5cm}>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor picks`}\label{table4} For each risk factor and period, this table displays the list of corresponding commodity factor picks along with the proportion of time they are individually held on the long and short legs of the corresponding risk factor mimicking portfolio. The factor portfolios are constructed from the whole set of the twenty four US traded commodities considered which is also used as asset pool for the commodity picking exercise. See section \ref{methods} for more details.}\\
\toprule
Factor & Period & Asset & Long & Short\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor & Period & Asset & Long & Short\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
CHP & 1997/2003 & XCBT - Oats & 97\% & 0\%\\
CHP & 1997/2003 & XCBT - Soy-oil & 42\% & 32\%\\
CHP & 1997/2003 & XCBT - Wheat-SRW & 29\% & 9\%\\
CHP & 1997/2003 & XCEC - Copper & 43\% & 18\%\\
CHP & 1997/2003 & XCEC - Silver & 100\% & 0\%\\
CHP & 1997/2003 & XNYM - Platinum & 91\% & 4\%\\
CHP & 1997/2003 & IFUS - Coffee & 22\% & 17\%\\
CHP & 1997/2003 & IFUS - Orange juice & 67\% & 18\%\\
Market & 1997/2003 & XCBT - Corn & 100\% & 0\%\\
Market & 1997/2003 & XCBT - Oats & 100\% & 0\%\\
Market & 1997/2003 & XCBT - Soy-meal & 100\% & 0\%\\
Market & 1997/2003 & XCBT - Wheat-SRW & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Crude oil & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Gasoline & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Heating oil & 100\% & 0\%\\
Market & 1997/2003 & XNYM - Natural gas & 99\% & 0\%\\
OI & 1997/2003 & XCME - Feeder cattle & 27\% & 39\%\\
OI & 1997/2003 & XCME - Lean hogs & 36\% & 30\%\\
OI & 1997/2003 & XNYM - Crude oil & 24\% & 28\%\\
OI & 1997/2003 & XNYM - Natural gas & 46\% & 24\%\\
OI & 1997/2003 & XCEC - Gold & 39\% & 28\%\\
OI & 1997/2003 & XNYM - Palladium & 46\% & 33\%\\
OI & 1997/2003 & IFUS - Coffee & 39\% & 23\%\\
OI & 1997/2003 & XCME - Lumber & 42\% & 33\%\\
TS & 1997/2003 & XCBT - Soy-meal & 71\% & 0\%\\
TS & 1997/2003 & XNYM - Crude oil & 53\% & 9\%\\
TS & 1997/2003 & XNYM - Gasoline & 61\% & 17\%\\
TS & 1997/2003 & XNYM - Heating oil & 35\% & 24\%\\
TS & 1997/2003 & XCEC - Gold & 10\% & 1\%\\
TS & 1997/2003 & XNYM - Palladium & 61\% & 0\%\\
TS & 1997/2003 & XNYM - Platinum & 96\% & 0\%\\
TS & 1997/2003 & IFUS - Sugar & 60\% & 26\%\\
CHP & 2004/2007 & XCBT - Oats & 100\% & 0\%\\
CHP & 2004/2007 & XCEC - Copper & 44\% & 37\%\\
CHP & 2004/2007 & XCEC - Gold & 100\% & 0\%\\
CHP & 2004/2007 & XCEC - Silver & 100\% & 0\%\\
CHP & 2004/2007 & XNYM - Palladium & 100\% & 0\%\\
CHP & 2004/2007 & XNYM - Platinum & 100\% & 0\%\\
CHP & 2004/2007 & IFUS - Coffee & 15\% & 5\%\\
CHP & 2004/2007 & IFUS - Orange juice & 66\% & 15\%\\
Market & 2004/2007 & XCBT - Corn & 100\% & 0\%\\
Market & 2004/2007 & XCBT - Soybeans & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Heating oil & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Natural gas & 100\% & 0\%\\
Market & 2004/2007 & XCEC - Silver & 100\% & 0\%\\
Market & 2004/2007 & XNYM - Palladium & 100\% & 0\%\\
OI & 2004/2007 & XCME - Lean hogs & 38\% & 22\%\\
OI & 2004/2007 & XCME - Live cattle & 32\% & 35\%\\
OI & 2004/2007 & XNYM - Natural gas & 32\% & 35\%\\
OI & 2004/2007 & XNYM - Palladium & 43\% & 28\%\\
OI & 2004/2007 & IFUS - Cocoa & 18\% & 36\%\\
OI & 2004/2007 & IFUS - Coffee & 34\% & 28\%\\
OI & 2004/2007 & IFUS - Cotton & 33\% & 17\%\\
OI & 2004/2007 & IFUS - Sugar & 48\% & 16\%\\
TS & 2004/2007 & XCEC - Copper & 86\% & 0\%\\
TS & 2004/2007 & XCME - Feeder cattle & 86\% & 0\%\\
TS & 2004/2007 & XCME - Lean hogs & 44\% & 46\%\\
TS & 2004/2007 & XCME - Live cattle & 56\% & 23\%\\
TS & 2004/2007 & XCEC - Gold & 28\% & 0\%\\
TS & 2004/2007 & XNYM - Palladium & 0\% & 6\%\\
TS & 2004/2007 & XNYM - Platinum & 95\% & 0\%\\
TS & 2004/2007 & IFUS - Cocoa & 15\% & 12\%\\
CHP & 2008/2013 & XCBT - Soy-meal & 79\% & 4\%\\
CHP & 2008/2013 & XNYM - Gasoline & 46\% & 0\%\\
CHP & 2008/2013 & XNYM - Heating oil & 1\% & 18\%\\
CHP & 2008/2013 & XCEC - Gold & 96\% & 0\%\\
CHP & 2008/2013 & XCEC - Silver & 90\% & 0\%\\
CHP & 2008/2013 & XNYM - Palladium & 100\% & 0\%\\
CHP & 2008/2013 & XNYM - Platinum & 100\% & 0\%\\
CHP & 2008/2013 & IFUS - Orange juice & 96\% & 0\%\\
Market & 2008/2013 & XCBT - Corn & 100\% & 0\%\\
Market & 2008/2013 & XCBT - Oats & 100\% & 0\%\\
Market & 2008/2013 & XCEC - Copper & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Heating oil & 100\% & 0\%\\
Market & 2008/2013 & XCEC - Silver & 100\% & 0\%\\
Market & 2008/2013 & XNYM - Palladium & 100\% & 0\%\\
OI & 2008/2013 & XCBT - Corn & 21\% & 37\%\\
OI & 2008/2013 & XCBT - Soy-meal & 37\% & 34\%\\
OI & 2008/2013 & XNYM - Crude oil & 23\% & 31\%\\
OI & 2008/2013 & XNYM - Gasoline & 33\% & 27\%\\
OI & 2008/2013 & XNYM - Heating oil & 27\% & 34\%\\
OI & 2008/2013 & IFUS - Orange juice & 32\% & 43\%\\
OI & 2008/2013 & IFUS - Sugar & 36\% & 31\%\\
OI & 2008/2013 & XCME - Lumber & 35\% & 38\%\\
TS & 2008/2013 & XCBT - Soy-meal & 69\% & 1\%\\
TS & 2008/2013 & XCBT - Soybeans & 59\% & 0\%\\
TS & 2008/2013 & XNYM - Crude oil & 25\% & 30\%\\
TS & 2008/2013 & XNYM - Gasoline & 61\% & 6\%\\
TS & 2008/2013 & XNYM - Heating oil & 33\% & 6\%\\
TS & 2008/2013 & XCEC - Silver & 64\% & 0\%\\
TS & 2008/2013 & XNYM - Palladium & 38\% & 0\%\\
TS & 2008/2013 & XNYM - Platinum & 47\% & 1\%\\
CHP & 2014/2017 & XCEC - Copper & 0\% & 79\%\\
CHP & 2014/2017 & XNYM - Gasoline & 13\% & 0\%\\
CHP & 2014/2017 & XCEC - Gold & 78\% & 0\%\\
CHP & 2014/2017 & XCEC - Silver & 69\% & 0\%\\
CHP & 2014/2017 & XNYM - Palladium & 100\% & 0\%\\
CHP & 2014/2017 & XNYM - Platinum & 100\% & 0\%\\
CHP & 2014/2017 & IFUS - Orange juice & 62\% & 2\%\\
CHP & 2014/2017 & XCME - Lumber & 38\% & 53\%\\
Market & 2014/2017 & XCBT - Soy-meal & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Crude oil & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Gasoline & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Heating oil & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Natural gas & 100\% & 0\%\\
Market & 2014/2017 & XNYM - Palladium & 100\% & 0\%\\
Market & 2014/2017 & IFUS - Coffee & 100\% & 0\%\\
Market & 2014/2017 & IFUS - Sugar & 100\% & 0\%\\
OI & 2014/2017 & XCBT - Oats & 32\% & 45\%\\
OI & 2014/2017 & XCBT - Soy-meal & 39\% & 27\%\\
OI & 2014/2017 & XCEC - Copper & 48\% & 15\%\\
OI & 2014/2017 & XCEC - Gold & 39\% & 23\%\\
OI & 2014/2017 & XCEC - Silver & 44\% & 28\%\\
OI & 2014/2017 & IFUS - Cocoa & 33\% & 34\%\\
OI & 2014/2017 & IFUS - Cotton & 42\% & 31\%\\
OI & 2014/2017 & XCME - Lumber & 50\% & 30\%\\
TS & 2014/2017 & XCBT - Soy-meal & 68\% & 10\%\\
TS & 2014/2017 & XCBT - Soybeans & 56\% & 10\%\\
TS & 2014/2017 & XCME - Feeder cattle & 75\% & 10\%\\
TS & 2014/2017 & XCME - Live cattle & 65\% & 24\%\\
TS & 2014/2017 & IFUS - Cocoa & 54\% & 1\%\\
TS & 2014/2017 & IFUS - Cotton & 40\% & 2\%\\
TS & 2014/2017 & IFUS - Orange juice & 41\% & 23\%\\
TS & 2014/2017 & XCME - Lumber & 41\% & 39\%\\*
\end{longtabu}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:correlations}\label{table5} This table displays the average pairwise nearby futures returns correlations for various sets of commodities including the multiple sets of risk factor commodity picks, the whole cross-section of US traded commodities considered in the study, the latter with metals omitted, and the cross-section of US traded metals. For each period and commodity set the correlations are calculated independently over the whole period as well as over \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes. See section \ref{methods} for more details.}\\
\toprule
Asset pool & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
CHP factor picks & High & 0.09 & 0.24 & 0.32 & 0.22\\
CHP factor picks & Low & 0.10 & 0.28 & 0.37 & 0.19\\
CHP factor picks & None & 0.10 & 0.27 & 0.34 & 0.21\\
Market factor picks & High & 0.23 & 0.24 & 0.40 & 0.22\\
Market factor picks & Low & 0.20 & 0.28 & 0.42 & 0.19\\
Market factor picks & None & 0.21 & 0.26 & 0.41 & 0.21\\
OI factor picks & High & 0.03 & 0.04 & 0.28 & 0.15\\
OI factor picks & Low & 0.03 & 0.09 & 0.22 & 0.07\\
OI factor picks & None & 0.03 & 0.07 & 0.25 & 0.11\\
TS factor picks & High & 0.16 & 0.15 & 0.41 & 0.11\\
TS factor picks & Low & 0.11 & 0.17 & 0.48 & 0.09\\
TS factor picks & None & 0.13 & 0.16 & 0.44 & 0.10\\
US - Commodity & High & 0.08 & 0.11 & 0.25 & 0.13\\
US - Commodity & Low & 0.07 & 0.15 & 0.22 & 0.09\\
US - Commodity & None & 0.08 & 0.13 & 0.24 & 0.11\\
US - Commodity - No metals & High & 0.09 & 0.10 & 0.26 & 0.12\\
US - Commodity - No metals & Low & 0.08 & 0.13 & 0.17 & 0.10\\
US - Commodity - No metals & None & 0.09 & 0.12 & 0.22 & 0.11\\
US - Metals & High & 0.23 & 0.55 & 0.52 & 0.44\\
US - Metals & Low & 0.19 & 0.52 & 0.61 & 0.42\\
US - Metals & None & 0.21 & 0.53 & 0.55 & 0.43\\
\bottomrule
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor pick regressions - picks:picks`}\label{table6} This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby futures returns on risk factors commodity picks while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in table \ref{table4}. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. See section \ref{methods} for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & 1.45\% & 20.37\% & 16.95\% & 8.81\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & 7.6\% & 21.82\% & 10.29\% & 4.84\%\\
TS factor picks & \textasciitilde{} CHP factor & None & 3.42\% & 21.2\% & 14.09\% & 5.06\%\\
TS factor picks & \textasciitilde{} Market factor & High & 28.64\% & 27.15\% & 48.8\% & 21.54\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 25.53\% & 28.87\% & 54.4\% & 21.22\%\\
TS factor picks & \textasciitilde{} Market factor & None & 27.23\% & 28.26\% & 50.95\% & 21.63\%\\
TS factor picks & \textasciitilde{} OI factor & High & 0.07\% & 6.42\% & 4.08\% & 2.13\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 2.48\% & 0.81\% & 2.68\% & 0.61\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.68\% & 2.26\% & 1.81\% & 0.56\%\\
TS factor picks & \textasciitilde{} TS factor & High & 2.04\% & 7.65\% & 5.2\% & 2.98\%\\
TS factor picks & \textasciitilde{} TS factor & Low & 2.22\% & 6.03\% & 2\% & 1.85\%\\
TS factor picks & \textasciitilde{} TS factor & None & 0.53\% & 6.91\% & 4.16\% & 2.02\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 3.02\% & 5.9\% & 4.56\% & 10.95\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & 1.32\% & 9.37\% & 7.5\% & 5.5\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 1.96\% & 6.96\% & 3.95\% & 6.36\%\\
OI factor picks & \textasciitilde{} Market factor & High & 14.4\% & 15.55\% & 39.51\% & 24.98\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 14.99\% & 20.4\% & 32.69\% & 17.31\%\\
OI factor picks & \textasciitilde{} Market factor & None & 14.86\% & 18\% & 36.91\% & 21.08\%\\
OI factor picks & \textasciitilde{} OI factor & High & 1.67\% & 1.82\% & 0.87\% & 0.64\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 0.67\% & -0.38\% & 0.95\% & 3.52\%\\
OI factor picks & \textasciitilde{} OI factor & None & 0.29\% & 0.77\% & 0.09\% & 0.74\%\\
OI factor picks & \textasciitilde{} TS factor & High & 3.24\% & 7.15\% & 13.59\% & 3.69\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 2.92\% & 2.7\% & 4.11\% & 3\%\\
OI factor picks & \textasciitilde{} TS factor & None & 3.07\% & 5.36\% & 9.97\% & 3.08\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 5.74\% & 15.05\% & 12.05\% & 12.46\%\\
CHP factor picks & \textasciitilde{} CHP factor & Low & 7.05\% & 10.85\% & 8.7\% & 14.6\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 5.83\% & 13.23\% & 10.46\% & 12.91\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 21.09\% & 35.23\% & 41.13\% & 32.66\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 21.6\% & 38.7\% & 46.97\% & 29.72\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 21.27\% & 37.36\% & 43.49\% & 31.36\%\\
CHP factor picks & \textasciitilde{} OI factor & High & 1.69\% & 3.9\% & 4.04\% & 0.74\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 3.51\% & 1.61\% & 1.03\% & 2.96\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 1.57\% & 1.64\% & 2.07\% & 0.23\%\\
CHP factor picks & \textasciitilde{} TS factor & High & 4.77\% & 7.1\% & 1.98\% & 4.78\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & 5.36\% & 5.57\% & 0.95\% & -0.21\%\\
CHP factor picks & \textasciitilde{} TS factor & None & 4.54\% & 5.1\% & 1.84\% & 1.66\%\\
Market factor picks & \textasciitilde{} CHP factor & High & 7.79\% & 8.62\% & 7.1\% & 7.84\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & 4.6\% & 7.78\% & 7.63\% & 7.73\%\\
Market factor picks & \textasciitilde{} CHP factor & None & 6.06\% & 8.31\% & 6.22\% & 7.07\%\\
Market factor picks & \textasciitilde{} Market factor & High & 32.85\% & 34.83\% & 48.14\% & 33.71\%\\
Market factor picks & \textasciitilde{} Market factor & Low & 30.4\% & 37.35\% & 49.31\% & 29.36\%\\
Market factor picks & \textasciitilde{} Market factor & None & 31.43\% & 36.1\% & 48.63\% & 32.31\%\\
Market factor picks & \textasciitilde{} OI factor & High & 3.74\% & 0.77\% & 2.62\% & 0.52\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.41\% & -0.45\% & 1.22\% & -0.05\%\\
Market factor picks & \textasciitilde{} OI factor & None & 1.33\% & 0.29\% & 0.78\% & 0.53\%\\
Market factor picks & \textasciitilde{} TS factor & High & 6.64\% & 5.3\% & 2.45\% & 7.59\%\\
Market factor picks & \textasciitilde{} TS factor & Low & 8.53\% & 4.21\% & 11.16\% & 4.7\%\\
Market factor picks & \textasciitilde{} TS factor & None & 7.65\% & 5.2\% & 3.25\% & 6.24\%\\*
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor pick regressions - picks: US commos`}\label{table7} This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby futures returns on the commodities forming the entire cross-section of US traded commodities considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in table \ref{table4}. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. See section \ref{methods} for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & 1.45\% & 20.37\% & 16.95\% & 8.81\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & 7.6\% & 21.82\% & 10.29\% & 4.84\%\\
TS factor picks & \textasciitilde{} CHP factor & None & 3.42\% & 21.2\% & 14.09\% & 5.06\%\\
TS factor picks & \textasciitilde{} Market factor & High & 28.64\% & 27.15\% & 48.8\% & 21.54\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 25.53\% & 28.87\% & 54.4\% & 21.22\%\\
TS factor picks & \textasciitilde{} Market factor & None & 27.23\% & 28.26\% & 50.95\% & 21.63\%\\
TS factor picks & \textasciitilde{} OI factor & High & 0.07\% & 6.42\% & 4.08\% & 2.13\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 2.48\% & 0.81\% & 2.68\% & 0.61\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.68\% & 2.26\% & 1.81\% & 0.56\%\\
TS factor picks & \textasciitilde{} TS factor & High & 2.04\% & 7.65\% & 5.2\% & 2.98\%\\
TS factor picks & \textasciitilde{} TS factor & Low & 2.22\% & 6.03\% & 2\% & 1.85\%\\
TS factor picks & \textasciitilde{} TS factor & None & 0.53\% & 6.91\% & 4.16\% & 2.02\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 3.02\% & 5.9\% & 4.56\% & 10.95\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & 1.32\% & 9.37\% & 7.5\% & 5.5\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 1.96\% & 6.96\% & 3.95\% & 6.36\%\\
OI factor picks & \textasciitilde{} Market factor & High & 14.4\% & 15.55\% & 39.51\% & 24.98\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 14.99\% & 20.4\% & 32.69\% & 17.31\%\\
OI factor picks & \textasciitilde{} Market factor & None & 14.86\% & 18\% & 36.91\% & 21.08\%\\
OI factor picks & \textasciitilde{} OI factor & High & 1.67\% & 1.82\% & 0.87\% & 0.64\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 0.67\% & -0.38\% & 0.95\% & 3.52\%\\
OI factor picks & \textasciitilde{} OI factor & None & 0.29\% & 0.77\% & 0.09\% & 0.74\%\\
OI factor picks & \textasciitilde{} TS factor & High & 3.24\% & 7.15\% & 13.59\% & 3.69\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 2.92\% & 2.7\% & 4.11\% & 3\%\\
OI factor picks & \textasciitilde{} TS factor & None & 3.07\% & 5.36\% & 9.97\% & 3.08\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 5.74\% & 15.05\% & 12.05\% & 12.46\%\\
CHP factor picks & \textasciitilde{} CHP factor & Low & 7.05\% & 10.85\% & 8.7\% & 14.6\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 5.83\% & 13.23\% & 10.46\% & 12.91\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 21.09\% & 35.23\% & 41.13\% & 32.66\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 21.6\% & 38.7\% & 46.97\% & 29.72\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 21.27\% & 37.36\% & 43.49\% & 31.36\%\\
CHP factor picks & \textasciitilde{} OI factor & High & 1.69\% & 3.9\% & 4.04\% & 0.74\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 3.51\% & 1.61\% & 1.03\% & 2.96\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 1.57\% & 1.64\% & 2.07\% & 0.23\%\\
CHP factor picks & \textasciitilde{} TS factor & High & 4.77\% & 7.1\% & 1.98\% & 4.78\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & 5.36\% & 5.57\% & 0.95\% & -0.21\%\\
CHP factor picks & \textasciitilde{} TS factor & None & 4.54\% & 5.1\% & 1.84\% & 1.66\%\\
Market factor picks & \textasciitilde{} CHP factor & High & 7.79\% & 8.62\% & 7.1\% & 7.84\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & 4.6\% & 7.78\% & 7.63\% & 7.73\%\\
Market factor picks & \textasciitilde{} CHP factor & None & 6.06\% & 8.31\% & 6.22\% & 7.07\%\\
Market factor picks & \textasciitilde{} Market factor & High & 32.85\% & 34.83\% & 48.14\% & 33.71\%\\
Market factor picks & \textasciitilde{} Market factor & Low & 30.4\% & 37.35\% & 49.31\% & 29.36\%\\
Market factor picks & \textasciitilde{} Market factor & None & 31.43\% & 36.1\% & 48.63\% & 32.31\%\\
Market factor picks & \textasciitilde{} OI factor & High & 3.74\% & 0.77\% & 2.62\% & 0.52\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.41\% & -0.45\% & 1.22\% & -0.05\%\\
Market factor picks & \textasciitilde{} OI factor & None & 1.33\% & 0.29\% & 0.78\% & 0.53\%\\
Market factor picks & \textasciitilde{} TS factor & High & 6.64\% & 5.3\% & 2.45\% & 7.59\%\\
Market factor picks & \textasciitilde{} TS factor & Low & 8.53\% & 4.21\% & 11.16\% & 4.7\%\\
Market factor picks & \textasciitilde{} TS factor & None & 7.65\% & 5.2\% & 3.25\% & 6.24\%\\*
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor pick regressions - picks: UK commos`}\label{table8} This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby futures returns on the six LME traded metals considered in the study while the regressors are returns on mimicking portfolios for risk factors where the factor portfolios are constructed using the corresponding set of risk factor commodity picks shown in Table \ref{table4}. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. See section \ref{methods} for more details.}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & High & -0.01\% & 2.1\% & 3.87\% & 1.05\%\\
TS factor picks & \textasciitilde{} CHP factor & Low & -0.37\% & 17.79\% & 3.32\% & 0.8\%\\
TS factor picks & \textasciitilde{} CHP factor & None & -0.05\% & 7.89\% & 0.65\% & -0.26\%\\
TS factor picks & \textasciitilde{} Market factor & High & 0.65\% & 17.55\% & 30.72\% & 5.69\%\\
TS factor picks & \textasciitilde{} Market factor & Low & 1.62\% & 22.57\% & 27.92\% & 0.68\%\\
TS factor picks & \textasciitilde{} Market factor & None & 1.3\% & 20.58\% & 29.6\% & 3.01\%\\
TS factor picks & \textasciitilde{} OI factor & High & -0.36\% & 4.97\% & 0.31\% & -0.65\%\\
TS factor picks & \textasciitilde{} OI factor & Low & 0.06\% & 1.29\% & -0.33\% & -0.11\%\\
TS factor picks & \textasciitilde{} OI factor & None & 0.04\% & 3.02\% & -0.02\% & -0.18\%\\
TS factor picks & \textasciitilde{} TS factor & High & 0.45\% & 4.72\% & 0.22\% & -0.22\%\\
TS factor picks & \textasciitilde{} TS factor & Low & 0.01\% & -0.67\% & 0.22\% & 1.72\%\\
TS factor picks & \textasciitilde{} TS factor & None & -0.12\% & 2.11\% & 0.44\% & 1.33\%\\
OI factor picks & \textasciitilde{} CHP factor & High & 1.63\% & 0.34\% & -0.37\% & 5.03\%\\
OI factor picks & \textasciitilde{} CHP factor & Low & -0.04\% & 8.57\% & 3.01\% & 0.43\%\\
OI factor picks & \textasciitilde{} CHP factor & None & 0.01\% & 3.54\% & 0.55\% & 0.72\%\\
OI factor picks & \textasciitilde{} Market factor & High & -0.24\% & 4.93\% & 28.78\% & 12.91\%\\
OI factor picks & \textasciitilde{} Market factor & Low & 3.59\% & 10.46\% & 16.76\% & 7.99\%\\
OI factor picks & \textasciitilde{} Market factor & None & 1.17\% & 7.97\% & 24.03\% & 10.37\%\\
OI factor picks & \textasciitilde{} OI factor & High & -0.18\% & -0.68\% & 0.43\% & 0.76\%\\
OI factor picks & \textasciitilde{} OI factor & Low & 1.79\% & 0.78\% & -0.32\% & 3.01\%\\
OI factor picks & \textasciitilde{} OI factor & None & 0.54\% & -0.33\% & 0.07\% & -0.12\%\\
OI factor picks & \textasciitilde{} TS factor & High & -0.26\% & -0.34\% & 7.07\% & -0.26\%\\
OI factor picks & \textasciitilde{} TS factor & Low & 0.3\% & 1.01\% & 0.58\% & -0.31\%\\
OI factor picks & \textasciitilde{} TS factor & None & 0.17\% & 0.11\% & 4.57\% & -0.17\%\\
CHP factor picks & \textasciitilde{} CHP factor & High & 0.48\% & -0.1\% & 2.96\% & 0.46\%\\
CHP factor picks & \textasciitilde{} CHP factor & Low & 0.15\% & -0.55\% & 2.7\% & 0.77\%\\
CHP factor picks & \textasciitilde{} CHP factor & None & 0.05\% & -0.43\% & 0.48\% & 0.05\%\\
CHP factor picks & \textasciitilde{} Market factor & High & 4.22\% & 18.7\% & 26.41\% & 15.02\%\\
CHP factor picks & \textasciitilde{} Market factor & Low & 4.4\% & 30.16\% & 27.34\% & 14.77\%\\
CHP factor picks & \textasciitilde{} Market factor & None & 4.48\% & 25.09\% & 26.83\% & 14.99\%\\
CHP factor picks & \textasciitilde{} OI factor & High & 0.28\% & 6.51\% & 0\% & -0.57\%\\
CHP factor picks & \textasciitilde{} OI factor & Low & 0.41\% & 0.18\% & 0.39\% & 0.65\%\\
CHP factor picks & \textasciitilde{} OI factor & None & 0.49\% & 3.89\% & 0.29\% & -0.1\%\\
CHP factor picks & \textasciitilde{} TS factor & High & -0.36\% & 4.19\% & -0.19\% & 2.09\%\\
CHP factor picks & \textasciitilde{} TS factor & Low & -0.26\% & -0.52\% & -0.17\% & 0.28\%\\
CHP factor picks & \textasciitilde{} TS factor & None & -0.24\% & 2.07\% & 0.13\% & 1.33\%\\
Market factor picks & \textasciitilde{} CHP factor & High & -0.22\% & 3.62\% & 0.64\% & 1.88\%\\
Market factor picks & \textasciitilde{} CHP factor & Low & -0.33\% & 9.62\% & 1.72\% & 1.12\%\\
Market factor picks & \textasciitilde{} CHP factor & None & -0.1\% & 6.32\% & -0.08\% & 1.61\%\\
Market factor picks & \textasciitilde{} Market factor & High & -0.11\% & 8.95\% & 36.58\% & 11.77\%\\
Market factor picks & \textasciitilde{} Market factor & Low & 2.08\% & 9.85\% & 29.63\% & 2.79\%\\
Market factor picks & \textasciitilde{} Market factor & None & 0.88\% & 9.93\% & 33.73\% & 7.69\%\\
Market factor picks & \textasciitilde{} OI factor & High & 0.24\% & -0.58\% & 0.84\% & -0.04\%\\
Market factor picks & \textasciitilde{} OI factor & Low & 0.66\% & 0.56\% & -0.17\% & -0.42\%\\
Market factor picks & \textasciitilde{} OI factor & None & -0.04\% & 0.12\% & 0.62\% & -0.24\%\\
Market factor picks & \textasciitilde{} TS factor & High & -0.38\% & 1.37\% & 2.9\% & -0.37\%\\
Market factor picks & \textasciitilde{} TS factor & Low & -0.43\% & 1.41\% & 0.32\% & 0.39\%\\
Market factor picks & \textasciitilde{} TS factor & None & -0.23\% & 1.65\% & 2.15\% & 0.12\%\\*
\end{longtabu}
\end{landscape}
\endgroup{}

\newpage
\begingroup\fontsize{10}{12}\selectfont

\begin{landscape}
\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedright\arraybackslash}p{2cm}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:`factor pick regressions - picks: UK commos - legs`}\label{table9} This table shows the average time series adjusted $R^{2}$s for a set of commodity factor models. The dependent variables are individual nearby futures returns on the six LME traded metals considered in the study while the regressors are returns on the long and short legs of mimicking portfolios for risk factors considered independently, where the factor portfolios are constructed using the corresponding set of risk factor commodity picks. For each period and model, the corresponding individual commodity adjusted $R^{2}$s are averaged. For each period, the models are implemented over the whole period, as well as over \textbf{contango} (high) and \textbf{backwardation} (low) \textbf{CHP} regimes independently. See section \ref{methods} for more details.}\\
\toprule
Factor asset pool & Model & Leg & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endfirsthead
\caption[]{continued}\\
\toprule
Factor asset pool & Model & Leg & CHP regime & 1997/2003 & 2004/2007 & 2008/2013 & 2014/2017\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
TS factor picks & \textasciitilde{} CHP factor & Long & High & 0.91\% & 11.5\% & 13.22\% & -0.15\%\\
TS factor picks & \textasciitilde{} CHP factor & Long & Low & 1.51\% & 29.68\% & 34.77\% & 1.21\%\\
TS factor picks & \textasciitilde{} CHP factor & Long & None & 1.39\% & 19.71\% & 20.35\% & 1.09\%\\
TS factor picks & \textasciitilde{} CHP factor & Short & High & 0.01\% & 1.47\% & 28.59\% & 4.63\%\\
TS factor picks & \textasciitilde{} CHP factor & Short & Low & -0.15\% & -0.48\% & 26.72\% & -0.11\%\\
TS factor picks & \textasciitilde{} CHP factor & Short & None & 0.16\% & 0.9\% & 27.93\% & 1.23\%\\
TS factor picks & \textasciitilde{} Market factor & Long & High & 0.65\% & 17.55\% & 30.72\% & 5.69\%\\
TS factor picks & \textasciitilde{} Market factor & Long & Low & 1.62\% & 22.57\% & 27.92\% & 0.68\%\\
TS factor picks & \textasciitilde{} Market factor & Long & None & 1.3\% & 20.58\% & 29.6\% & 3.01\%\\
TS factor picks & \textasciitilde{} OI factor & Long & High & 0.36\% & 4.73\% & 23.81\% & 1.72\%\\
TS factor picks & \textasciitilde{} OI factor & Long & Low & 1.54\% & 11.1\% & 30.88\% & -0.57\%\\
TS factor picks & \textasciitilde{} OI factor & Long & None & 0.99\% & 8.22\% & 25.87\% & 0.26\%\\
TS factor picks & \textasciitilde{} OI factor & Short & High & 0.56\% & 22.57\% & 20.33\% & 2.35\%\\
TS factor picks & \textasciitilde{} OI factor & Short & Low & 2.01\% & 20.14\% & 28.09\% & 0.2\%\\
TS factor picks & \textasciitilde{} OI factor & Short & None & 1.42\% & 21.62\% & 22.99\% & 1.8\%\\
TS factor picks & \textasciitilde{} TS factor & Long & High & -0.14\% & 24.07\% & 19.08\% & 0.51\%\\
TS factor picks & \textasciitilde{} TS factor & Long & Low & 1.18\% & 22.59\% & 23.52\% & -0.54\%\\
TS factor picks & \textasciitilde{} TS factor & Long & None & 0.37\% & 23.8\% & 20.46\% & -0.36\%\\
TS factor picks & \textasciitilde{} TS factor & Short & High & 1.14\% & 4.16\% & 25.71\% & 3.89\%\\
TS factor picks & \textasciitilde{} TS factor & Short & Low & 0.6\% & 9.27\% & 34.98\% & 2.59\%\\
TS factor picks & \textasciitilde{} TS factor & Short & None & 1.06\% & 7.24\% & 28.35\% & 4.01\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & High & 0.19\% & 8.36\% & 15.6\% & 1.65\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & Low & 3.42\% & 16.64\% & 8.36\% & 6.14\%\\
OI factor picks & \textasciitilde{} CHP factor & Long & None & 0.03\% & 12.55\% & 13.29\% & 3.64\%\\
OI factor picks & \textasciitilde{} CHP factor & Short & High & 0.9\% & 0.79\% & 23.17\% & 13.29\%\\
OI factor picks & \textasciitilde{} CHP factor & Short & Low & 0.45\% & 0.65\% & 20.93\% & 3.06\%\\
OI factor picks & \textasciitilde{} CHP factor & Short & None & 1.01\% & 1.03\% & 22.48\% & 8.42\%\\
OI factor picks & \textasciitilde{} Market factor & Long & High & -0.24\% & 4.93\% & 28.78\% & 12.91\%\\
OI factor picks & \textasciitilde{} Market factor & Long & Low & 3.59\% & 10.46\% & 16.76\% & 7.99\%\\
OI factor picks & \textasciitilde{} Market factor & Long & None & 1.17\% & 7.97\% & 24.03\% & 10.37\%\\
OI factor picks & \textasciitilde{} OI factor & Long & High & -0.39\% & 1.54\% & 17.05\% & 3.59\%\\
OI factor picks & \textasciitilde{} OI factor & Long & Low & -0.18\% & 7.12\% & 16.74\% & 6.64\%\\
OI factor picks & \textasciitilde{} OI factor & Long & None & -0.22\% & 3.7\% & 17.08\% & 5.56\%\\
OI factor picks & \textasciitilde{} OI factor & Short & High & 0.01\% & 3.92\% & 24.04\% & 9.07\%\\
OI factor picks & \textasciitilde{} OI factor & Short & Low & 3.83\% & 2.41\% & 15.01\% & -1.06\%\\
OI factor picks & \textasciitilde{} OI factor & Short & None & 1.55\% & 3.14\% & 20.57\% & 2.71\%\\
OI factor picks & \textasciitilde{} TS factor & Long & High & -0.24\% & 5.47\% & 22.98\% & 3.39\%\\
OI factor picks & \textasciitilde{} TS factor & Long & Low & 2.1\% & 15.43\% & 16.87\% & 0.23\%\\
OI factor picks & \textasciitilde{} TS factor & Long & None & 0.74\% & 10.28\% & 20.88\% & 1.57\%\\
OI factor picks & \textasciitilde{} TS factor & Short & High & -0.35\% & 1.99\% & 10.81\% & 8.45\%\\
OI factor picks & \textasciitilde{} TS factor & Short & Low & 0.56\% & 1.8\% & 8.96\% & 0.35\%\\
OI factor picks & \textasciitilde{} TS factor & Short & None & 0.16\% & 2.25\% & 10.32\% & 4.29\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & High & 0.9\% & 10.35\% & 12.52\% & 5.81\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & Low & 3.58\% & 28.68\% & 29.36\% & 11.92\%\\
CHP factor picks & \textasciitilde{} CHP factor & Long & None & 1.87\% & 18.04\% & 18.32\% & 8.25\%\\
CHP factor picks & \textasciitilde{} CHP factor & Short & High & 4.73\% & 21.64\% & 25.57\% & 13.39\%\\
CHP factor picks & \textasciitilde{} CHP factor & Short & Low & 0.15\% & 22.77\% & 21.53\% & 7.38\%\\
CHP factor picks & \textasciitilde{} CHP factor & Short & None & 2.08\% & 22.85\% & 24.23\% & 10.59\%\\
CHP factor picks & \textasciitilde{} Market factor & Long & High & 4.22\% & 18.7\% & 26.41\% & 15.02\%\\
CHP factor picks & \textasciitilde{} Market factor & Long & Low & 4.4\% & 30.16\% & 27.34\% & 14.77\%\\
CHP factor picks & \textasciitilde{} Market factor & Long & None & 4.48\% & 25.09\% & 26.83\% & 14.99\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & High & 0.6\% & 1.73\% & 19.04\% & 8.75\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & Low & 1.32\% & 21.86\% & 29.51\% & 16.19\%\\
CHP factor picks & \textasciitilde{} OI factor & Long & None & 1.03\% & 9.08\% & 22.58\% & 11.61\%\\
CHP factor picks & \textasciitilde{} OI factor & Short & High & 2.03\% & 21.75\% & 17.79\% & 8.86\%\\
CHP factor picks & \textasciitilde{} OI factor & Short & Low & 3.48\% & 23.02\% & 30.31\% & 3.59\%\\
CHP factor picks & \textasciitilde{} OI factor & Short & None & 2.83\% & 22.98\% & 21.6\% & 6.75\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & High & 1.55\% & 24.47\% & 16.98\% & 4.23\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & Low & 2.4\% & 29.72\% & 21.59\% & 4.46\%\\
CHP factor picks & \textasciitilde{} TS factor & Long & None & 1.91\% & 27.37\% & 18.54\% & 5.2\%\\
CHP factor picks & \textasciitilde{} TS factor & Short & High & 2.04\% & 2.25\% & 21.16\% & 11.49\%\\
CHP factor picks & \textasciitilde{} TS factor & Short & Low & 0.13\% & 15.94\% & 28.57\% & 15.29\%\\
CHP factor picks & \textasciitilde{} TS factor & Short & None & 1.15\% & 8.14\% & 23.54\% & 12.87\%\\
Market factor picks & \textasciitilde{} CHP factor & Long & High & 0.14\% & 14.47\% & 24.85\% & 13.11\%\\
Market factor picks & \textasciitilde{} CHP factor & Long & Low & 1.14\% & 19\% & 33.01\% & 7.1\%\\
Market factor picks & \textasciitilde{} CHP factor & Long & None & 0.66\% & 17\% & 27.85\% & 10.89\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & High & -0.19\% & 1.81\% & 31.9\% & 5.71\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & Low & 1.53\% & 3.86\% & 35.32\% & -0.39\%\\
Market factor picks & \textasciitilde{} CHP factor & Short & None & 0.66\% & 2.85\% & 32.76\% & 2.67\%\\
Market factor picks & \textasciitilde{} Market factor & Long & High & -0.11\% & 8.95\% & 36.58\% & 11.77\%\\
Market factor picks & \textasciitilde{} Market factor & Long & Low & 2.08\% & 9.85\% & 29.63\% & 2.79\%\\
Market factor picks & \textasciitilde{} Market factor & Long & None & 0.88\% & 9.93\% & 33.73\% & 7.69\%\\
Market factor picks & \textasciitilde{} OI factor & Long & High & -0.32\% & 2.73\% & 34.04\% & 7.23\%\\
Market factor picks & \textasciitilde{} OI factor & Long & Low & 0.07\% & 5.36\% & 35.07\% & 1.56\%\\
Market factor picks & \textasciitilde{} OI factor & Long & None & 0.07\% & 4.4\% & 34.38\% & 4.78\%\\
Market factor picks & \textasciitilde{} OI factor & Short & High & -0.03\% & 6.86\% & 24.78\% & 9.62\%\\
Market factor picks & \textasciitilde{} OI factor & Short & Low & 2.28\% & 9.51\% & 27.82\% & 1.38\%\\
Market factor picks & \textasciitilde{} OI factor & Short & None & 0.43\% & 8.38\% & 25.68\% & 6.05\%\\
Market factor picks & \textasciitilde{} TS factor & Long & High & -0.03\% & 11.75\% & 30.86\% & 8.57\%\\
Market factor picks & \textasciitilde{} TS factor & Long & Low & 0.84\% & 8.74\% & 30.58\% & 4.49\%\\
Market factor picks & \textasciitilde{} TS factor & Long & None & 0.48\% & 10.83\% & 30.76\% & 7.34\%\\
Market factor picks & \textasciitilde{} TS factor & Short & High & -0.43\% & 2.19\% & 25.29\% & 6.6\%\\
Market factor picks & \textasciitilde{} TS factor & Short & Low & 2.06\% & 2.77\% & 24.04\% & 0.2\%\\
Market factor picks & \textasciitilde{} TS factor & Short & None & 0.53\% & 2.81\% & 24.88\% & 3.76\%\\*
\end{longtabu}
\end{landscape}
\endgroup{}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{assets}{%
\subsection{Assets}\label{assets}}

\begingroup\fontsize{10}{12}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:`appendix - assets`}This table shows details on the individual commodities considered in this study including trading country, sector, subsector as well as trading exchange Market Identifier Codes (MIC, ISO 10383).}\\
\toprule
Country & Sector & Subsector & Exchange & Name\\
\midrule
US & Agriculturals & Grains & XCBT & Corn\\
US & Agriculturals & Grains & XCBT & Oats\\
US & Agriculturals & Grains & XCBT & Soy-meal\\
US & Agriculturals & Grains & XCBT & Soy-oil\\
US & Agriculturals & Grains & XCBT & Soybeans\\
US & Agriculturals & Grains & XCBT & Wheat-SRW\\
US & Agriculturals & Livestock & XCME & Feeder cattle\\
US & Agriculturals & Livestock & XCME & Lean hogs\\
US & Agriculturals & Livestock & XCME & Live cattle\\
US & Agriculturals & Softs & IFUS & Cocoa\\
US & Agriculturals & Softs & IFUS & Coffee\\
US & Agriculturals & Softs & IFUS & Cotton\\
US & Agriculturals & Softs & IFUS & Orange juice\\
US & Agriculturals & Softs & IFUS & Sugar\\
US & Agriculturals & Softs & XCME & Lumber\\
US & Energy & Petroleum & XNYM & Crude oil\\
US & Energy & Petroleum & XNYM & Gasoline\\
US & Energy & Petroleum & XNYM & Heating oil\\
US & Energy & Petroleum & XNYM & Natural gas\\
US & Metals & Industrial & XCEC & Copper\\
US & Metals & Precious & XCEC & Gold\\
US & Metals & Precious & XCEC & Silver\\
US & Metals & Precious & XNYM & Palladium\\
US & Metals & Precious & XNYM & Platinum\\
UK & Metals & Industrial & XLME & Aluminium\\
UK & Metals & Industrial & XLME & Copper\\
UK & Metals & Industrial & XLME & Lead\\
UK & Metals & Industrial & XLME & Nickel\\
UK & Metals & Industrial & XLME & Tin\\
UK & Metals & Industrial & XLME & Zinc\\
\bottomrule
\end{longtabu}
\endgroup{}

\newpage

\hypertarget{market-factor-lemma}{%
\subsection{Market factor lemma}\label{market-factor-lemma}}

\[\overline{R_{r_{mkt}, p}^{2}} = \frac{1}{n} \sum_{j} R_{r_{j}, r_{mkt}, p}^{2} = \frac{1}{n}\]
\(\overline{R_{mkt, p}^{2}}\equiv\) period \(p\) average \(R^{2}\) for
market risk factor returns.\\
\(R_{j, mkt, p}^{2}\equiv\) period \(p\) \(R^{2}\) for the regression
where the dependent variable is the vector of returns on commodity \(j\)
nearby contract and the regressor is the vector of returns on the market
portfolio; \(j \in Y_{p}^{mkt}\).\\
\(Y_{p}^{mkt}\equiv\) commodity set considered for the formation of the
mimicking portfolio for market risk over period \(p\).\\
\(n\equiv\) number of commodities in \(Y_{p}^{mkt}\).

\textbf{Proof}

Week \(t\) market factor return:
\[r_{mkt, t} = \sum_{j} w_{j, p} \cdot r_{j, t}\] \(w_{j, p}\equiv\)
period \(p\) weight for commodity \(j\) in the mimicking portfolio for
market risk in returns.\\
\(r_{j, t}\equiv\) week \(t\) return for commodity \(j\).

Assumptions:

\begin{math}
  \begin{aligned}
   w_{j, p} &= w_{k, p} = \frac{1}{n} \: \forall \: \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt} \\
   \sigma_{r_{j}, p} &= \sigma_{r_{k}, p} \: \forall \: \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt} \\
   cov \left ( r_{j}, r_{k} \right )_{p} &= 0 \: \forall \left \{ j, \: k \right \}; \: j, \: k \: \in Y_{p}^{mkt}
  \end{aligned}
\end{math}

\newpage

\[
\begin{array}{l c l}
  \sigma_{r_{j, p}}^{2} & = & \beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2} + \sigma_{\epsilon}^{2} \\
  \beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2} & = & \sigma_{r_{j, p}}^{2} - \sigma_{\epsilon}^{2} \\
  \frac{\beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2}}{\sigma_{r_{j, p}}^{2}} & = & \frac{\sigma_{r_{j, p}}^{2} - \sigma_{\epsilon}^{2}}{\sigma_{r_{j, p}}^{2}} \\
    \frac{\beta_{r_{j}, r_{mkt}, p}^{2} \cdot \sigma_{r_{mkt, p}}^{2}}{\sigma_{r_{j, p}}^{2}}  & = & R_{r_{j}, r_{mkt}, p}^{2} \\ \\
  \sigma_{r_{mkt, p}}^{2} & = & \sum_{j} w_{j, p}^{2} \cdot \sigma_{j, p}^{2} + 2 \cdot \sum_{j} \sum_{k \neq j} w_{j, p} \cdot w_{k, p} \cdot cov \left ( r_{j, p}, r_{k, p} \right ) \\
                          & = & \sum_{j} \left ( \frac{1}{n} \right )^{2} \cdot \sigma_{j, p}^{2} \\
                          & = & \frac{\sum_{j} \sigma_{j, p}^{2}}{n^{2}} \\ \\
  cov \left ( r_{j}, r_{mkt} \right )_{p} & = & cov \left ( r_{j}, \frac{1}{n} \sum_{k} r_{k} \right )_{p} \\
                                          & = & \frac{1}{n} \sum_{k} cov \left ( r_{j}, r_{k} \right )_{p} \\
                                          & = & \frac{1}{n} cov \left ( r_{j}, r_{j} \right )_{p} \\
                                          & = & \frac{\sigma_{r_{j}, p}^{2}}{n} \\ \\
  \beta_{j, mkt, p}^{2} & = & \left [ \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}}{\sigma_{r_{mkt, p}}^{2}} \right ]^{2} \\
                        & = & \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}^{2}}{\sigma_{r_{mkt, p}}^{4}} \\ \\
  \beta_{j, mkt, p}^{2} \cdot  \sigma_{r_{mkt, p}}^{2}  & = & \frac{cov \left ( r_{j}, r_{mkt} \right )_{p}^{2}}{\sigma_{r_{mkt, p}}^{2}} \\
                                                        & = & \left ( \frac{\sigma_{r_{j}, p}^{2}}{n} \right )^{2} \cdot \frac{1}{\sigma_{r_{mkt, p}}^{2}} \\
                                                        & = & \frac{\sigma_{r_{j}, p}^{4}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\ \\
\end{array}
\]

\newpage

\[
\begin{array}{l c l}
  R_{r_{j}, r_{mkt}, p}^{2} & = & \frac{\sigma_{r_{j}, p}^{4}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \cdot \frac{1}{\sigma_{r_{j, p}}^{2}} \\
                            & = & \frac{\sigma_{r_{j}, p}^{2}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\ \\  
                            & \therefore  & \\ \\
  \overline{R_{r_{mkt}, p}^{2}} & = & \frac{1}{n} \sum_{j} \frac{\sigma_{r_{j}, p}^{2}}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \\
                                & = & \frac{1}{n} \cdot \frac{1}{n^{2} \cdot \sigma_{r_{mkt, p}}^{2}} \cdot \sum_{j} \sigma_{r_{j}, p}^{2} \\
                                & = & \frac{1}{n} \cdot \frac{\sum_{j} \sigma_{r_{j}, p}^{2}}{n^{2} \cdot \frac{1}{n^{2}} \cdot \sum_{j} \sigma_{r_{j}, p}^{2}} \\
                                & = & \frac{1}{n} \\
\end{array}
\]

\newpage

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-acharya_limits_2013}{}%
Acharya, V.V., Lochstoer, L.A., Ramadorai, T., 2013. Limits to arbitrage
and hedging: Evidence from commodity markets. Journal of Financial
Economics 109, 441--465.
doi:\href{https://doi.org/10.1016/j.jfineco.2013.03.003}{10.1016/j.jfineco.2013.03.003}

\leavevmode\hypertarget{ref-anderson_hedger_1983}{}%
Anderson, R.W., Danthine, J.-P., 1983. Hedger diversity in futures
markets. The Economic Journal 93, 370--389.
doi:\href{https://doi.org/10.2307/2232798}{10.2307/2232798}

\leavevmode\hypertarget{ref-baker_financialization_2014}{}%
Baker, S.D., 2014. The financialization of storable commodities. Finance
Down Under 2014 Building on the Best from the Cellars of Finance.
doi:\href{https://doi.org/10.2139/ssrn.2348333\%20}{10.2139/ssrn.2348333 }

\leavevmode\hypertarget{ref-basak_model_2016}{}%
Basak, S., Pavlova, A., 2016. A model of financialization of
commodities. The Journal of Finance 71, 1511--1556.
doi:\href{https://doi.org/10.1111/jofi.12408}{10.1111/jofi.12408}

\leavevmode\hypertarget{ref-basu_capturing_2013}{}%
Basu, D., Miffre, J., 2013. Capturing the risk premium of commodity
futures: The role of hedging pressure. Journal of Banking \& Finance 37,
2652--2664.
doi:\href{https://doi.org/10.1016/j.jbankfin.2013.02.031}{10.1016/j.jbankfin.2013.02.031}

\leavevmode\hypertarget{ref-bohl_does_2013}{}%
Bohl, M.T., Stephan, P.M., 2013. Does futures speculation destabilize
spot prices? New evidence for commodity markets. Journal of Agricultural
and Applied Economics 45, 595--616.
doi:\href{https://doi.org/10.1017/S1074070800005150}{10.1017/S1074070800005150}

\leavevmode\hypertarget{ref-boons_stock_2012}{}%
Boons, M., De Roon, F., Szymanowska, M., 2012. The stock market price of
commodity risk, in: AFA 2012 Chicago Meetings Paper.

\leavevmode\hypertarget{ref-boyd_prevalence_2016}{}%
Boyd, N.E., Büyükşahin, B., Haigh, M.S., Harris, J.H., 2016. The
prevalence, sources, and effects of herding. Journal of Futures Markets
36, 671--694.
doi:\href{https://doi.org/10.1002/fut.21756}{10.1002/fut.21756}

\leavevmode\hypertarget{ref-brunetti_speculators_2016}{}%
Brunetti, C., Büyükşahin, B., Harris, J.H., 2016. Speculators, prices,
and market volatility. Journal of Financial and Quantitative Analysis
51, 1545--1574.
doi:\href{https://doi.org/10.1017/S0022109016000569}{10.1017/S0022109016000569}

\leavevmode\hypertarget{ref-brunetti_commodity_2014}{}%
Brunetti, C., Reiffen, D., 2014. Commodity index trading and hedging
costs. Journal of Financial Markets 21, 153--180.
doi:\href{https://doi.org/10.1016/j.finmar.2014.08.001}{10.1016/j.finmar.2014.08.001}

\leavevmode\hypertarget{ref-bruno_financialization_2017}{}%
Bruno, V.G., Büyükşahin, B., Robe, M.A., 2017. The financialization of
food? American Journal of Agricultural Economics 99, 243--264.
doi:\href{https://doi.org/10.1093/ajae/aaw059}{10.1093/ajae/aaw059}

\leavevmode\hypertarget{ref-buyuksahin_speculators_2011}{}%
Büyükşahin, B., Harris, J.H., 2011. Do speculators drive crude oil
futures prices? The Energy Journal 167--202.

\leavevmode\hypertarget{ref-buyuksahin_speculators_2014}{}%
Büyükşahin, B., Robe, M.A., 2014. Speculators, commodities and
cross-market linkages. Journal of International Money and Finance,
Understanding international commodity price fluctuations 42, 38--70.
doi:\href{https://doi.org/10.1016/j.jimonfin.2013.08.004}{10.1016/j.jimonfin.2013.08.004}

\leavevmode\hypertarget{ref-chang_returns_1985}{}%
Chang, E.C., 1985. Returns to speculators and the theory of normal
backwardation. The Journal of Finance 40, 193--208.
doi:\href{https://doi.org/10.1111/j.1540-6261.1985.tb04944.x}{10.1111/j.1540-6261.1985.tb04944.x}

\leavevmode\hypertarget{ref-cheng_convective_2014}{}%
Cheng, I.-H., Kirilenko, A., Xiong, W., 2014. Convective risk flows in
commodity futures markets. Review of Finance 19, 1733--1781.
doi:\href{https://doi.org/10.1093/rof/rfu043}{10.1093/rof/rfu043}

\leavevmode\hypertarget{ref-cheng_financialization_2014}{}%
Cheng, I.-H., Xiong, W., 2014. Financialization of commodity markets.
Annual Review of Financial Economics 6, 419--441.
doi:\href{https://doi.org/10.1146/annurev-financial-110613-034432}{10.1146/annurev-financial-110613-034432}

\leavevmode\hypertarget{ref-christoffersen_factor_2014}{}%
Christoffersen, P., Lunde, A., Olesen, K.V., 2014. Factor structure in
commodity futures return and volatility. Journal of Financial and
Quantitative Analysis 1--74.
doi:\href{https://doi.org/10.1017/S0022109018000765}{10.1017/S0022109018000765}

\leavevmode\hypertarget{ref-cootner_returns_1960}{}%
Cootner, P.H., 1960. Returns to speculators: Telser versus keynes.
Journal of political Economy 68, 396--404.
doi:\href{https://doi.org/10.1086/258347}{10.1086/258347}

\leavevmode\hypertarget{ref-cootner_speculation_1967}{}%
Cootner, P.H., 1967. Speculation and hedging. Food Research Institute
Studies 7, 65.

\leavevmode\hypertarget{ref-danthine_information_1978}{}%
Danthine, J.-P., 1978. Information, futures prices, and stabilizing
speculation. Journal of Economic Theory 17, 79--98.
doi:\href{https://doi.org/10.1016/0022-0531(78)90124-2}{10.1016/0022-0531(78)90124-2}

\leavevmode\hypertarget{ref-deschutter_food_2010}{}%
De Schutter, O., 2010. Food commodities speculation and food price
crises: Regulation to reduce the risks of price volatility. United
Nations Special Rapporteur on the Right to Food Briefing Note 2, 1--14.

\leavevmode\hypertarget{ref-domanski_financial_2007}{}%
Domanski, D., Heath, A., 2007. Financial investors and commodity
markets. Working Paper.

\leavevmode\hypertarget{ref-duffie_challenges_2014}{}%
Duffie, D., 2014. Challenges to a policy treatment of sSpeculative
trading motivated by differences in beliefs. The Journal of Legal
Studies 43, S173--S182.
doi:\href{https://doi.org/10.1086/677836}{10.1086/677836}

\leavevmode\hypertarget{ref-dusak_futures_1973}{}%
Dusak, K., 1973. Futures trading and investor returns: An investigation
of commodity market risk premiums. Journal of Political economy 81,
1387--1406. doi:\href{https://doi.org/10.1086/260133}{10.1086/260133}

\leavevmode\hypertarget{ref-ederington_who_2002}{}%
Ederington, L., Lee, J.H., 2002. Who trades futures and how: Evidence
from the heating oil futures market. The Journal of Business 75,
353--373. doi:\href{https://doi.org/10.1086/338706}{10.1086/338706}

\leavevmode\hypertarget{ref-ekeland_speculation_2016}{}%
Ekeland, I., Lautier, D., Villeneuve, B., 2016. Speculation in commodity
futures markets: A simple equilibrium model.

\leavevmode\hypertarget{ref-etula_broker-dealer_2013}{}%
Etula, E., 2013. Broker-dealer risk appetite and commodity returns.
Journal of Financial Econometrics 11, 486--521.
doi:\href{https://doi.org/10.1093/jjfinec/nbs024}{10.1093/jjfinec/nbs024}

\leavevmode\hypertarget{ref-fattouh_role_2013}{}%
Fattouh, B., Kilian, L., Mahadeva, L., 2013. The role of speculation in
oil markets: What have we learnt so far? The Energy Journal 34, 7--33.

\leavevmode\hypertarget{ref-fuertes_commodity_2015}{}%
Fuertes, A.-M., Miffre, J., Fernandez-Perez, A., 2015. Commodity
strategies based on momentum, term structure, and idiosyncratic
volatility. Journal of Futures Markets 35, 274--297.
doi:\href{https://doi.org/10.1002/fut.21656}{10.1002/fut.21656}

\leavevmode\hypertarget{ref-gilbert_how_2010}{}%
Gilbert, C.L., 2010a. How to understand high food prices. Journal of
Agricultural Economics 61, 398--425.
doi:\href{https://doi.org/10.1111/j.1477-9552.2010.00248.x}{10.1111/j.1477-9552.2010.00248.x}

\leavevmode\hypertarget{ref-gilbert_speculative_2010}{}%
Gilbert, C.L., 2010b. Speculative influences on commodity futures prices
2006-2008. UNCTAD, Geneva.

\leavevmode\hypertarget{ref-gilbert_role_2014}{}%
Gilbert, C.L., Pfuderer, S., 2014. The role of index trading in price
formation in the grains and oilseeds markets. Journal of Agricultural
Economics 65, 303--322.
doi:\href{https://doi.org/10.1111/1477-9552.12068}{10.1111/1477-9552.12068}

\leavevmode\hypertarget{ref-goldstein_speculation_2014}{}%
Goldstein, I., Li, Y., Yang, L., 2014. Speculation and hedging in
segmented markets. The Review of Financial Studies 27, 881--922.
doi:\href{https://doi.org/10.1093/rfs/hht059}{10.1093/rfs/hht059}

\leavevmode\hypertarget{ref-goldstein_commodity_2017}{}%
Goldstein, I., Yang, L., 2017. Commodity financialization and
information transmission (SSRN Scholarly Paper No. ID 2555996). Social
Science Research Network, Rochester, NY.
doi:\href{https://doi.org/10.2139/ssrn.2555996}{10.2139/ssrn.2555996}

\leavevmode\hypertarget{ref-hamilton_effects_2015}{}%
Hamilton, J.D., Wu, J.C., 2015. Effects of index-fund investing on
commodity futures prices. International Economic Review 56, 187--205.
doi:\href{https://doi.org/10.1111/iere.12099}{10.1111/iere.12099}

\leavevmode\hypertarget{ref-henderson_new_2015}{}%
Henderson, B.J., Pearson, N.D., Wang, L., 2015. New evidence on the
financialization of commodity markets. The Review of Financial Studies
28, 1285--1311.
doi:\href{https://doi.org/10.1093/rfs/hhu091}{10.1093/rfs/hhu091}

\leavevmode\hypertarget{ref-herman_not_2011}{}%
Herman, M.-O., Kelly, R., Nash, R., 2011. Not a game, speculation v.
Food security: Regulating financial markets to grow a better future.
Oxfam policy and practice: agriculture, food and land 11, 127--138.

\leavevmode\hypertarget{ref-hicks_value_1939}{}%
Hicks, J.R., 1939. Value and capital. Oxford University Press,
Cambridge.

\leavevmode\hypertarget{ref-hirshleifer_risk_1988}{}%
Hirshleifer, D., 1988. Risk, futures pricing, and the organization of
production in commodity markets. Journal of Political Economy 96,
1206--1220. doi:\href{https://doi.org/10.1086/261584}{10.1086/261584}

\leavevmode\hypertarget{ref-hirshleifer_determinants_1989}{}%
Hirshleifer, D., 1989. Determinants of hedging and risk premia in
commodity futures markets. Journal of Financial and Quantitative
Analysis 24, 313--331.
doi:\href{https://doi.org/10.2307/2330814}{10.2307/2330814}

\leavevmode\hypertarget{ref-hirshleifer_hedging_1990}{}%
Hirshleifer, D., 1990. Hedging pressure and futures price movements in a
general equilibrium model. Econometrica 58, 411--428.
doi:\href{https://doi.org/10.2307/2938209}{10.2307/2938209}

\leavevmode\hypertarget{ref-hong_what_2012}{}%
Hong, H., Yogo, M., 2012. What does futures market interest tell us
about the macroeconomy and asset prices? Journal of Financial Economics
105, 473--490.
doi:\href{https://doi.org/10.1016/j.jfineco.2012.04.005}{10.1016/j.jfineco.2012.04.005}

\leavevmode\hypertarget{ref-irwin_commodity_2013}{}%
Irwin, S.H., 2013. Commodity index investment and food prices: Does the
``masters hypothesis'' explain recent price spikes? Agricultural
Economics 44, 29--41.
doi:\href{https://doi.org/10.1111/agec.12048}{10.1111/agec.12048}

\leavevmode\hypertarget{ref-irwin_index_2011}{}%
Irwin, S.H., Sanders, D.R., 2011. Index funds, financialization, and
commodity futures markets. Applied Economic Perspectives and Policy 33,
1--31.
doi:\href{https://doi.org/10.1093/aepp/ppq032}{10.1093/aepp/ppq032}

\leavevmode\hypertarget{ref-irwin_financialization_2012}{}%
Irwin, S.H., Sanders, D.R., 2012. Financialization and structural change
in commodity futures markets. Journal of Agricultural and Applied
Economics 44, 371--396.
doi:\href{https://doi.org/10.1017/S1074070800000481}{10.1017/S1074070800000481}

\leavevmode\hypertarget{ref-irwin_devil_2009}{}%
Irwin, S.H., Sanders, D.R., Merrin, R.P., 2009. Devil or angel? The role
of speculation in the recent commodity price boom (and bust). Journal of
Agricultural and Applied Economics 41, 377--391.
doi:\href{https://doi.org/10.1017/S1074070800002856}{10.1017/S1074070800002856}

\leavevmode\hypertarget{ref-juvenal_speculation_2015}{}%
Juvenal, L., Petrella, I., 2015. Speculation in the oil market. Journal
of Applied Econometrics 30, 621--649.
doi:\href{https://doi.org/10.1002/jae.2388}{10.1002/jae.2388}

\leavevmode\hypertarget{ref-kang_tale_2017}{}%
Kang, W., Rouwenhorst, K.G., Tang, K., 2017. A tale of two premiums: The
role of hedgers and speculators in commodity futures markets.
doi:\href{https://doi.org/10.2139/ssrn.2449315}{10.2139/ssrn.2449315}

\leavevmode\hypertarget{ref-keynes_treatise_1930}{}%
Keynes, J.M., 1930. Treatise on money. Macmillan, London.

\leavevmode\hypertarget{ref-kilian_role_2014}{}%
Kilian, L., Murphy, D.P., 2014. The role of inventories and speculative
trading in the global market for crude oil. Journal of Applied
Econometrics 29, 454--478.
doi:\href{https://doi.org/10.1002/jae.2322}{10.1002/jae.2322}

\leavevmode\hypertarget{ref-kim_does_2015}{}%
Kim, A., 2015. Does futures speculation destabilize commodity markets?
Journal of Futures Markets 35, 696--714.
doi:\href{https://doi.org/10.1002/fut.21716}{10.1002/fut.21716}

\leavevmode\hypertarget{ref-knittel_simple_2016}{}%
Knittel, C.R., Pindyck, R.S., 2016. The simple economics of commodity
price speculation. American Economic Journal: Macroeconomics 8, 85--110.
doi:\href{https://doi.org/10.1257/mac.20140033}{10.1257/mac.20140033}

\leavevmode\hypertarget{ref-kolb_is_1992}{}%
Kolb, R.W., 1992. Is normal backwardation normal? Journal of Futures
Markets 12, 75--91.

\leavevmode\hypertarget{ref-leclercq_equilibrium_2014}{}%
Leclercq, E., Praz, R., 2014. Equilibrium commodity trading (SSRN
Scholarly Paper No. ID 2464400). Social Science Research Network,
Rochester, NY.
doi:\href{https://doi.org/10.2139/ssrn.2464400}{10.2139/ssrn.2464400}

\leavevmode\hypertarget{ref-manera_modelling_2016}{}%
Manera, M., Nicolini, M., Vignati, I., 2016. Modelling futures price
volatility in energy markets: Is there a role for financial speculation?
Energy Economics, Energy markets 53, 220--229.
doi:\href{https://doi.org/10.1016/j.eneco.2014.07.001}{10.1016/j.eneco.2014.07.001}

\leavevmode\hypertarget{ref-masters_testimony_2008}{}%
Masters, M.W., 2008. Testimony before the committee on homeland security
and governmental affairs. U.S. Senate 20.

\leavevmode\hypertarget{ref-masters_accidental_2008}{}%
Masters, M., White, A., 2008. The accidental hunt brothers: How
institutional investors are driving up food and energy prices. Working
paper.

\leavevmode\hypertarget{ref-moskowitz_time_2012}{}%
Moskowitz, T.J., Ooi, Y.H., Pedersen, L.H., 2012. Time series momentum.
Journal of Financial Economics, Special issue on investor sentiment 104,
228--250.
doi:\href{https://doi.org/10.1016/j.jfineco.2011.11.003}{10.1016/j.jfineco.2011.11.003}

\leavevmode\hypertarget{ref-sanders_impact_2011}{}%
Sanders, D.R., Irwin, S.H., 2011a. The impact of index funds in
commodity futures markets: A systems approach. The Journal of
Alternative Investments 14, 40--49.

\leavevmode\hypertarget{ref-sanders_new_2011}{}%
Sanders, D.R., Irwin, S.H., 2011b. New evidence on the impact of index
funds in u.s. Grain futures markets. Canadian Journal of Agricultural
Economics/Revue canadienne d'agroeconomie 59, 519--532.
doi:\href{https://doi.org/10.1111/j.1744-7976.2011.01226.x}{10.1111/j.1744-7976.2011.01226.x}

\leavevmode\hypertarget{ref-sanders_adequacy_2010}{}%
Sanders, D.R., Irwin, S.H., Merrin, R.P., 2010. The adequacy of
speculation in agricultural futures markets: Too much of a good thing?
Applied Economic Perspectives and Policy 32, 77--94.
doi:\href{https://doi.org/10.1093/aepp/ppp006}{10.1093/aepp/ppp006}

\leavevmode\hypertarget{ref-schumann_hunger_2011}{}%
Schumann, H., 2011. The hunger-makers: How deutsche bank, goldman sachs
and other financial institutions are speculating with food at the
expense of the poorest. Foodwatch, Berlin.

\leavevmode\hypertarget{ref-senate_excessive_2009}{}%
Senate, U., 2009. Excessive speculation in the wheat market. Majority
and Minority Staff Report. Permanent Subcommittee on Investigations 24,
107--108.

\leavevmode\hypertarget{ref-simsek_speculation_2013}{}%
Simsek, A., 2013. Speculation and risk sharing with new financial
assets. The Quarterly Journal of Economics 128, 1365--1396.
doi:\href{https://doi.org/10.1093/qje/qjt007}{10.1093/qje/qjt007}

\leavevmode\hypertarget{ref-singleton_investor_2013}{}%
Singleton, K.J., 2013. Investor flows and the 2008 boom/bust in oil
prices. Management Science 60, 300--318.
doi:\href{https://doi.org/10.1287/mnsc.2013.1756}{10.1287/mnsc.2013.1756}

\leavevmode\hypertarget{ref-sockin_informational_2015}{}%
Sockin, M., Xiong, W., 2015. Informational frictions and commodity
markets. The Journal of Finance 70, 2063--2098.
doi:\href{https://doi.org/10.1111/jofi.12261}{10.1111/jofi.12261}

\leavevmode\hypertarget{ref-stoll_commodity_2011}{}%
Stoll, H.R., Whaley, R.E., 2011. Commodity index investing: Speculation
or diversification? The Journal of Alternative Investments 14, 50--60.
doi:\href{https://doi.org/10.2139/ssrn.1633908}{10.2139/ssrn.1633908}

\leavevmode\hypertarget{ref-stout_why_1998}{}%
Stout, L.A., 1998. Why the law hates speculators: Regulation and private
ordering in the market for o.t.c. Derivatives. Duke Law Journal 48,
701--786.

\leavevmode\hypertarget{ref-stulz_rethink_1996}{}%
Stulz, R.M., 1996. Rethinking risk management. Journal of Applied
Corporate Finance 9, 8--25.
doi:\href{https://doi.org/10.1111/j.1745-6622.1996.tb00295.x}{10.1111/j.1745-6622.1996.tb00295.x}

\leavevmode\hypertarget{ref-szymanowska_anatomy_2014}{}%
Szymanowska, M., De Roon, F., Nijman, T., Van Den Goorbergh, R., 2014.
An anatomy of commodity futures risk premia. The Journal of Finance 69,
453--482.
doi:\href{https://doi.org/10.1111/jofi.12096}{10.1111/jofi.12096}

\leavevmode\hypertarget{ref-tang_index_2012}{}%
Tang, K., Xiong, W., 2012. Index investment and the financialization of
commodities. Financial Analysts Journal 68, 54--74.
doi:\href{https://doi.org/10.2469/faj.v68.n6.5}{10.2469/faj.v68.n6.5}

\leavevmode\hypertarget{ref-till_long_2007}{}%
Till, H., 2007. A long-term perspective on commodity futures returns,
in: Intelligent Commodity Investing. Risk books, London.

\leavevmode\hypertarget{ref-tokic_speculation_2012}{}%
Tokic, D., 2012. Speculation and the 2008 oil bubble: The d.c.o.t.
Report analysis. Energy Policy 45, 541--550.
doi:\href{https://doi.org/10.1016/j.enpol.2012.02.069}{10.1016/j.enpol.2012.02.069}

\leavevmode\hypertarget{ref-unctad_global_2009}{}%
UNCTAD, S.T.F. on S.I., Cooperation, E., 2009. The global economic
crisis: Systemic failures and multilateral remedies. UNCTAD, Geneva.

\leavevmode\hypertarget{ref-working_theory_1948}{}%
Working, H., 1948. Theory of the inverse carrying charge in futures
markets. Journal of Farm Economics 30, 1--28.
doi:\href{https://doi.org/10.2307/1232678}{10.2307/1232678}

\leavevmode\hypertarget{ref-working_hedging_1953}{}%
Working, H., 1953. Hedging reconsidered. Journal of Farm Economics 35,
544--561. doi:\href{https://doi.org/10.2307/1233368}{10.2307/1233368}

\end{document}