### 2024-06-11 others/Paper highlights.Rmd

---
title: "Results highlights"
author: "Olivier Bauthéac"
date: ""
output: html_document
---

# Methods
* We consider the entire cross-section of liquid commodities (24 US, 6 UK) in 
  contrast to most other studies that focus on individual commodities or 
  sub-sectors.
* We use a futures-based asset pricing framework that includes factors 
  constructed using both price (market, term structure) and non-price (CHP, 
  open interest) attributes.   
  Helps isolating common driving factors and detecting co-movement.
* We extend the Keynesian individual hedging pressure paradigm by incorporating 
  aggregate CHP (**CHP**), a market wide measure of hedging pressure first 
  considered in @hong_what_2012.

# Results
* A non-price based factor is best able to detect the effects of commodity 
  financialization.
* The effects of the financial crisis and its aftermath are most clearly visible 
  through the performance of the market factor. 
* The effects of financialization seem to go beyond mechanical effects induced 
  by the rise of indexing.

* Over the first period the cross sectional average returns during 
  **backwardation** for both the US and UK commodities are positive with those 
  for the US statistically significant, and significantly higher than during 
  **contango**, providing broad support for an aggregate version of the 
  Keynesian hedging pressure hypothesis. Over the second period this pattern 
  reverses, with both the US and UK commodities having significant positive 
  returns during **contango** with these returns being significantly higher 
  than during **backwardation**. This change in pattern suggests the onset of 
  financialization altered the traditional risk shifting nature of hedger’s 
  behaviour, implied by the Keynesian hypothesis.
  Financial investors may have been liquidity demanders, leading hedgers to 
  become liquidity providers: because financial investors tend to be long only 
  the are liquidity demanders on the long side. Hedgers could therefore have 
  provided the required liquidity by being short as a group (**contango**) and 
  been compensated correspondingly (positive returns).

* Factors constructed with US commos, performance analysis on US commos.
  + The CHP factor shows the greatest improvement in performance in the second 
    period. 
  + The dramatic improvement of the CHP factor in the second period is driven by 
    precious and precious-base metals.  
  + The best performing commodities for the CHP factor exhibited very low 
    pair-wise correlation in the first period, around the average for US traded 
    commodities, while during financialization the average pairwise correlation 
    between the top performing CHP factor commodities is higher than that for 
    the market factor and double of the average for all the US traded 
    commodities.
  + This increase is due to the dramatic rise in average pairwise correlation 
    amongst the US traded metals all of which exhibited high R^2^ with respect 
    to the CHP factor during financialization which in turn is due to their 
    common low level of CHP over this period.  
  + The CHP factor model is therefore able to establish the strengthening link 
    between a non-price attribute and price dynamics in the US metal sub-sector 
    during financialization:

* Factors constructed with the top eight commodities for a given factor, 
  performance analysis on commodity pool used for construction + on US commos.  
  + Results strongly suggest the hedging pressure co-movement and price 
    co-movement link was concentrated in the metals sub-sector.

* Factors constructed with the top eight commodities for a given factor, 
  performance analysis on UK metals.  
  + Results show there was co-movement in global (as opposed to US only) metals 
    returns in the second period and that this co-movement could be detected by 
    our hedging pressure based model.

* Pattern of return for the US traded commodities across **CHP** regimes 
  reverted to Keynesian paradigm over the crisis and aftermath period.

* Pricing results indicate the emergence of a systematic factor across the 
  entire cross-section of the US commodities as well as strong evidence of 
  cross-market linkages.

# Conclusions
* Financialization was a phenomenon endogenous to the commodity markets 
  transmitted via the commodity futures markets and mainly concentrated in a 
  specific sector (metals).
* The crisis and its aftermath, in contrast, seems to have delivered an 
  exogenous shock across the entire cross-section of global commodities.