### 2024-06-11 others/paper/custom/title_page.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[letterpaper, left = 1in, right = 1in, top = 1.5in, bottom = 1.5in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Financialization of Commodities: an Asset Pricing Perspective},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{jf}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Financialization of Commodities: an Asset Pricing Perspective}
    % \author{Devraj Basu \\ Olivier Bauthéac}
  \author{ Devraj Basu  \\  Olivier Bauthéac  
	  \thanks{ Basu (corresponding author): University of Strathclyde, Glasgow, UK, \href{mailto:devraj.basu@strath.ac.uk}{\nolinkurl{devraj.basu@strath.ac.uk}}.  Bauthéac: University of Strathclyde, Glasgow, UK, \href{mailto:olivier.bautheac@strath.ac.uk}{\nolinkurl{olivier.bautheac@strath.ac.uk}}. 	  
	   \newline We thank seminar participants at IIM Ahmedabad, IIM Calcutta, SKEMA
Business School and University of Gottingen, as well as Griffith
inaugural alternative investment conference attendants for their
feedback. We also thank Joelle Miffre, Martijn Boons, John Fan, Robert
Bianchi, Chardin Wese Simen, Andrew Kaleel and particularly Olaf Korn
for their extremely useful comments and suggestions.  }
  }
    \date{}

\usepackage{longtable, tabu, pdflscape, booktabs, caption, setspace}
\doublespacing
\renewcommand*{\arraystretch}{0.65}

\begin{document}
\maketitle

\newpage

\textbf{Disclosure}

\begin{itemize}
\tightlist
\item
  Devraj Basu: I am a co-author of this paper and wish to state that I
  have no conflict of interest, financial or otherwise. I do not hold
  paid or unpaid positions at any organization whose goals relate to the
  article.\\
  \includegraphics{pics/signature - Devraj.png}\\
\end{itemize}

\newpage

\textbf{Disclosure}

\begin{itemize}
\tightlist
\item
  Olivier Bauthéac: I am a co-author of this paper and wish to state
  that I have no conflict of interest, financial or otherwise. I do not
  hold paid or unpaid positions at any organization whose goals relate
  to the article.\\
  \includegraphics{pics/signature - Olivier.png}\\
\end{itemize}

\newpage

\bibliography{BibTeX.bib}

\end{document}