### 2024-06-11 others/revision-jfm.qmd

---
title: "journal of futures markets - revision"
execute:
  echo: false
  message: false
  warning: false
format:
  html:
    page-layout: full
    toc: true
    toc-title: Contents
    toc-depth: 5
    theme: 
      - Minty
      - style.css
editor: visual
bibliography: references.bib
---

<style>
<!-- .main-container { -->
<!--   max-width: 95vw; /* Adjust this value as needed */ -->
<!--   margin: 0 auto; /* Center align the content */ -->
<!--   padding: 0; -->
<!-- } -->

#quarto-content {
  max-width: 95vw;
  margin: 0 auto;
  padding: 0;
}

.no-wrap-table th, .no-wrap-table td {
  white-space: nowrap;
}
</style>

```{r globals, echo=FALSE}
url_css <- "https://www.dropbox.com/scl/fi/hca7tog3830ir7ro9fszt/styles.css?rlkey=pma97454kb75qzuj9znv86zrf&dl=1"
url_references <- "https://www.dropbox.com/scl/fi/n7t8g3hn32d8e5gf2rcti/references.bib?rlkey=xy251n44q3y0t3omxpqk0dvyf&dl=1"
file_css <- "style.css"
file_references <- "references.bib"
urls <- c(url_css, url_references)
files <- c(file_css, file_references)

purrr::walk2(urls, files, ~download.file(.x, .y))

####
library(magrittr)

####
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

####
tables_collapse_rows_align <- "top"
tables_cut <- TRUE
tables_hline <- "major"

output_type <- "html"
```

# Correlation analysis

```{r correlations-load, echo=FALSE}
correlations <- readr::read_rds(paste0(here::here(), "/explore/tables/revision-jfm/correlations.rds"))
```

## Pairwise relative change correlations by period

```{r correlations-relative-periods, echo=FALSE}
filtered_correlations <- dplyr::filter(correlations, timespan == "period", type == "return") %>%
  dplyr::select(-c(timespan, year, type)) %>% dplyr::relocate(period, .before = regime) %>%
  dplyr::relocate(field, .before = frequency)

align <- "lllllllrllr"

knitr::kable(filtered_correlations, format = "html", table.attr = 'class="no-wrap-table"') %>% 
  kableExtra::kable_styling(full_width = TRUE) %>%
  kableExtra::scroll_box(width = "100%", height = "800px")
```