### 2024-08-21 others/revision-jfm - notes.Rmd
Analysis of new factor regressions results

---
title: "revision-jfm - notes"
output: html_document
date: '2024-08-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Factor regressions results overview ####

### All, all, all ####

#### First period ####

##### Market factor ####

- Best overall performance: fits story.
- Overall performance consistent across periods of high and low CHP: fits story.

##### CHP factor ####

- Lower performance relative to market factor: fits story.
- Higher performance in backwardation: fits overall story but not current text, rewrite.

##### Open interest nearby factor ####

- Lower performance relative to market factor but higher than CHP.
- Higher performance in contango.
- Not included in story yet, could potentially include, not sure how.

##### Open interest aggregate factor ####

- Lowest performance overall for the factor itself.
- Moderate performance (slightly lower than that for OI nearby) for both legs. Indicates that the legs are negatively correlated.
- Higher performance in contango.
- Not included in story yet, could potentially include, not sure how.

##### Term structure factor ####

- Similar pattern to OI aggregate.
- Not included in story yet, could potentially include, not sure how.

#### Second period ####

##### Market factor ####

- Substantial improvement in performance (19.70% vs. 9.69%) albeit not as strong as that for the CHP factor in a relative sense: fits story.
- Stronger performance in backwardation: not included in story.

##### CHP factor ####

- Strongest improvement in performance in relative sense (5.39% vs 1.93%). Improvement comes from long leg (16.95% vs. 5.77%), not as strong for short leg (10.49% vs. 5.41%): fits story.
- Stronger performance in backwardation: not included in story.

##### Open interest nearby factor ####

- No notable change in performance for the factor itself (5.63% vs 5.98%).
- Both legs show substantial improvement however (long: 14.18% vs 8.12%; short: 14.04% vs. 6.13%), indicating that they become negatively correlated.
- Stronger performance switches to backwardation.
- Not included in story, include?

##### Open interest aggregate factor ####

- Increase in performance similar in a relative sense for the factor itself and the legs indicating legs maintain similar level of negative correlation.
- Performance of the factor itself remains low.
- Not included in story, include?

##### Term structure factor ####

- Increase in performance (2.95%vs 1.52%) although performance remains low. Better performance switches to contango.
- More pronouced increase for the short leg (15.77% vs. 5.28%).
- Gap widens between performance of factor itselfs and that of legs; indicates legs become less positively correlated.
- Better performance switches to backwardation on both legs.