### 2024-09-16 docs/_extensions/quarto-journals/elsevier/partials/before-body.tex
Copy html output
To create an index.html file for deployment

\begin{frontmatter}
\title{$title$$if(subtitle)$ \\\large{$subtitle$} $endif$}
$for(by-author)$\author[$for(by-author.affiliations)$$it.number$$sep$,$endfor$]{$by-author.name.literal$%
$if(by-author.attributes.corresponding)$\corref{cor1}$endif$%
$if(by-author.note.text)$\fnref{fn$by-author.note.number$}$endif$}
$if(by-author.email)$ \ead{$by-author.email$} $endif$
$endfor$
$for(by-affiliation)$
\affiliation[$it.number$]{organization={$it.name$$if(it.department)$, $it.department$$endif$},$if(it.address)$addressline={$it.address$}$endif$$if(it.city)$$if(it.address)$,$endif$city={$it.city$}$endif$$if(it.country)$$if(it.city)$,$elseif(it.address)$,$endif$country={$it.country$},countrysep={,}$endif$$if(it.postal-code)$$if(it.country)$,$elseif(it.city)$,$elseif(it.address)$,$endif$postcode={$it.postal-code$}$endif$,postcodesep={}}
$endfor$

\cortext[cor1]{Corresponding author}
$for(by-author)$
$if(by-author.note.text)$\fntext[fn$by-author.note.number$]{$by-author.note.text$}$endif$
$endfor$

$if(abstract)$
\begin{abstract}
$abstract$
\end{abstract}
$endif$

$if(journal.graphical-abstract)$\begin{graphicalabstract}
$journal.graphical-abstract$
\end{graphicalabstract}
$endif$

$if(journal.highlights)$\begin{highlights}
$for(journal.highlights)$\item $it$$endfor$
\end{highlights}
$endif$

$if(keywords)$
\begin{keyword}
    $for(keywords/allbutlast)$$keywords$ \sep $endfor$
    $for(keywords/last)$$keywords$$endfor$
\end{keyword}
$endif$
\end{frontmatter}