### 2024-10-30 others/regime difference test results analysis.qmd
make regime difference test results analysis dashboard

---
title: "Analysis of regime difference test results"
format: 
  html:
    grid:
      body-width: 1000px
---

```{r}
#| label: globals
#| include: false

# Load required packages
library(dplyr)
library(readr)
library(here)
library(slituR)
library(BBGsymbols)
library(fewISOs)

# Load data
data("tickers_futures", "tickers_cftc", package = "BBGsymbols")
data("exchanges", package = "fewISOs")

results_directory_path <- here::here("explore", "results", "revision-jfm")
commodity_futures_tickers <- c(
  "BOA Comdty", "C A Comdty", "CCA Comdty", "CLA Comdty", "CTA Comdty", 
  "FCA Comdty", "GCA Comdty", "HGA Comdty", "HOA Comdty", "JOA Comdty", 
  "KCA Comdty", "LAA Comdty", "LBA Comdty", "LCA Comdty", "LHA Comdty", 
  "LLA Comdty", "LNA Comdty", "LPA Comdty", "LTA Comdty", "LXA Comdty", 
  "NGA Comdty", "O A Comdty", "PAA Comdty", "PLA Comdty", "S A Comdty", 
  "SBA Comdty", "SIA Comdty", "SMA Comdty", "W A Comdty", "XBWA Comdty",
  "LAA Comdty", "LPA Comdty", "LLA Comdty", "LNA Comdty", "LTA Comdty",
  "LXA Comdty"
)
period_levels <- c("past", "financialization", "crisis", "present")
sector_levels <- c("all", "agriculturals", "energy", "metals")
subsector_levels <- c("all", "grains", "livestock", "softs", "gas", "petroleum", "base", "precious")

results <- readr::read_rds(
  slituR::paste_forward_slash(results_directory_path, "regime-difference-tests.rds")
)
ticker_asset_map <- dplyr::left_join(
  dplyr::filter(tickers_futures, ticker %in% commodity_futures_tickers) %>% 
    dplyr::select(ticker, name, sector, subsector, MIC), 
  dplyr::select(exchanges, MIC, country), 
  by = "MIC"
)
results <- 
  dplyr::left_join(results, ticker_asset_map, by = c("ticker" = "ticker")) %>%
  dplyr::mutate(
    country = factor(country, levels = c("US", "GB")),
    moment = ifelse(moment == "mean", "mean", "volatility"),
    name = ifelse(is.na(name), ticker, name),
    name = ifelse(is.na(MIC), name, paste0(name, " (", MIC, ")")),
    period = factor(period, levels = period_levels),
    sector = factor(sector, levels = sector_levels),
    subsector = factor(subsector, levels = subsector_levels)
  ) %>%
  dplyr::arrange(country, sector, subsector, name, period) %>%
  dplyr::select(country, name, moment, period, `dominant regime`, `p-value`)

individuals <- dplyr::filter(results, !name %in% c("GB commodities", "US commodities"))
countries <- dplyr::filter(results, name %in% c("GB commodities", "US commodities")) %>%
  dplyr::select(-country)
```

## Individuals

### Numbers
```{r}
#| label: individuals-numbers
#| echo: false

summary_individuals <- group_by(individuals, country, period, moment) %>%
  summarise(
    `10%` = sum(`p-value` > 0.05 & `p-value` <= 0.1, na.rm = TRUE),
    `5%` = sum(`p-value` > 0.01 & `p-value` <= 0.05, na.rm = TRUE),
    `1%` = sum(`p-value` > 0 & `p-value` <= 0.01, na.rm = TRUE),
    .groups = "drop"
  )

knitr::kable(summary_individuals)
```

### Names
```{r}
#| label: individuals-names
#| echo: false

summary_individuals <- group_by(individuals, country, period, moment) %>%
  summarise(
    `10%` = list(name[`p-value` >= 0.05 & `p-value` < 0.1]),
    `5%` = list(name[`p-value` >= 0.01 & `p-value` < 0.05]),
    `1%` = list(name[`p-value` >= 0 & `p-value` < 0.01]),
    .groups = "drop"
  )

knitr::kable(summary_individuals)
```

## countries
```{r}
#| label: countries
#| echo: false

significance <- function(x) {
  case_when(
    x > 0.1 ~ "",
    x > 0.05 ~ "*",
    x > 0.01 ~ "**",
    TRUE ~ "***"
  )
}

summary_countries <- dplyr::mutate(
  countries, significance = significance(as.numeric(`p-value`))
  )

knitr::kable(summary_countries)
```